### 2024-04-03 Intro Memo/intro_memo.tex
Initial Memo Prior to First Meeting

\documentclass{article}

\usepackage{amsmath}
\usepackage{geometry}
\usepackage{graphicx}

\author{Gideon Moore}
\date{April 3, 2024}
\title{Introductory Memo: \\ Demand Estimation Using Movie Descriptions}

\begin{document}

\maketitle

\begin{abstract}
    Traditional BLP-based demand estimation requires structured characteristic data to identify cross-elasticities. Many markets, however, feature unstructured text descriptions of products without well-structured characteristics. Using modern text-to-data tools, I hope to estimate cross-elasticities in the presence of unstructured text alone. I exploit specific characteristics of the movie industry to pilot this methodology. Uniform pricing is ubiquitous in this market; thus, we can make arguments about demand shifts from quantity changes alone. Moreover, long and inflexible production pipelines have the potential to produce plausibly-exogeneous product characteristics; this argument is supported by the presence of ``twin films,'' movies released simultaneously which are so similar they cannibalize each other's sales. 
\end{abstract}

\section{Research Question}

BLP is a workhorse for estimating cross-elasticities between products; however, the technique lives in the characteristic space. We need an ontological understanding of these products: what they are, and why each characteristic matters. The econometrician makes a judgement call about whether, e.g., car size is a factor in people's decision to purchase a car. Moreover, these characteristics need to be somehow pre-digested: without a table of car sizes, the econometrician cannot include it in the model even if it is relevant.

Many products contain text descriptions of their characteristics. This could be a Amazon listing for a product, a Yelp summary of a restaurant, or a description of a movie. These descriptions by design include the most important information consumers need about a product, as sellers attempt to attract as many consumers as possible. Thus, if we could incorporate this data into our model, it should be highly informative about the characteristics of the product. Moreover, it removes the judgement call of the economist: the characteristics included are precisely those the merchants themselves believe to be most important.

Thus, the question is: can we use text data to estimate cross-elasticities between products as a supplement (or substitute) for structured characteristic data?

Because text-to-data techniques have evolved so rapidly, the literature is relatively primitive; existing work usually uses algorithms from the pre-LLM era (e.g., word2vec) which are anecdotally much lower quality. I have identified two working papers from the past few months which are particularly relevant.

Compiani et al., released last October as a CESifo working paper, uses Amazon listing text and image data to estimate a combinatorial nested logit for consumer electronics. I believe this would be my closest ``substitute'' in the literature; however, I believe I improve on their results in at least two ways. My first iteration on this paper would be an improvement in the embedding method--they use a bag-of-words, a TF-IDF, a sentence encoder, and Sentence-BERT; notably, all of these embeddings predate the advent of recent LLMs. Secondly, to me the setting for this paper is notably inferior; specifically, I do not believe there is any exogeneous variation in demand. The BLP assumption clearly does not hold here--the characteristics of the Kindle Fire 7 Kids were clearly chosen keeping the characteristics of the Kindle Fire 7 in mind. Thus, the paper seems to be entirely observational.

Michalopoulos and Rauh, published two weeks ago as an NBER working paper, is a complementary paper which also uses movie descriptions as an input for text analysis. However, rather than using this data to estimate distances between different movies, this paper compares each movie to local myths and legends to estimate the impact of cultural proximity on movie success. In many ways this can act as a template for some of my text analysis work in the movie context; however, the thrust of my paper is almost entirely unrelated. Again, this uses a 2018 sentence encoder and the Latent Dirichlet Analysis method first published in 2003--both well predating modern text analysis tools.

A personal note: while I think this is in some ways interesting on its own, I'm not an econometrician--for me, this is basically an intermediate good. I have another project thinking about changes in field subject matter over time; e.g., studying how a 1990 computer science major fares in the 2020 market for software engineers. I'm hoping to use text data such as course descriptions and textbook content to estimate the size of these changes by field, potentially examining substitution between hiring different majors. In that sense this is very much a ``yogurt'' paper--if I can show text data works acceptably in this simple context, I'm hoping to apply it credibly in the more complex setting down the line.

\section{Institutional Setting: Movies}

I suggest piloting this methodology in the movie industry. There are two institutional details which make this a good setting (in conjunction with data availability, which I will cover in the next section).

\subsection{The Uniform-Price Puzzle}

As is well-documented (in one of Liran's papers, among others), the movie industry exhibits a peculiar pricing standard: all movie tickets are the same price. Thus--conditional on the movie being produced--supply of the movie is perfectly elastic. This may or may not be profit maximizing, but I leave that for other papers to explore. The point is that when demand shifts, this manifests \emph{entirely} as a change in quantity. Therefore--unlike other settings where price responds endogeneously--we can reason about demand shifts from price data alone.

\subsection{The BLP Assumption: Twin Movies}

A critical assumption of the BLP model is that product characteristics cannot be determined in strategic response to competitors. While I think this is indefensible in the majority of industries, there exists anecdotal evidence it is true in the movie industry.

The concept of ``twin films'' is well-known enough to have its own--quite developed--Wikipedia page. The idea is that, likely as a result of market researchers reporting similar trends to separate studios independently, two studios develop movies which are remarkably similar. Examples include \emph{A Bug's Life} and \emph{Antz} both releasing in 1998, \emph{Knocked Up} and \emph{Juno} both releasing in 2007, and \emph{No Strings Attached} and \emph{Friends with Benefits} both releasing in 2011.

Generally, twin films are considered unfortunate occurances--audiences will generally see one or the other, rather than both, leading to a significant decline in demand. If firms could avoid them, then, they would; their continued existence suggests that movie studios are in some sense inflexible in the characteristics of their products, satisfying the BLP assumption

\section{Data}

The other virtue of the movie industry is the availability of data. While IMDb is the ``preferred'' movie platform, their data is quite expensive. Fortunately, there exists an community-made competitor, TMDb, with a free API. While IMDb is certainly more comprehensive, for my purposes TMDb should be sufficient (and if this project takes off, I can consider spending the cash for the IMDb data). Figure \ref{fig:star_wars} shows the page for the original \emph{Star Wars} to illustrate what is available. All of these qualities can be retrieved via the free API.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{star_wars_tmdb.png}
    \caption{Screenshot of the TMDb page for \emph{Star Wars}}
    \label{fig:star_wars}
\end{figure}

As far as outcome variables are concerned, I have two options as measures of ``quantity.'' The classic approach would be to use some version of box office receipts over time. Alternatively, I also considered using Google Search trends. To me, both of these seem like expressions of interest in the movie; I will likely perform my analysis using both.

\section{Research Design}

The goal is to show that the unstructured text data is a strong approximation for the structured data. Since we have the virtue of both here, in my mind the ``ideal'' result would be three models: one using structured data, one using text data, and one using both. In a perfect world, all three of these models would generate the same cross-elasticities, suggesting that the text data is a good approximation for the structured data. Moving forward, then, researchers could use exclusively text data even in the absence of structured data with some amount of comfort. One somewhat strange aspect of this approach is it fails if we think text is \emph{better} than the structured data, as in that case the cross elasticities would still be different. We could also try a cross-validation approach to see which model is better at predicting cross-elasticities: train it on 90\% of the sample, and attempt to predict the remaining 10\%.

Generally, we will want to predict the impact of a new entrant on sales of existing movies \emph{conditional on the characteristics of each movie}. Generally, we expect sales of a movie to decline gradually over time. The question, then, is can we detect \emph{excess} decline when movies enter the market which are similar to the incumbents. I think this is feasible with a panel of movies alone; however, if I can find rollout data for movies, I could potentially use a staggered rollout approach as well.

\section{Planning}

At the end of last quarter, I performed some exploratory work to ensure this project was feasible. I pulled and embedded descriptions for three sets of movies using the OpenAI ADA embedding API:
\begin{itemize}
    \item Five Pixar films
    \item Five Tarantino films
    \item Three pairs of ``twin films''
\end{itemize}

Given these embeddings, I performed a t-distributed stochastic neighbor embedding (t-SNE) process to cluster the data in two dimensions. The results are visible in figure \ref{fig:tsne}. I was surprised how effective this was even using only terse, uncleaned movie descriptions: the Tarantino films are clearly separate from the Pixar films, the twin movie's nearest neighbors are always their twin, and twin pair 1 (\emph{Dispicable Me} and \emph{Megamind}, two animated children's films) are grouped with the Pixar films. This leads me to believe distance measures constructed using these embeddings really do have ``substance.''

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{tsne.png}
    \caption{t-SNE of movie descriptions}
    \label{fig:tsne}
\end{figure}

The first step on my docket is simply to construct my sample. From earlier discussions with Liran, it seems worthwhile to trim my sample of movies somehow: a couple hundred movies release each year in the US, so isolating the impact of one movie may be difficult. Given the distribution of movie performance is highly skew, it may be acceptable to focus solely on a subset of movies expected to perform well. I think limiting only to, e.g., ``top 50'' performing movies creates endogeneity concerns, as movies definitionally are in the top 50 because they perform well. However, limiting only to a handful of studios, or only to movies with a certain budget or scale of initial release (two factors set ex-ante), may be more defensible while still providing a large enough sample to work with. Per our discussion in class, this is a very clear ``doing'' task.

As far as ``thinking'' tasks go, to me the primary challenge of this paper is thinking about how to translate the text embeddings to some measure of similarity and competition. Some initial desiderata:
\begin{itemize}
    \item A movie in a more crowded market should experience more competition
    \item A movie with more similar competitors should experience more competition
\end{itemize}

How to balance these is non-obvious: how should I think about, for example, a pair of twin films in duopoly versus a distinctive film in a crowded market?

Similarly, I need to think about how to compute distance between movies. OpenAI provides very high-dimensional embeddings, such that traditional Euclidean distance measures are not feasible under the curse of dimensionality. Some initial ideas:
\begin{itemize}
    \item The simplest (and I think weakest) answer is feeding each embedding value into a simple regression. This is a \emph{very} wide dataset, such that each component likely won't have much power. One low-cost, high-return improvement would be to apply some sort of regularization such as LASSO to trim out the less important components, identifying only the components carrying the most information. 
    \item My understanding of the literature in this space is that generally distance in text embeddings is understood via cosine similarity, representing the angle between the two embedding vectors. This works well for \emph{pairwise} distances, but is not particularly interpretable, and loses any sort of ``directionality'' in the difference.
    \item I could try to reduce the dimensionality of the embeddings. Usually in this kind of model, my understanding is a standard approach is to perform some version of PCA on the data, plot the eigenvalues corresponding to the components, and choose the set of components prior to some inflection point to reduce the number of components. I could then use these components as standard ``Xs'' in a BLP model. 
\end{itemize}

A third-year student in the econometrics group, Amar Venugopal, recently presented some preliminary work in metrics lunch on using text embeddings as a form of treatment variable. He solicited applications for this type of technique during his talk, such that I am optimistic he would find the idea compelling. If the above sounds good to you, I've considered reaching out to him to hear his thoughts, or as a potential coauthor if he finds the project a particularly good fit.

\end{document}

### 2024-04-05 tmdb/code/embed_descriptions.py
Migrated existing scraper

@@ -1,0 +1,44 @@
+from openai import OpenAI
+import csv
+import pandas as pd
+import pickle
+
+# Fetch API key from secret file
+OPENAI_API_KEY = open('openai.secret', 'r').read()
+
+def get_embedding(description):
+
+    client = OpenAI(
+        api_key=OPENAI_API_KEY
+    )
+
+    response = client.embeddings.create(
+        input=[description],
+        model="text-embedding-ada-002"
+    )
+
+    return response
+
+# Read Tarantino movie descriptions into dataframe
+tarantino = pd.read_csv('descriptions/tarantino.csv', encoding="ISO-8859-1")
+tarantino['embedding'] = tarantino['description'].apply(get_embedding)
+
+# Save to disk to avoid re-embedding
+with open('embeddings/tarantino.pkl', 'wb') as f:
+    pickle.dump(tarantino, f)
+
+# Read Pixar movie descriptions into dataframe
+pixar = pd.read_csv('descriptions/pixar.csv', encoding="ISO-8859-1")
+pixar['embedding'] = pixar['description'].apply(get_embedding)
+
+# Save to disk to avoid re-embedding
+with open('embeddings/pixar.pkl', 'wb') as f:
+    pickle.dump(pixar, f)
+
+# Read twin movie descriptions into dataframe
+twins = pd.read_csv('descriptions/twins.csv', encoding="ISO-8859-1")
+twins['embedding'] = twins['description'].apply(get_embedding)
+
+# Save to disk to avoid re-embedding
+with open('embeddings/twins.pkl', 'wb') as f:
+    pickle.dump(twins, f)

### 2024-04-05 tmdb/code/plots.py

@@ -1,0 +1,43 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import pickle
+from sklearn.manifold import TSNE
+import numpy as np
+
+# Load Tarantino embeddings
+with open('embeddings/tarantino.pkl', 'rb') as f:
+    tarantino = pickle.load(f)
+
+# Load Pixar embeddings 
+with open('embeddings/pixar.pkl', 'rb') as f:
+    pixar = pickle.load(f)
+
+# Load twin embeddings
+with open('embeddings/twins.pkl', 'rb') as f:
+    twins = pickle.load(f)
+
+# Concatenate all embeddings
+embeddings = pd.concat([tarantino, pixar, twins])
+
+# Create numpy array from embeddings
+X = embeddings['embedding'].apply(lambda x: x.data[0].embedding).to_list()
+X = np.array(X)
+
+# Create t-SNE model
+tsne = TSNE(n_components=2, random_state=0, perplexity=10)
+reduced_descriptions = tsne.fit_transform(X)
+
+# Add t-SNE results to dataframe
+embeddings['x'] = reduced_descriptions[:, 0]
+embeddings['y'] = reduced_descriptions[:, 1]
+
+# Plot t-SNE results
+plt.figure(figsize=(10, 10))
+for group in embeddings['group'].unique():
+    group_df = embeddings[embeddings['group'] == group]
+    plt.scatter(group_df['x'], group_df['y'], label=group)
+plt.legend()
+plt.title('t-SNE of movie descriptions')
+
+# Save plot to file
+plt.savefig('plots/tsne.png')

### 2024-04-05 tmdb/code/scrape_movies.py

@@ -1,0 +1,59 @@
+import tmdbsimple as tmdb
+import csv
+
+TARANTINO = [466272, 24, 16869, 68718, 273248]
+PIXAR = [862, 585, 12, 9806, 10681]
+TWINS = [
+    50544, # No Strings Attached
+    41630, # Friends with Benefits
+    20352, # Despicable Me
+    38055, # Megamind
+    374720, # Dunkirk
+    399404, # Darkest Hour
+]
+
+# Read API key from file
+with open('tmdb.secret', 'r') as f:
+    api_key = f.read().strip()
+    tmdb.API_KEY = api_key
+
+# Get movie descriptions
+def get_movie_descriptions(movie_ids):
+    descriptions = []
+    for movie_id in movie_ids:
+        movie = tmdb.Movies(movie_id)
+        response = movie.info()
+        descriptions.append(response['overview'])
+    return descriptions
+
+# Get descriptions for Tarantino movies
+tarantino_descriptions = get_movie_descriptions(TARANTINO)
+
+# Write descriptions to file
+with open('descriptions/tarantino.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(['movie_id', 'description', 'group'])
+    for i, description in enumerate(tarantino_descriptions):
+        writer.writerow([TARANTINO[i], description, "tarantino"])
+
+# Get descriptions for Pixar movies
+pixar_descriptions = get_movie_descriptions(PIXAR)
+
+# Write descriptions to file
+with open('descriptions/pixar.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(['movie_id', 'description', 'group'])
+    for i, description in enumerate(pixar_descriptions):
+        writer.writerow([PIXAR[i], description, "pixar"])
+
+# Get descriptions for twin movies
+twin_descriptions = get_movie_descriptions(TWINS)
+
+# Write descriptions to file
+with open('descriptions/twins.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(['movie_id', 'description', 'group'])
+    counter = 0
+    for i, description in enumerate(twin_descriptions):
+        writer.writerow([TWINS[i], description, "twin " + str(counter // 2)])
+        counter += 1

### 2024-04-05 tmdb/code/get_tmdb_info.py
First pass at full TMDB scraper using opusdata

@@ -1,0 +1,167 @@
+import tmdbsimple as tmdb
+from unidecode import unidecode
+import pandas as pd
+import string
+import csv
+
+def clean_name(name):
+    """
+    Function to clean movie name
+    """
+    if not name:
+        return None 
+    
+    # Remove accents and asian characters
+    clean_name = unidecode(clean_name)
+    
+    # Remove punctuation
+    clean_name = name.translate(str.maketrans('', '', string.punctuation))
+
+    # Lowercase
+    clean_name = clean_name.lower()
+
+    # Handcoding some exceptions
+    if clean_name == 'chang jiang qi hao  cj7':
+        clean_name = 'cj7'
+    if clean_name == 'stan helsing a parody':
+        clean_name = 'stan helsing'
+    elif clean_name == 'san suk si gin':
+        clean_name = 'shinjuku incident'
+    elif clean_name == 'shi yue wei cheng':
+        clean_name = 'bodyguards and assassins'
+    elif clean_name == 'les intouchables':
+        clean_name = 'the intouchables'
+    elif clean_name == 'lee daniels the butler':
+        clean_name = 'the butler'
+    elif clean_name == 'mr popperss penguins':
+        clean_name = 'mr poppers penguins'
+    elif clean_name == 'doctor seuss the lorax':
+        clean_name = 'the lorax'
+    elif clean_name == 'jin ling shi san chai':
+        clean_name = 'the flowers of war'
+    elif clean_name == 'san cheng ji':
+        clean_name = 'a tale of three cities'
+    elif clean_name == 'savva serdtse voyna':
+        clean_name = 'savva heart of the warrior'
+    elif clean_name == 'star wars ep vii the force awakens':
+        clean_name = 'star wars the force awakens'
+    elif clean_name == 'chai dan zhuanjia':
+        clean_name = 'shock wave'
+    elif clean_name == 'star wars ep viii the last jedi':
+        clean_name = 'star wars the last jedi'
+    elif clean_name == 'spiderman into the spiderverse 3d':
+        clean_name = 'spiderman into the spiderverse'
+    elif clean_name == 'walle':
+        clean_name = 'WALL·E'
+    elif clean_name == 'halloween 2':
+        clean_name = 'halloween ii'
+    elif clean_name == 'michael jacksons this is it':
+        clean_name = 'this is it'
+    elif clean_name == 'disneys a christmas carol':
+        clean_name = 'a christmas carol'
+
+    return clean_name
+
[…]

### 2024-04-05 tmdb/code/scrape_movies.py

@@ -1,5 +1,31 @@
 import tmdbsimple as tmdb
+import pandas as pd
 import csv
+
+# Read API key from file
+with open('tmdb.secret', 'r') as f:
+    api_key = f.read().strip()
+    tmdb.API_KEY = api_key
+
+# Get movie list from OpusData sample
+opusdata = pd.read_csv('../OpusData/MovieData.csv')
+
+print(opusdata.head())
+# Get name and year of first movie
+movie_name = opusdata['movie_name'][0]
+movie_year = opusdata['production_year'][0]
+
+search = tmdb.Search()
+response = search.movie(query=movie_name, year=movie_year)
+
+# Ensure only one result is found
+assert search.total_results > 0, "No results found for movie"
+assert search.total_results == 1, "Multiple results found for movie"
+
+movie_info = search.results[0]
+movie_id = movie_info['id']
+print(movie_info)
+kill
 
 TARANTINO = [466272, 24, 16869, 68718, 273248]
 PIXAR = [862, 585, 12, 9806, 10681]
@@ -12,10 +38,7 @@ TWINS = [
     399404, # Darkest Hour
 ]
 
-# Read API key from file
-with open('tmdb.secret', 'r') as f:
-    api_key = f.read().strip()
-    tmdb.API_KEY = api_key
+
 
 # Get movie descriptions
 def get_movie_descriptions(movie_ids):

### 2024-04-05 tmdb/code/get_tmdb_info.py
Handcoding additions

@@ -4,6 +4,191 @@ import pandas as pd
 import string
 import csv
 
+YEAR_HANDCODES = movie_years = {
+    'romance and cigarettes': 2005,
+    'slow burn': 2005,
+    'red tails': 2012,
+    'august osage county': 2013,
+    'sin city a dame to kill for': 2014,
+    'edge of tomorrow': 2014,
+    'the last song': 2010,
+    'faster': 2010,
+    'valentines day': 2010,
+    'killers': 2010,
+    'the roommate': 2011,
+    'the beaver': 2011,
+    'the help': 2011,
+    'the eagle': 2011,
+    'no strings attached': 2011,
+    'hanna': 2011,
+    'winnie the pooh': 2011,
+    'sanctum': 2011,
+    'the tree of life': 2011,
+    'abduction': 2011,
+    'the rite': 2011,
+    'unknown': 2011,
+    'paul': 2011,
+    'the big year': 2011,
+    'red riding hood': 2011,
+    'the mechanic': 2011,
+    'super 8': 2011,
+    'upside down': 2012,
+    'priest': 2011,
+    'hop': 2011,
+    'battle los angeles': 2011,
+    'the dilemma': 2011,
+    'movie 43': 2013,
+    'lol': 2012,
+    'quartet': 2012,
+    'project x': 2012,
+    'chronicle': 2012,
+    'stoker': 2013,
+    'the possession': 2012,
+    'promised land': 2012,
+    'vamps': 2012,
+    'the apparition': 2012,
+    'red lights': 2012,
+    'lockout': 2012,
+    'ride along': 2014,
+    'the raven': 2012,
+    'lawless': 2012,
+    'the vow': 2012,
+    'the three stooges': 2012,
+    'the butler': 2013,
+    'safe': 2012,
+    'this is 40': 2012,
+    'parker': 2013, 
+    'riddick': 2013,
+    'the impossible': 2012,
+    'big miracle': 2012,
+    'taken 2': 2012,
+    'ted': 2012,
+    'the equalizer': 2014,
+    'the watch': 2012,
+    'lincoln': 2012,
+    'the dictator': 2012,
+    'the hunger games': 2012,
+    'mirror mirror': 2012,
+    'the campaign': 2012,
+    'life of pi': 2012,
+    'prometheus': 2012,
+    'world war z': 2013,
+    'battleship': 2012,
+    'the avengers': 2012,
+    'john carter': 2012,
[…]

### 2024-04-05 tmdb/code/get_tmdb_info.py
Cleanup of scraper script

@@ -160,6 +160,7 @@ YEAR_HANDCODES = movie_years = {
     'safe house': 2012,
     'tomorrowland': 2015,
     'peter rabbit': 2018,
+    'warrior': 2011,
 }
 
 
@@ -230,15 +231,12 @@ def get_movie_info(row):
     search.movie(query=cleaned_name, year=year)
 
     if not search.total_results:
-        #print(f"Movie {row['movie_name']} not found in year {row['production_year']}, trying previous year")
         search.movie(query=cleaned_name, year= year - 1)
 
     if not search.total_results:
-        #print(f"Movie {row['movie_name']} not found in year {row['production_year']}, trying next year")
         search.movie(query=cleaned_name, year = year + 1)
 
     if not search.total_results:
-        #print(f"Movie {row['movie_name']} not found in year {row['production_year']}, trying two years out")
         search.movie(query=cleaned_name, year = year + 2)
 
     # If no results
@@ -247,7 +245,6 @@ def get_movie_info(row):
         return None
 
     return search.results[0]
-
 
 # Read API key from file
 with open('tmdb.secret', 'r') as f:
@@ -260,6 +257,7 @@ opusdata = pd.read_csv('../OpusData/MovieData.csv')
 # Get movie info for each movie
 query_results = opusdata.apply(get_movie_info, axis=1)
 
+# Parse TMDB info into dataframe
 opusdata['tmdb_id'] = [query_result['id'] if query_result else None for query_result in query_results]
 opusdata['tmdb_name'] = [query_result['title'] if query_result else None for query_result in query_results]
 opusdata['tmdb_description'] = [query_result['overview'] if query_result else None for query_result in query_results]

### 2024-04-05 tmdb/code/embed_descriptions.py
Consolidate preliminary scripts into single proof-of-concept

@@ -1,44 +1,0 @@
-from openai import OpenAI
-import csv
-import pandas as pd
-import pickle
-
-# Fetch API key from secret file
-OPENAI_API_KEY = open('openai.secret', 'r').read()
-
-def get_embedding(description):
-
-    client = OpenAI(
-        api_key=OPENAI_API_KEY
-    )
-
-    response = client.embeddings.create(
-        input=[description],
-        model="text-embedding-ada-002"
-    )
-
-    return response
-
-# Read Tarantino movie descriptions into dataframe
-tarantino = pd.read_csv('descriptions/tarantino.csv', encoding="ISO-8859-1")
-tarantino['embedding'] = tarantino['description'].apply(get_embedding)
-
-# Save to disk to avoid re-embedding
-with open('embeddings/tarantino.pkl', 'wb') as f:
-    pickle.dump(tarantino, f)
-
-# Read Pixar movie descriptions into dataframe
-pixar = pd.read_csv('descriptions/pixar.csv', encoding="ISO-8859-1")
-pixar['embedding'] = pixar['description'].apply(get_embedding)
-
-# Save to disk to avoid re-embedding
-with open('embeddings/pixar.pkl', 'wb') as f:
-    pickle.dump(pixar, f)
-
-# Read twin movie descriptions into dataframe
-twins = pd.read_csv('descriptions/twins.csv', encoding="ISO-8859-1")
-twins['embedding'] = twins['description'].apply(get_embedding)
-
-# Save to disk to avoid re-embedding
-with open('embeddings/twins.pkl', 'wb') as f:
-    pickle.dump(twins, f)

### 2024-04-05 tmdb/code/plots.py

@@ -1,43 +1,0 @@
-import matplotlib.pyplot as plt
-import pandas as pd
-import pickle
-from sklearn.manifold import TSNE
-import numpy as np
-
-# Load Tarantino embeddings
-with open('embeddings/tarantino.pkl', 'rb') as f:
-    tarantino = pickle.load(f)
-
-# Load Pixar embeddings 
-with open('embeddings/pixar.pkl', 'rb') as f:
-    pixar = pickle.load(f)
-
-# Load twin embeddings
-with open('embeddings/twins.pkl', 'rb') as f:
-    twins = pickle.load(f)
-
-# Concatenate all embeddings
-embeddings = pd.concat([tarantino, pixar, twins])
-
-# Create numpy array from embeddings
-X = embeddings['embedding'].apply(lambda x: x.data[0].embedding).to_list()
-X = np.array(X)
-
-# Create t-SNE model
-tsne = TSNE(n_components=2, random_state=0, perplexity=10)
-reduced_descriptions = tsne.fit_transform(X)
-
-# Add t-SNE results to dataframe
-embeddings['x'] = reduced_descriptions[:, 0]
-embeddings['y'] = reduced_descriptions[:, 1]
-
-# Plot t-SNE results
-plt.figure(figsize=(10, 10))
-for group in embeddings['group'].unique():
-    group_df = embeddings[embeddings['group'] == group]
-    plt.scatter(group_df['x'], group_df['y'], label=group)
-plt.legend()
-plt.title('t-SNE of movie descriptions')
-
-# Save plot to file
-plt.savefig('plots/tsne.png')

### 2024-04-05 tmdb/code/proof_of_concept.py

@@ -1,0 +1,175 @@
+import tmdbsimple as tmdb
+import pandas as pd
+import csv
+import pickle
+from sklearn.manifold import TSNE
+import numpy as np
+import matplotlib.pyplot as plt
+from openai import OpenAI
+
+
+########## Pull data
+
+# Read API key from file
+with open('tmdb.secret', 'r') as f:
+    api_key = f.read().strip()
+    tmdb.API_KEY = api_key
+
+# Get movie list from OpusData sample
+opusdata = pd.read_csv('../OpusData/MovieData.csv')
+
+print(opusdata.head())
+# Get name and year of first movie
+movie_name = opusdata['movie_name'][0]
+movie_year = opusdata['production_year'][0]
+
+search = tmdb.Search()
+response = search.movie(query=movie_name, year=movie_year)
+
+# Ensure only one result is found
+assert search.total_results > 0, "No results found for movie"
+assert search.total_results == 1, "Multiple results found for movie"
+
+movie_info = search.results[0]
+movie_id = movie_info['id']
+print(movie_info)
+
+TARANTINO = [466272, 24, 16869, 68718, 273248]
+PIXAR = [862, 585, 12, 9806, 10681]
+TWINS = [
+    50544, # No Strings Attached
+    41630, # Friends with Benefits
+    20352, # Despicable Me
+    38055, # Megamind
+    374720, # Dunkirk
+    399404, # Darkest Hour
+]
+
+
+
+# Get movie descriptions
+def get_movie_descriptions(movie_ids):
+    descriptions = []
+    for movie_id in movie_ids:
+        movie = tmdb.Movies(movie_id)
+        response = movie.info()
+        descriptions.append(response['overview'])
+    return descriptions
+
+# Get descriptions for Tarantino movies
+tarantino_descriptions = get_movie_descriptions(TARANTINO)
+
+# Write descriptions to file
+with open('descriptions/tarantino.csv', 'w') as f:
+    writer = csv.writer(f)
+    writer.writerow(['movie_id', 'description', 'group'])
+    for i, description in enumerate(tarantino_descriptions):
+        writer.writerow([TARANTINO[i], description, "tarantino"])
+
+# Get descriptions for Pixar movies
+pixar_descriptions = get_movie_descriptions(PIXAR)
+
+# Write descriptions to file
+with open('descriptions/pixar.csv', 'w') as f:
[…]

### 2024-04-05 tmdb/code/scrape_movies.py

@@ -1,82 +1,0 @@
-import tmdbsimple as tmdb
-import pandas as pd
-import csv
-
-# Read API key from file
-with open('tmdb.secret', 'r') as f:
-    api_key = f.read().strip()
-    tmdb.API_KEY = api_key
-
-# Get movie list from OpusData sample
-opusdata = pd.read_csv('../OpusData/MovieData.csv')
-
-print(opusdata.head())
-# Get name and year of first movie
-movie_name = opusdata['movie_name'][0]
-movie_year = opusdata['production_year'][0]
-
-search = tmdb.Search()
-response = search.movie(query=movie_name, year=movie_year)
-
-# Ensure only one result is found
-assert search.total_results > 0, "No results found for movie"
-assert search.total_results == 1, "Multiple results found for movie"
-
-movie_info = search.results[0]
-movie_id = movie_info['id']
-print(movie_info)
-kill
-
-TARANTINO = [466272, 24, 16869, 68718, 273248]
-PIXAR = [862, 585, 12, 9806, 10681]
-TWINS = [
-    50544, # No Strings Attached
-    41630, # Friends with Benefits
-    20352, # Despicable Me
-    38055, # Megamind
-    374720, # Dunkirk
-    399404, # Darkest Hour
-]
-
-
-
-# Get movie descriptions
-def get_movie_descriptions(movie_ids):
-    descriptions = []
-    for movie_id in movie_ids:
-        movie = tmdb.Movies(movie_id)
-        response = movie.info()
-        descriptions.append(response['overview'])
-    return descriptions
-
-# Get descriptions for Tarantino movies
-tarantino_descriptions = get_movie_descriptions(TARANTINO)
-
-# Write descriptions to file
-with open('descriptions/tarantino.csv', 'w') as f:
-    writer = csv.writer(f)
-    writer.writerow(['movie_id', 'description', 'group'])
-    for i, description in enumerate(tarantino_descriptions):
-        writer.writerow([TARANTINO[i], description, "tarantino"])
-
-# Get descriptions for Pixar movies
-pixar_descriptions = get_movie_descriptions(PIXAR)
-
-# Write descriptions to file
-with open('descriptions/pixar.csv', 'w') as f:
-    writer = csv.writer(f)
-    writer.writerow(['movie_id', 'description', 'group'])
-    for i, description in enumerate(pixar_descriptions):
-        writer.writerow([PIXAR[i], description, "pixar"])
-
-# Get descriptions for twin movies
[…]

### 2024-04-06 tmdb/code/embed_descriptions.py
Handle missing descriptions

@@ -1,0 +1,33 @@
+from openai import OpenAI
+import pandas as pd
+import pickle
+
+# Input OpenAI API key
+OPENAI_API_KEY = open('openai.secret', 'r').read()
+
+def get_embedding(description):
+    """
+    Function taking in a description and returnning an embedding object
+    """
+    client = OpenAI(
+        api_key=OPENAI_API_KEY
+    )
+
+    if not description:
+        print("No description!")
+        return None
+
+    response = client.embeddings.create(
+        input=[description],
+        model="text-embedding-3-small"
+    )
+
+    return response
+
+# Get descriptions
+movie_data = pd.read_csv('descriptions/movie_descriptions.csv', encoding="ISO-8859-1")
+
+movie_data['embedding'] = movie_data['tmdb_description'].apply(get_embedding)
+
+with open('embeddings/movie_descriptions.pkl', 'wb') as f:
+    pickle.dump(movie_data, f)

### 2024-04-06 tmdb/code/get_tmdb_info.py

@@ -161,6 +161,7 @@ YEAR_HANDCODES = movie_years = {
     'tomorrowland': 2015,
     'peter rabbit': 2018,
     'warrior': 2011,
+    'mama': 2013,
 }
 
 

### 2024-04-28 google_trends/code/pull_trends.py
Misc changes

@@ -1,0 +1,59 @@
+from pytrends.request import TrendReq
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Create pytrends object
+pytrends = TrendReq(hl='en-US', tz=360)
+
+# Load data from OpusData
+opusdata = pd.read_csv('../OpusData/MovieData.csv')
+
+# Get list of movie names
+movie_names = opusdata['movie_name'].to_list()
+movie_years = opusdata['production_year'].to_list()
+movie_ids = opusdata['movie_odid'].to_list()
+
+# Create dataframe to store movie interests
+# Has three columns: movie id, date, and interest
+movie_interests_long = pd.DataFrame(columns=['movie_id', 'date', 'interest'])
+
+for i in range(len(movie_names)):
+    # Unpack movie data
+    movie_name = movie_names[i]
+    movie_year = movie_years[i]
+    movie_id = movie_ids[i]
+
+    # Get Google Trends data
+    pytrends.build_payload(
+        kw_list=[movie_name + " movie"],
+        timeframe=f'{movie_year - 1}-01-01 {movie_year + 1}-12-31',
+    )
+
+    # Get interest over time
+    interest_over_time = pytrends.interest_over_time()
+
+    # Drop isPartial column
+    try:
+        interest_over_time = interest_over_time.drop(columns='isPartial')
+    except KeyError:
+        print(f"Error with {movie_name} {movie_year}")
+        print(interest_over_time)
+
+    # Drop if value is zero
+    interest_over_time = interest_over_time[interest_over_time[movie_name + ' movie'] != 0]
+
+    # Add movie id to column
+    interest_over_time['movie_id'] = movie_id
+
+    # Reset index
+    interest_over_time = interest_over_time.reset_index()
+
+    # Rename columns
+    interest_over_time.columns = ['date', 'interest', 'movie_id']
+
+    # Add results to end of long dataframe
+    movie_interests_long = pd.concat([movie_interests_long, interest_over_time])
+
+# Save movie_interests_long
+movie_interests_long.to_csv('output/movie_interests_long.csv', index=False)
+

### 2024-04-28 tmdb/code/embed_descriptions.py

@@ -1,6 +1,9 @@
 from openai import OpenAI
 import pandas as pd
 import pickle
+from tqdm import tqdm
+
+tqdm.pandas()
 
 # Input OpenAI API key
 OPENAI_API_KEY = open('openai.secret', 'r').read()
@@ -27,7 +30,7 @@ def get_embedding(description):
 # Get descriptions
 movie_data = pd.read_csv('descriptions/movie_descriptions.csv', encoding="ISO-8859-1")
 
-movie_data['embedding'] = movie_data['tmdb_description'].apply(get_embedding)
+movie_data['embedding'] = movie_data['tmdb_description'].progress_apply(get_embedding)
 
 with open('embeddings/movie_descriptions.pkl', 'wb') as f:
     pickle.dump(movie_data, f)

### 2024-04-28 tmdb/code/plot_genres.py

@@ -1,0 +1,32 @@
+import pickle
+import pandas as pd
+from sklearn.manifold import TSNE
+import numpy as np
+import matplotlib.pyplot as plt
+
+# Load in the embeddings
+with open('embeddings/movie_descriptions.pkl', 'rb') as f:
+    movie_data = pickle.load(f)
+
+X = movie_data['embedding'].apply(lambda x: x.data[0].embedding).to_list()
+X = np.array(X)
+
+# Create t-SNE model
+tsne = TSNE(n_components=2, random_state=0, perplexity=10)
+reduced_descriptions = tsne.fit_transform(X)
+
+# Add t-SNE results to dataframe
+movie_data['x'] = reduced_descriptions[:, 0]
+movie_data['y'] = reduced_descriptions[:, 1]
+
+# Plot t-SNE results
+plt.figure(figsize=(10, 10))
+
+# Plot each genre in a different color
+for genre in movie_data['genre'].unique():
+    genre_data = movie_data[movie_data['genre'] == genre]
+    plt.scatter(genre_data['x'], genre_data['y'], label=genre)
+plt.legend()
+
+# Save the plot
+plt.savefig('plots/tsne_genres.png')

### 2024-04-29 draft/draft.tex
Initial cross-elasticity estimate

--- 2024-04-03 Intro Memo/intro_memo.tex
@@
 
-\usepackage{amsmath}
 \usepackage{geometry}
-\usepackage{graphicx}
-
-\author{Gideon Moore}
-\date{April 3, 2024}
-\title{Introductory Memo: \\ Demand Estimation Using Movie Descriptions}
 
@@
 
-\maketitle
-
-\begin{abstract}
-    Traditional BLP-based demand estimation requires structured characteristic data to identify cross-elasticities.
-Many markets, however, feature unstructured text descriptions of products without well-structured characteristics.
-Using modern text-to-data tools, I hope to estimate cross-elasticities in the presence of unstructured text alone.
-I exploit specific characteristics of the movie industry to pilot this methodology.
-Uniform pricing is ubiquitous in this market; thus, we can make arguments about demand shifts from quantity changes alone.
-Moreover, long and inflexible production pipelines have the potential to produce plausibly-exogeneous product characteristics; this argument is supported by the presence of ``twin films,'' movies released simultaneously which are so similar they cannibalize each other's sales.
-\end{abstract}
-
-\section{Research Question}
-
-BLP is a workhorse for estimating cross-elasticities between products; however, the technique lives in the characteristic space.
-We need an ontological understanding of these products: what they are, and why each characteristic matters.
-The econometrician makes a judgement call about whether, e.g., car size is a factor in people's decision to purchase a car.
-Moreover, these characteristics need to be somehow pre-digested: without a table of car sizes, the econometrician cannot include it in the model even if it is relevant.
-
-Many products contain text descriptions of their characteristics.
-This could be a Amazon listing for a product, a Yelp summary of a restaurant, or a description of a movie.
-These descriptions by design include the most important information consumers need about a product, as sellers attempt to attract as many consumers as possible.
-Thus, if we could incorporate this data into our model, it should be highly informative about the characteristics of the product.
-Moreover, it removes the judgement call of the economist: the characteristics included are precisely those the merchants themselves believe to be most important.
-
-Thus, the question is: can we use text data to estimate cross-elasticities between products as a supplement (or substitute) for structured characteristic data?
-
-Because text-to-data techniques have evolved so rapidly, the literature is relatively primitive; existing work usually uses algorithms from the pre-LLM era (e.g., word2vec) which are anecdotally much lower quality.
-I have identified two working papers from the past few months which are particularly relevant.
-
-Compiani et al., released last October as a CESifo working paper, uses Amazon listing text and image data to estimate a combinatorial nested logit for consumer electronics.
-I believe this would be my closest ``substitute'' in the literature; however, I believe I improve on their results in at least two ways.
-My first iteration on this paper would be an improvement in the embedding method--they use a bag-of-words, a TF-IDF, a sentence encoder, and Sentence-BERT; notably, all of these embeddings predate the advent of recent LLMs.
-Secondly, to me the setting for this paper is notably inferior; specifically, I do not believe there is any exogeneous variation in demand.
-The BLP assumption clearly does not hold here--the characteristics of the Kindle Fire 7 Kids were clearly chosen keeping the characteristics of the Kindle Fire 7 in mind.
-Thus, the paper seems to be entirely observational.
-
-Michalopoulos and Rauh, published two weeks ago as an NBER working paper, is a complementary paper which also uses movie descriptions as an input for text analysis.
-However, rather than using this data to estimate distances between different movies, this paper compares each movie to local myths and legends to estimate the impact of cultural proximity on movie success.
-In many ways this can act as a template for some of my text analysis work in the movie context; however, the thrust of my paper is almost entirely unrelated.
-Again, this uses a 2018 sentence encoder and the Latent Dirichlet Analysis method first published in 2003--both well predating modern text analysis tools.
-
-A personal note: while I think this is in some ways interesting on its own, I'm not an econometrician--for me, this is basically an intermediate good.
-I have another project thinking about changes in field subject matter over time; e.g., studying how a 1990 computer science major fares in the 2020 market for software engineers.
-I'm hoping to use text data such as course descriptions and textbook content to estimate the size of these changes by field, potentially examining substitution between hiring different majors.
-In that sense this is very much a ``yogurt'' paper--if I can show text data works acceptably in this simple context, I'm hoping to apply it credibly in the more complex setting down the line.
-
-\section{Institutional Setting: Movies}
-
-I suggest piloting this methodology in the movie industry.
-There are two institutional details which make this a good setting (in conjunction with data availability, which I will cover in the next section).
-
-\subsection{The Uniform-Price Puzzle}
-
-As is well-documented (in one of Liran's papers, among others), the movie industry exhibits a peculiar pricing standard: all movie tickets are the same price.
-Thus--conditional on the movie being produced--supply of the movie is perfectly elastic.
-This may or may not be profit maximizing, but I leave that for other papers to explore.
-The point is that when demand shifts, this manifests \emph{entirely} as a change in quantity.
-Therefore--unlike other settings where price responds endogeneously--we can reason about demand shifts from price data alone.
-
-\subsection{The BLP Assumption: Twin Movies}
-
-A critical assumption of the BLP model is that product characteristics cannot be determined in strategic response to competitors.
-While I think this is indefensible in the majority of industries, there exists anecdotal evidence it is true in the movie industry.
-
-The concept of ``twin films'' is well-known enough to have its own--quite developed--Wikipedia page.
-The idea is that, likely as a result of market researchers reporting similar trends to separate studios independently, two studios develop movies which are remarkably similar.
-Examples include \emph{A Bug's Life} and \emph{Antz} both releasing in 1998, \emph{Knocked Up} and \emph{Juno} both releasing in 2007, and \emph{No Strings Attached} and \emph{Friends with Benefits} both releasing in 2011.
-
-Generally, twin films are considered unfortunate occurances--audiences will generally see one or the other, rather than both, leading to a significant decline in demand.
-If firms could avoid them, then, they would; their continued existence suggests that movie studios are in some sense inflexible in the characteristics of their products, satisfying the BLP assumption
-
-\section{Data}
-
-The other virtue of the movie industry is the availability of data.
-While IMDb is the ``preferred'' movie platform, their data is quite expensive.
-Fortunately, there exists an community-made competitor, TMDb, with a free API.
-While IMDb is certainly more comprehensive, for my purposes TMDb should be sufficient (and if this project takes off, I can consider spending the cash for the IMDb data).
-Figure \ref{fig:star_wars} shows the page for the original \emph{Star Wars} to illustrate what is available.
-All of these qualities can be retrieved via the free API.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{star_wars_tmdb.png}
-    \caption{Screenshot of the TMDb page for \emph{Star Wars}}
-    \label{fig:star_wars}
-\end{figure}
-
-As far as outcome variables are concerned, I have two options as measures of ``quantity.'' The classic approach would be to use some version of box office receipts over time.
-Alternatively, I also considered using Google Search trends.
-To me, both of these seem like expressions of interest in the movie; I will likely perform my analysis using both.
-
-\section{Research Design}
-
-The goal is to show that the unstructured text data is a strong approximation for the structured data.
-Since we have the virtue of both here, in my mind the ``ideal'' result would be three models: one using structured data, one using text data, and one using both.
-In a perfect world, all three of these models would generate the same cross-elasticities, suggesting that the text data is a good approximation for the structured data.
-Moving forward, then, researchers could use exclusively text data even in the absence of structured data with some amount of comfort.
-One somewhat strange aspect of this approach is it fails if we think text is \emph{better} than the structured data, as in that case the cross elasticities would still be different.
-We could also try a cross-validation approach to see which model is better at predicting cross-elasticities: train it on 90\% of the sample, and attempt to predict the remaining 10\%.
-
-Generally, we will want to predict the impact of a new entrant on sales of existing movies \emph{conditional on the characteristics of each movie}.
-Generally, we expect sales of a movie to decline gradually over time.
-The question, then, is can we detect \emph{excess} decline when movies enter the market which are similar to the incumbents.
-I think this is feasible with a panel of movies alone; however, if I can find rollout data for movies, I could potentially use a staggered rollout approach as well.
-
-\section{Planning}
-
-At the end of last quarter, I performed some exploratory work to ensure this project was feasible. I pulled and embedded descriptions for three sets of movies using the OpenAI ADA embedding API:
-\begin{itemize}
-    \item Five Pixar films
-    \item Five Tarantino films
-    \item Three pairs of ``twin films''
-\end{itemize}
-
-Given these embeddings, I performed a t-distributed stochastic neighbor embedding (t-SNE) process to cluster the data in two dimensions.
-The results are visible in figure \ref{fig:tsne}.
-I was surprised how effective this was even using only terse, uncleaned movie descriptions: the Tarantino films are clearly separate from the Pixar films, the twin movie's nearest neighbors are always their twin, and twin pair 1 (\emph{Dispicable Me} and \emph{Megamind}, two animated children's films) are grouped with the Pixar films.
-This leads me to believe distance measures constructed using these embeddings really do have ``substance.''
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{tsne.png}
-    \caption{t-SNE of movie descriptions}
-    \label{fig:tsne}
-\end{figure}
-
-The first step on my docket is simply to construct my sample.
-From earlier discussions with Liran, it seems worthwhile to trim my sample of movies somehow: a couple hundred movies release each year in the US, so isolating the impact of one movie may be difficult.
-Given the distribution of movie performance is highly skew, it may be acceptable to focus solely on a subset of movies expected to perform well.
-I think limiting only to, e.g., ``top 50'' performing movies creates endogeneity concerns, as movies definitionally are in the top 50 because they perform well.
-However, limiting only to a handful of studios, or only to movies with a certain budget or scale of initial release (two factors set ex-ante), may be more defensible while still providing a large enough sample to work with.
-Per our discussion in class, this is a very clear ``doing'' task.
-
-As far as ``thinking'' tasks go, to me the primary challenge of this paper is thinking about how to translate the text embeddings to some measure of similarity and competition.
-Some initial desiderata:
-\begin{itemize}
-    \item A movie in a more crowded market should experience more competition
-    \item A movie with more similar competitors should experience more competition
-\end{itemize}
-
-How to balance these is non-obvious: how should I think about, for example, a pair of twin films in duopoly versus a distinctive film in a crowded market?
-
-Similarly, I need to think about how to compute distance between movies.
-OpenAI provides very high-dimensional embeddings, such that traditional Euclidean distance measures are not feasible under the curse of dimensionality.
-Some initial ideas:
-\begin{itemize}
-    \item The simplest (and I think weakest) answer is feeding each embedding value into a simple regression.
-This is a \emph{very} wide dataset, such that each component likely won't have much power.
-One low-cost, high-return improvement would be to apply some sort of regularization such as LASSO to trim out the less important components, identifying only the components carrying the most information.
-    \item My understanding of the literature in this space is that generally distance in text embeddings is understood via cosine similarity, representing the angle between the two embedding vectors.
-This works well for \emph{pairwise} distances, but is not particularly interpretable, and loses any sort of ``directionality'' in the difference.
-    \item I could try to reduce the dimensionality of the embeddings.
-Usually in this kind of model, my understanding is a standard approach is to perform some version of PCA on the data, plot the eigenvalues corresponding to the components, and choose the set of components prior to some inflection point to reduce the number of components.
-I could then use these components as standard ``Xs'' in a BLP model.
-\end{itemize}
-
-A third-year student in the econometrics group, Amar Venugopal, recently presented some preliminary work in metrics lunch on using text embeddings as a form of treatment variable.
-He solicited applications for this type of technique during his talk, such that I am optimistic he would find the idea compelling.
-If the above sounds good to you, I've considered reaching out to him to hear his thoughts, or as a potential coauthor if he finds the project a particularly good fit.
-
 \end{document}

### 2024-04-29 OpusData/code/clean_opusdata.py

@@ -1,2 +1,229 @@
 import pandas as pd
+from unidecode import unidecode
+import string
 
+YEAR_HANDCODES = {
+    'romance and cigarettes': 2005,
+    'slow burn': 2005,
+    'red tails': 2012,
+    'august osage county': 2013,
+    'sin city a dame to kill for': 2014,
+    'edge of tomorrow': 2014,
+    'the last song': 2010,
+    'faster': 2010,
+    'valentines day': 2010,
+    'killers': 2010,
+    'the roommate': 2011,
+    'the beaver': 2011,
+    'the help': 2011,
+    'the eagle': 2011,
+    'no strings attached': 2011,
+    'hanna': 2011,
+    'winnie the pooh': 2011,
+    'sanctum': 2011,
+    'the tree of life': 2011,
+    'abduction': 2011,
+    'the rite': 2011,
+    'unknown': 2011,
+    'paul': 2011,
+    'the big year': 2011,
+    'red riding hood': 2011,
+    'the mechanic': 2011,
+    'super 8': 2011,
+    'upside down': 2012,
+    'priest': 2011,
+    'hop': 2011,
+    'battle los angeles': 2011,
+    'the dilemma': 2011,
+    'movie 43': 2013,
+    'lol': 2012,
+    'quartet': 2012,
+    'project x': 2012,
+    'chronicle': 2012,
+    'stoker': 2013,
+    'the possession': 2012,
+    'promised land': 2012,
+    'vamps': 2012,
+    'the apparition': 2012,
+    'red lights': 2012,
+    'lockout': 2012,
+    'ride along': 2014,
+    'the raven': 2012,
+    'lawless': 2012,
+    'the vow': 2012,
+    'the three stooges': 2012,
+    'the butler': 2013,
+    'safe': 2012,
+    'this is 40': 2012,
+    'parker': 2013, 
+    'riddick': 2013,
+    'the impossible': 2012,
+    'big miracle': 2012,
+    'taken 2': 2012,
+    'ted': 2012,
+    'the equalizer': 2014,
+    'the watch': 2012,
+    'lincoln': 2012,
+    'the dictator': 2012,
+    'the hunger games': 2012,
+    'mirror mirror': 2012,
+    'the campaign': 2012,
+    'life of pi': 2012,
+    'prometheus': 2012,
+    'world war z': 2013,
+    'battleship': 2012,
+    'the avengers': 2012,
+    'john carter': 2012,
+    'admission': 2013,
[…]

### 2024-04-29 elasticity_estimate/code/estimate_elasticity.py

@@ -1,0 +1,202 @@
+import pandas as pd
+import numpy as np
+import pickle
+from sklearn.decomposition import PCA
+from scipy.optimize import minimize
+import matplotlib.pyplot as plt
+
+# Load the embeddings
+embeddings = pickle.load(open('../tmdb/embeddings/movie_descriptions.pkl', 'rb'))
+
+# Rename movie_odid to movie_id
+embeddings.rename(columns={'movie_odid': 'movie_id'}, inplace=True)
+
+# Merge embedding data onto google tren
+
+raw_embeddings = embeddings['embedding'].apply(lambda x: x.data[0].embedding).to_list()
+
+raw_embeddings = np.array(raw_embeddings)
+reduced_embeddings = PCA(n_components=6).fit_transform(raw_embeddings)
+
+for i in range(6):
+    embeddings[f'pca_{i}'] = reduced_embeddings[:, i]
+
+# Load google trends
+google_trends = pd.read_csv('../google_trends/output/trends.csv')
+
+# Filter only to dates with 100 interest
+premier_dates = google_trends[google_trends['interest'] == 100]
+
+# In the event of a tie, take the first date
+premier_dates = premier_dates.groupby('movie_id').first()
+premier_dates.reset_index(inplace=True)
+
+# Rename columns
+premier_dates.columns = ['movie_id', 'premier_date', 'interest']
+premier_dates = premier_dates[['movie_id', 'premier_date']]
+
+# Merge premier_dates onto trends
+google_trends = google_trends.merge(premier_dates, on='movie_id')
+
+# Keep if date is within three months of premier
+google_trends['date'] = pd.to_datetime(google_trends['date'])
+google_trends['premier_date'] = pd.to_datetime(google_trends['premier_date'])
+google_trends['days_from_premier'] = (google_trends['date'] - google_trends['premier_date']).dt.days
+google_trends = google_trends[google_trends['days_from_premier'].between(0, 90)]
+
+# Keep if date has positive interest
+google_trends = google_trends[google_trends['interest'] > 0]
+
+# Limit only to movies from 2007
+google_trends = google_trends[google_trends['premier_date'].dt.year == 2007]
+
+# Compute log interest
+google_trends['ln_interest'] = np.log(google_trends['interest'])
+
+# De-Mean log interest within a month
+google_trends['month'] = google_trends['date'].dt.to_period('M')
+google_trends['ln_interest'] = google_trends.groupby('month')['ln_interest'].transform(lambda x: x - x.mean())
+
[…]

### 2024-04-29 google_trends/code/concat_trends.py

@@ -1,0 +1,59 @@
+import pandas as pd 
+import os 
+import matplotlib.pyplot as plt
+
+# Load each of the files pulled from google trends
+trends_files = os.listdir('raw')
+
+# Load each of the files into a dataframe
+trends_data = []
+for file in trends_files:
+    data = pd.read_csv(f'raw/{file}')
+    trends_data.append(data)
+
+# Concatenate all the dataframes
+trends_data = pd.concat(trends_data)
+trends_data['date'] = pd.to_datetime(trends_data['date'])
+
+# Set multiindex of movie id and date
+trends_data.set_index(['movie_id', 'date'], inplace=True)
+
+new_index = pd.MultiIndex.from_product(trends_data.index.levels)
+trends_data = trends_data.reindex(new_index)
+
+# Fill in missing values with 0
+trends_data.fillna(0, inplace=True)
+
+trends_data.reset_index(inplace=True)
+
+# Filter to the Dark Knight
+dark_knight = trends_data[trends_data['movie_id'] == 20100]
+
+# Plot the data
+# Show data labels yearly on the x axis
+plt.plot(
+    dark_knight['date'], 
+    dark_knight['interest'],
+)
+plt.xticks(
+    dark_knight['date'][::52],
+    [str(date.year) for date in dark_knight['date'][::52]],
+    rotation=45,
+)
+
+plt.xlabel('Date')
+plt.ylabel('Interest')
+plt.title('Interest in The Dark Knight')
+
+# Create datetime object for july 18, 2008
+release_date = pd.to_datetime('2008-07-18')
+plt.axvline(release_date, color='red', linestyle='--')
+
+#plt.show()
+plt.savefig('output/dark_knight.png')
+
+# Save the data
+trends_data.to_csv(
+    'output/trends.csv',
+    index=False,
+)

### 2024-04-29 google_trends/code/pull_trends.py

@@ -1,21 +1,27 @@
 from pytrends.request import TrendReq
 import pandas as pd
 import matplotlib.pyplot as plt
+import os
+import time
+from requests.exceptions import RetryError
 
 # Create pytrends object
-pytrends = TrendReq(hl='en-US', tz=360)
+pytrends = TrendReq(
+    hl='en-US', 
+    tz=360,
+    retries=3,
+)
 
 # Load data from OpusData
-opusdata = pd.read_csv('../OpusData/MovieData.csv')
+opusdata = pd.read_csv('../OpusData/output/opus_cleaned.csv')
 
 # Get list of movie names
-movie_names = opusdata['movie_name'].to_list()
+movie_names = opusdata['cleaned_name'].to_list()
 movie_years = opusdata['production_year'].to_list()
 movie_ids = opusdata['movie_odid'].to_list()
 
-# Create dataframe to store movie interests
-# Has three columns: movie id, date, and interest
-movie_interests_long = pd.DataFrame(columns=['movie_id', 'date', 'interest'])
+# Get list of movie ids already scraped
+existing_output = os.listdir('raw')
 
 for i in range(len(movie_names)):
     # Unpack movie data
@@ -23,21 +29,43 @@ for i in range(len(movie_names)):
     movie_year = movie_years[i]
     movie_id = movie_ids[i]
 
-    # Get Google Trends data
-    pytrends.build_payload(
-        kw_list=[movie_name + " movie"],
-        timeframe=f'{movie_year - 1}-01-01 {movie_year + 1}-12-31',
-    )
+    # Skip if already scraped
+    filename = f'{movie_id}.csv'
+    if filename in existing_output:
+        continue
 
-    # Get interest over time
-    interest_over_time = pytrends.interest_over_time()
+    # Get Google Trends data
+    try:
+        pytrends.build_payload(
+            kw_list=[movie_name + " movie"],
+            timeframe=f'{movie_year - 1}-01-01 {movie_year + 1}-12-31',
+        )
+
+        # Get interest over time
+        interest_over_time = pytrends.interest_over_time()
+    except RetryError:
+        # Too many requests; wait a while and try again
+        print('Too many requests; waiting 60 seconds')
+        time.sleep(60)
+        pytrends.build_payload(
+            kw_list=[movie_name + " movie"],
+            timeframe=f'{movie_year - 1}-01-01 {movie_year + 1}-12-31',
+        )
+
+        # Get interest over time
+        interest_over_time = pytrends.interest_over_time()
+    
+
[…]

### 2024-04-29 tmdb/code/get_tmdb_info.py

@@ -4,228 +4,14 @@ import pandas as pd
 import string
 import csv
 
-YEAR_HANDCODES = movie_years = {
-    'romance and cigarettes': 2005,
-    'slow burn': 2005,
-    'red tails': 2012,
-    'august osage county': 2013,
-    'sin city a dame to kill for': 2014,
-    'edge of tomorrow': 2014,
-    'the last song': 2010,
-    'faster': 2010,
-    'valentines day': 2010,
-    'killers': 2010,
-    'the roommate': 2011,
-    'the beaver': 2011,
-    'the help': 2011,
-    'the eagle': 2011,
-    'no strings attached': 2011,
-    'hanna': 2011,
-    'winnie the pooh': 2011,
-    'sanctum': 2011,
-    'the tree of life': 2011,
-    'abduction': 2011,
-    'the rite': 2011,
-    'unknown': 2011,
-    'paul': 2011,
-    'the big year': 2011,
-    'red riding hood': 2011,
-    'the mechanic': 2011,
-    'super 8': 2011,
-    'upside down': 2012,
-    'priest': 2011,
-    'hop': 2011,
-    'battle los angeles': 2011,
-    'the dilemma': 2011,
-    'movie 43': 2013,
-    'lol': 2012,
-    'quartet': 2012,
-    'project x': 2012,
-    'chronicle': 2012,
-    'stoker': 2013,
-    'the possession': 2012,
-    'promised land': 2012,
-    'vamps': 2012,
-    'the apparition': 2012,
-    'red lights': 2012,
-    'lockout': 2012,
-    'ride along': 2014,
-    'the raven': 2012,
-    'lawless': 2012,
-    'the vow': 2012,
-    'the three stooges': 2012,
-    'the butler': 2013,
-    'safe': 2012,
-    'this is 40': 2012,
-    'parker': 2013, 
-    'riddick': 2013,
-    'the impossible': 2012,
-    'big miracle': 2012,
-    'taken 2': 2012,
-    'ted': 2012,
-    'the equalizer': 2014,
-    'the watch': 2012,
-    'lincoln': 2012,
-    'the dictator': 2012,
-    'the hunger games': 2012,
-    'mirror mirror': 2012,
-    'the campaign': 2012,
-    'life of pi': 2012,
-    'prometheus': 2012,
-    'world war z': 2013,
-    'battleship': 2012,
-    'the avengers': 2012,
-    'john carter': 2012,
[…]

### 2024-05-07 draft/draft.tex
Overview of results

@@
 \usepackage{geometry}
+\usepackage{graphicx}
+
+\author{Gideon Moore}
+\title{Mid-Quarter Check-In}
+\date{May 1, 2024}
 
@@
 
+\maketitle
+
+\section{Overview}
+
+Since our last meeting, I have collected data from OpusData, Google Trends, and TMDB about a sample of major movies from 2006 to 2018.
+I have managed to link these data together, such that I can connect embeddings of movie descriptions from TMDB to search interest from Google Trends in the period around movie publication dates taken from OpusData.
+
+My text embedding algorithm seems to work as-expected; without supervision, it manages to cluster movies by genre tag (as taken from TMDB).
+Moreover, the cross-genre clusters it produces are reasonable--for example, ``action'' and ``adventure'' movies are very close, whereas ``romantic comedy'' movies are on the other side of the chart.
+
+Following Magnolfi et al.
+(2022), I estimated cross-elasticities of demand using a log-linear demand model.
+Cross-elasticity is estimated as a cubic function of normalized Euclidean distance in 6-space.
+The results are encouraging: on the support of the data, the cross-elasticities are monotonically decreasing in distance, as expected.
+
+I have a few next steps in mind, but am very open to other suggestions:
+\begin{itemize}
+    \item At the moment I estimate cross-elasticities only for movies published in 2007.
+I intend to extend this to the whole time span, but under my currend specification movie distance computation is quadratic in the number of movies.
+With some clever dynamic programming this should be much more tractable, but I haven't implemented this yet.
+    \item For now I am using 5,012-dimensional text embeddings reduced to 6 dimensions using PCA.
+Since I created these embeddings, OpenAI has released a new version of their model which allows for custom embedding dimensions.
+I intend to use this feature to create 6-dimensional embeddings directly, rather than using PCA.
+    \item My Google Trends data is incomplete, as they've blocked my IP.
+While I anticipate regaining access at some point, I also think this is the weakest part of my data product at the moment.
+Trends numbers are sort of ``mushy''--what does a 10\% decline in search interest mean?
+How does interest in one movie compare to another?
+    \item At some point I should speak to Amar Venugopal, a third-year PhD student in the metrics group.
+He is currently researching causal inference using text data; now that I have something like an intermediate result, I think I would feel comfortable asking for his input.
+\end{itemize}
+
+\section{Motivation}
+
+As we've discussed previously, I am generally underwhelmed be existing text- and image-data for demand estimation methods.
+I believe the closest analog to my project is Compiani et al.
+(2023); however, as mentioned at our previous meeting, their team generally uses more rudimentary text analysis tools relative to the frontier.
+Moreover, their emphasis is primarily on image data anyway.
+I think their setting is also quite a bit less conducive to identifying parameters of interest; they examine cross-elasticities of demand for various consumer electronics on Amazon.
+However, they treat prices and characteristics as exogeneous.
+Clearly this is not reasonable: the Kindle 7 receives a price drop with the introduction of the Kindle 8, and its features are almost certainly a subset of its successor's.
+In contrast, the movie industry resolves both of these problems: uniform pricing norms mean we can treat prices as exogeneously constant, and movies in theaters at the same time generally had scripts and production approved well in advance of any point where they could have observed the other movie (see: twin movies).
+
+Magnolfi et al.
+(2022) has been very helpful for thinking about demand estimation using embeddings.
+While triplet embeddings are nice to have, I think text embeddings are generally preferrable: you don't need to pay external judges to produce triplets, and triplets are not subject to questions about the survey population.
+So long as the researcher has non-trivial (probably > 1 sentence) text, whether from an API or scraped from the web, embeddings should produce meaningful approximations of product characteristics.
+
+\section{Datasets}
+
+\subsection{OpusData}
+
+OpusData is the vendor behind ``The Numbers.com,'' the premier box office data provider.
+I solicited them for weekly box office returns.
+While I did not want to pay for their data (yet), they also provided a free sample of movies meeting the following criteria:
+\begin{itemize}
+    \item Production year between 2006 and 2018
+    \item Production budget at least \$10 million
+    \item Figures available for box office receipts
+\end{itemize}
+
+Per our discussion at the start of the term, none of these appear to be obviously of concern for selection into the sample. Thus, I have used this sample as the spine to attach other data.
+
+\subsection{Google Trends}
+
+In the absence of box office data, I have used Google Trends to measure search interest in movies.
+After cleaning movie titles from OpusData, I query ``[movie name] movie'' to get a time series of search interest.
+I performed these queries using PyTrends, which worked well for ~1,000 movies, but have since been IP blocked.
+Thus, I will continue to work only with this sample for now.
+
+This measure is somewhat odd because interest is normalized out of 100.
+Thus, we know how well a movie does \emph{relative to its earlier self,} but not relative to other movies.
+Given we are estimating a log-linear demand model (and thus working in percentages) I think this should be fine, but it is worth noting.
+This will be more of an issue if Google's normalization is nonlinear; this I don't have an answer for.
+Similarly I only have three years of data for each movie; I retrieve data for the year before, during, and after the movie's release, and set all other weeks to zero.
+
+The other issue with search interest relative to box office receipts is that we do not observe the release date directly.
+However, I proxy for the week of release using the week with highest interest (breaking ties with the earlier week).
+See the figure below for an example of this process for \textit{The Dark Knight}, plotting the release date from Wikipedia on the chart in red:
+
+\begin{center}
+    \includegraphics[width=.6\textwidth]{../google_trends/output/dark_knight.png}
+\end{center}
+
+\subsection{TMDB}
+
+I'll admit that IMDB is probably the ``preferred'' source for movie data, but their API costs tens of thousands of dollars.
+On the other hand, the community-driven TMDB API is free.
+Using their search interface, I query the name of each movie from the OpusData sample in order to get the description from TMDB.
+There are generally quite concise, but I think they should carry enough information to be meaningful.
+For example, here is the description for \textit{The Dark Knight}:
+
+\begin{quote}
+    Batman raises the stakes in his war on crime.
+With the help of Lt.
+Jim Gordon and District Attorney Harvey Dent, Batman sets out to dismantle the remaining criminal organizations that plague the streets.
+The partnership proves to be effective, but they soon find themselves prey to a reign of chaos unleashed by a rising criminal mastermind known to the terrified citizens of Gotham as the Joker.
+\end{quote}
+
+Indeed, short texts like this are where embeddings do particularly well relative to tradition text-to-data methods: whereas a bag-of-words model would struggle to capture the meaning of this text, embeddings can leverage their under-the-hood training corpus to understand the meaning of this short snippet.
+
+\section{Initial Results}
+
+\subsection{Embeddings}
+
+Before getting to the cross-elasticity estimates, first I should demonstrate that the OpenAI embeddings are capturing something ``meaningful.'' Below, find a t-SNE plot of the embeddings for the movies in the sample.
+I have colored the points by the genre of the movie, as provided by TMDB.
+
+\begin{center}
+    \includegraphics[width=\textwidth]{../tmdb/plots/tsne_genres.png}
+\end{center}
+
+You can see that movies within a genre appear to cluster together.
+Moreover, similar genres appear to group: see the cloud of ``Action/Adventure'' to the bottom-right, and the intersection of ``Horror'' and `Thriller'' in the bottom-left.
+Note certain colors have multiple meanings superimposed (e.g.
+green is both Drama and Documentary); if I wish to show all movies on the plot, there is a limited pallette to choose from, and wanted to present \emph{all} the movies to show that even in the noisiest specification we can see the clusters.
+
+\subsection{Cross-Elasticities}
+
+Creating the cross-movie distances is quite intensive since the computations grow quadratically in the number of movies, so for now I have limited only to movies published in 2007.
+For the full analysis I will likely push the analysis to Sherlock rather than running it on my laptop.
+
+Following Magnolfi et al.
+(2022), I use a log-linear demand model dependent on a cubic in normalized Euclidean distance between goods.
+Embeddings are reduced from 5,012 dimensions to 6 using PCA to follow the authors; just this month, OpenAI has released an updated model which allows embeddings of custom dimensions, so for the final version I will likely use a direct six-dimensional embedding rather than PCA.
+The final regression specification is:
+
+$$\ln(Interest_{jt}) = \alpha_j + \alpha_t + \sum_{k \neq j} \left(\gamma_1 d_{jk} + \gamma_2 d_{jk}^2 + \gamma_3 d_{jk}^3\right) + \varepsilon_{jt}$$
+
+Note here we have omitted both the own- and cross-price terms, as uniform pricing in the movie industry means that all the price terms reduce to constants and are absorbed into the intercept term.
+This is estimated using GMM, but I have not yet implemented the standard errors.
+
+The cross-price elasticity function is given in the figure below:
+\begin{center}
+    \includegraphics[width=.6\textwidth]{../elasticity_estimate/output/elasticity_function.png}
+\end{center}
+
+While initially I was concerned that the cross-elasticity function was not monotone, on inspecting the data I found that the support of the distribution was predominantly between 0.3 and 0.7: precisely the monotonically declining part of the function:
+
+\begin{center}
+    \includegraphics[width=.6\textwidth]{../elasticity_estimate/output/histogram_distances.png}
+\end{center}
+
+Thus, these results are quite encouraging to me: the cross-elasticities are correct-signed and--on the support of the data--monotonically declining in distance, as expected.
+
 \end{document}

### 2024-05-07 elasticity_estimate/code/estimate_elasticity.py

@@ -60,12 +60,14 @@ google_trends['ln_interest'] = google_trends.groupby('month')['ln_interest'].tra
 # De-Mean log interest within a movie
 google_trends['ln_interest'] = google_trends.groupby('movie_id')['ln_interest'].transform(lambda x: x - x.mean())
 
+
+
 # Limit sample to movies with trends and embeddings
 movies_with_trends = google_trends['movie_id'].unique()
 movies_with_embeddings = embeddings['movie_id'].unique()
 movies = np.intersect1d(movies_with_trends, movies_with_embeddings)
 
-
+print(f'Number of movies satisfying all criteria: {len(movies)}')
 
 google_trends = google_trends[google_trends['movie_id'].isin(movies)]
 embeddings = embeddings[embeddings['movie_id'].isin(movies)]
@@ -170,23 +172,37 @@ empirical_distances = []
 for observation in observations:
     empirical_distances.extend(observation.distances)
 
+# Get bottom and top decile
+distances = np.array(empirical_distances)
+bottom_decile = np.percentile(distances, 10)
+top_decile = np.percentile(distances, 90)
+median = np.percentile(distances, 50)
+
 plt.hist(empirical_distances, bins=20)
 plt.xlabel('Distance')
 plt.ylabel('Frequency')
 plt.title('Histogram of Distances')
-plt.show()
+
+plt.axvline(x=bottom_decile, color='r', linestyle='--')
+plt.axvline(x=top_decile, color='r', linestyle='--')
+plt.axvline(x=median, color='g', linestyle='--')
+
+# Add labels
+plt.text(bottom_decile, 0, 'Bottom Decile', rotation=90)
+plt.text(top_decile, 0, 'Top Decile', rotation=90)
+plt.text(median, 0, 'Median', rotation=90)
+
+plt.savefig('output/histogram_distances.png')
+
+
+# New plot
+plt.figure()
 
 # Plot elasticity function
 plt.plot(distance_grid, elasticity)
 plt.xlabel('Distance')
 plt.ylabel('Elasticity')
 plt.title('Estimated Elasticity Function')
-
-# Get bottom and top decile
-distances = np.array(empirical_distances)
-bottom_decile = np.percentile(distances, 10)
-top_decile = np.percentile(distances, 90)
-median = np.percentile(distances, 50)
 
 # Add lines for median and top/bottom deciles
[…]

### 2024-05-07 guru/code/guru_scraper.py
Scraper for Box Office Guru

@@ -1,0 +1,100 @@
+import requests 
+from bs4 import BeautifulSoup
+import re
+import pandas as pd
+import time
+import random
+
+ARCHIVES_ROOT = "http://www.boxofficeguru.com/archives2.htm"
+URL_ROOT = "http://www.boxofficeguru.com/"
+DATE_REGEX = re.compile(r"\w+\. \d{1,2} - \d{1,2}")
+
+def parse_html(filename):
+    with open(f"html/{filename}") as file:
+        html = file.read()
+
+    # Parse the HTML
+    soup = BeautifulSoup(html, "html.parser")
+
+    # Get the table element
+    table = soup.find("table")
+
+    # Convert the table to a DataFrame
+    df = pd.read_html(
+        str(table),
+        match="\$",
+    )[0]
+
+    converters = {c: lambda x: str(x) for c in df.columns}
+    df = pd.read_html(
+        str(table),
+        match="\$",
+        converters=converters
+    )[0]
+
+    # Convert all columns to strings
+    df = df.applymap(str)
+
+    # Drop first row
+    df = df[1:]
+
+    # Convert the first row to column names
+    colnames = df.iloc[0]
+    colnames[2] = "Week Sales"
+    colnames[3] = "Lag Week Sales"
+    df.columns = colnames
+    df = df[1:]
+
+    # Replace string nan with actual NaN
+    df = df.replace("nan", pd.NA)
+
+    # Drop if missing title
+    df = df.dropna(subset=["Title"])
+
+    # Drop if title contains "Top 5" or "Top 10"
+    df = df[~df["Title"].str.contains("Top 5|Top 10")]
+
+    # Drop first column
+    df = df.drop(df.columns[0], axis=1)
+
+    # Convert sales columns to numeric
+    for column in ["Week Sales", "Lag Week Sales", "Cumulative", "AVG"]:
+        df[column] = df[column].str.replace("$", "").str.replace(",", "").astype(pd.Int64Dtype())
+
+    # Add a column for the date
+    df["Date"] = filename.split(".")[0]
+
+    # Convert date to datetime
+    df["Date"] = pd.to_datetime(df["Date"], format="%m%d%y")
+
+    return df
+
+# Pull the archives page
+homepage = requests.get(ARCHIVES_ROOT)
+
+# Get the set of links with text matching the date regex
+homepage_html = BeautifulSoup(homepage.text, "html.parser")
+links = homepage_html.find_all("a", text=DATE_REGEX)
+
[…]

### 2024-05-08 guru/code/clean_data.py
Fully scrape and clean data

@@ -1,0 +1,120 @@
+import pandas as pd
+import os
+from bs4 import BeautifulSoup
+import re
+import progressbar
+
+table_regex = re.compile(r"Top\s*5")
+
+def parse_html(filename):
+    with open(f"html/{filename}") as file:
+        html = file.read()
+
+    # Remove all newlines
+    html = html.replace("\n", " ")
+
+    # Parse the HTML
+    soup = BeautifulSoup(html, "html.parser")
+
+    # Get the list of table elements and select the last one
+    tables = soup.find_all("table")
+    if len(tables) == 0:
+        raise ValueError("No Tables")
+    table = tables[-1]
+
+
+    # Convert the table to a DataFrame
+    df = pd.read_html(
+        str(table),
+        match=table_regex,
+    )
+    #assert len(df) == 1
+    df = df[0]
+
+    converters = {c: lambda x: str(x) for c in df.columns}
+    df = pd.read_html(
+        str(table),
+        match=table_regex,
+        converters=converters
+    )[0]
+
+    # Replace string nan with actual NaN
+    df = df.replace("nan", pd.NA)
+
+    # Drop empty rows at the top
+    while pd.isna(df[1][0]):
+        df = df[1:].reset_index(drop=True)
+
+    # Convert the first row to column names
+    colnames = list(df.iloc[0])
+
+    colnames[2] = "Week Sales"
+    colnames[3] = "Lag Week Sales"
+
+    # If "Dist." in column names, replace with "Distributor"
+    colnames = [col.replace("Dist.", "Distributor") for col in colnames]
+    colnames = ["Cumulative" if "Cume" in col else col for col in colnames]
+    colnames = [col.replace("Cume", "Cumulative") for col in colnames]
+
+    for important_col in ["#", "Title", "Cumulative", "Weeks"]:
+        if filename != "070797.htm":
+            assert important_col in colnames, f"{important_col} not in {colnames}"
+
+    df.columns = colnames
+    df = df[1:]
+
+    # Drop columns to right of "Distributor"
+    if "Distributor" in df.columns:
+        df = df.iloc[:, :df.columns.get_loc("Distributor") + 1]
+
+    # Drop if missing title
+    df = df.dropna(subset=["Title"])
+
+        
+    # Drop if title contains "Top 5" or "Top 10"
+    df = df[~df["Title"].str.contains("Top 5|Top 10|Top 20")]
+
[…]

### 2024-05-08 guru/code/guru_scraper.py

@@ -4,82 +4,39 @@ import re
 import pandas as pd
 import time
 import random
+import os
+import progressbar
 
 ARCHIVES_ROOT = "http://www.boxofficeguru.com/archives2.htm"
 URL_ROOT = "http://www.boxofficeguru.com/"
-DATE_REGEX = re.compile(r"\w+\. \d{1,2} - \d{1,2}")
-
-def parse_html(filename):
-    with open(f"html/{filename}") as file:
-        html = file.read()
-
-    # Parse the HTML
-    soup = BeautifulSoup(html, "html.parser")
-
-    # Get the table element
-    table = soup.find("table")
-
-    # Convert the table to a DataFrame
-    df = pd.read_html(
-        str(table),
-        match="\$",
-    )[0]
-
-    converters = {c: lambda x: str(x) for c in df.columns}
-    df = pd.read_html(
-        str(table),
-        match="\$",
-        converters=converters
-    )[0]
-
-    # Convert all columns to strings
-    df = df.applymap(str)
-
-    # Drop first row
-    df = df[1:]
-
-    # Convert the first row to column names
-    colnames = df.iloc[0]
-    colnames[2] = "Week Sales"
-    colnames[3] = "Lag Week Sales"
-    df.columns = colnames
-    df = df[1:]
-
-    # Replace string nan with actual NaN
-    df = df.replace("nan", pd.NA)
-
-    # Drop if missing title
-    df = df.dropna(subset=["Title"])
-
-    # Drop if title contains "Top 5" or "Top 10"
-    df = df[~df["Title"].str.contains("Top 5|Top 10")]
-
-    # Drop first column
-    df = df.drop(df.columns[0], axis=1)
-
-    # Convert sales columns to numeric
-    for column in ["Week Sales", "Lag Week Sales", "Cumulative", "AVG"]:
-        df[column] = df[column].str.replace("$", "").str.replace(",", "").astype(pd.Int64Dtype())
-
-    # Add a column for the date
-    df["Date"] = filename.split(".")[0]
-
-    # Convert date to datetime
-    df["Date"] = pd.to_datetime(df["Date"], format="%m%d%y")
-
-    return df
+DATE_REGEX = re.compile(r"\w+\.? \d{1,2} - (\w+\.? )?\d{1,2}")
 
 # Pull the archives page
 homepage = requests.get(ARCHIVES_ROOT)
 
+# Replace line breaks with spaces
[…]

### 2024-05-08 guru/code/test_plots.py

@@ -1,0 +1,45 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Load the data
+df = pd.read_csv("data/weekly_sales.csv")
+
+# Convert Date to datetime
+df["Date"] = pd.to_datetime(df["Date"])
+
+# Convert week sales to int
+df["Week Sales"] = df["Week Sales"].str.replace("$", "").str.replace(",", "").astype(int)
+df["ln Week Sales"] = np.log(df["Week Sales"])
+
+# Histogram of weekly sales
+plt.hist(df["ln Week Sales"], bins=20)
+plt.title("Weekly Sales")
+plt.xlabel("ln Week Sales")
+plt.ylabel("Count")
+plt.show()
+
+
+# Sum revenue by week
+weekly_revenue = df.groupby("Date")["Week Sales"].sum()
+ln_weekly_revenue = np.log(weekly_revenue)
+
+# Plot the data
+plt.plot(ln_weekly_revenue)
+plt.title("Revenue per Week")
+plt.xlabel("Date")
+plt.show()
+
+# Open movie info
+movie_info = pd.read_csv("data/movies.csv")
+
+# Bar chart of top 10 distributors
+distributor_counts = movie_info["Distributor"].value_counts()
+top_10 = distributor_counts.head(10)
+top_10.plot(kind="bar")
+# Increase plot area
+plt.tight_layout()
+plt.title("Top 10 Distributors")
+plt.xlabel("Distributor")
+plt.ylabel("Count")
+plt.show()

### 2024-05-08 tmdb/code/get_tmdb_info.py
Move TMDB scraper to Guru data

@@ -10,8 +10,8 @@ def get_movie_info(row):
     Function to apply to each row in the movie data
     Takes info from that row and returns the movie info from TMDB
     """
-    cleaned_name = row['cleaned_name']
-    year = row['production_year']
+    cleaned_name = row['Title']
+    year = row['year']
 
     # Search for movie given name and year
     search = tmdb.Search()
@@ -38,21 +38,31 @@ with open('tmdb.secret', 'r') as f:
     api_key = f.read().strip()
     tmdb.API_KEY = api_key
 
-opusdata = pd.read_csv('../OpusData/output/opus_cleaned.csv')
+gurudata = pd.read_csv('../guru/data/weekly_sales.csv')
+
+# Convert Date to datetime
+gurudata['Date'] = pd.to_datetime(gurudata['Date'])
+
+# Get the first date of each movie
+movies = gurudata.groupby('Title')['Date'].min().reset_index()
+movies['year'] = movies['Date'].dt.year
+
+# Merge premier onto movie data
 
 # Get movie info for each movie
-query_results = opusdata.apply(get_movie_info, axis=1)
+query_results = movies.apply(get_movie_info, axis=1)
 
 # Parse TMDB info into dataframe
-opusdata['tmdb_id'] = [query_result['id'] if query_result else None for query_result in query_results]
-opusdata['tmdb_name'] = [query_result['title'] if query_result else None for query_result in query_results]
-opusdata['tmdb_description'] = [query_result['overview'] if query_result else None for query_result in query_results]
-opusdata['tmdb_release_date'] = [query_result['release_date'] if query_result else None for query_result in query_results]
-opusdata['tmdb_genres'] = [query_result['genre_ids'] if query_result else None for query_result in query_results]
-opusdata['tmdb_language'] = [query_result['original_language'] if query_result else None for query_result in query_results]
+movies = movies.copy()
+movies['tmdb_id'] = [query_result['id'] if query_result else None for query_result in query_results]
+movies['tmdb_name'] = [query_result['title'] if query_result else None for query_result in query_results]
+movies['tmdb_description'] = [query_result['overview'] if query_result else None for query_result in query_results]
+movies['tmdb_release_date'] = [query_result['release_date'] if query_result else None for query_result in query_results]
+movies['tmdb_genres'] = [query_result['genre_ids'] if query_result else None for query_result in query_results]
[…]

### 2024-05-08 tmdb/code/get_tmdb_info.py
Pull descs given Guru data

@@ -3,6 +3,10 @@ from unidecode import unidecode
 import pandas as pd
 import string
 import csv
+from tqdm import tqdm
+import re
+
+PAREN_REGEX = re.compile(r'\([^)]*\)')
 
 
 def get_movie_info(row):
@@ -13,18 +17,52 @@ def get_movie_info(row):
     cleaned_name = row['Title']
     year = row['year']
 
+    # Flag rereleases to change search year constraints
+    rerelease = False
+    if "RE" in cleaned_name or "rerelease" in cleaned_name.lower() or "reissue" in cleaned_name.lower() or "re-release" in cleaned_name.lower():
+        cleaned_name = cleaned_name.replace("RE", "")
+        cleaned_name = cleaned_name.replace("rerelease", "")
+        cleaned_name = cleaned_name.replace("reissue", "")
+        cleaned_name = cleaned_name.replace("re-release", "")
+        rerelease = True
+
+    # Remove anything in parentheses
+    cleaned_name = re.sub(PAREN_REGEX, '', cleaned_name)
+
+    # Remove anything after colon
+    cleaned_name = cleaned_name.split(":")[0]
+
+    # Remove accents
+    cleaned_name = unidecode(cleaned_name)
+
+    # Remove punctuation
+    cleaned_name = cleaned_name.translate(str.maketrans('', '', string.punctuation))
+
+    # Need to wait until after above to successfully flag RE
+    cleaned_name = cleaned_name.lower()
+    if "imax" in cleaned_name:
+        cleaned_name = cleaned_name.replace(" imax", " ")
+    cleaned_name = cleaned_name.replace(" pt", " Part")
+                                        
+    
+    
+    
+
     # Search for movie given name and year
     search = tmdb.Search()
-    search.movie(query=cleaned_name, year=year)
+    if rerelease:
+        search.movie(query=cleaned_name)
+    else:
+        search.movie(query=cleaned_name, year=year)
 
-    if not search.total_results:
-        search.movie(query=cleaned_name, year= year - 1)
+        if not search.total_results:
+            search.movie(query=cleaned_name, year= year - 1)
 
-    if not search.total_results:
-        search.movie(query=cleaned_name, year = year + 1)
+        if not search.total_results:
+            search.movie(query=cleaned_name, year = year + 1)
 
-    if not search.total_results:
-        search.movie(query=cleaned_name, year = year + 2)
+        if not search.total_results:
[…]

### 2024-05-08 tmdb/code/embed_descriptions.py
Updated embedding script

@@ -1,4 +1,4 @@
-from openai import OpenAI
+from openai import OpenAI, BadRequestError
 import pandas as pd
 import pickle
 from tqdm import tqdm
@@ -20,16 +20,26 @@ def get_embedding(description):
         print("No description!")
         return None
 
-    response = client.embeddings.create(
-        input=[description],
-        model="text-embedding-3-small"
-    )
+    try:
+        response = client.embeddings.create(
+            input=[description],
+            model="text-embedding-3-small",
+            dimensions=6
+        )
+    except:
+        print("Could not process description!")
+        print("Description:", description)
+        raise
 
     return response
 
 # Get descriptions
 movie_data = pd.read_csv('descriptions/movie_descriptions.csv', encoding="ISO-8859-1")
 
+# Drop if descriptions are null
+movie_data = movie_data.dropna(subset=['tmdb_description'])
+
+# Embed descriptions
 movie_data['embedding'] = movie_data['tmdb_description'].progress_apply(get_embedding)
 
 with open('embeddings/movie_descriptions.pkl', 'wb') as f:

### 2024-05-10 guru/code/clean_data.py
Move cleaning for TMDB into cleaning script

@@ -3,7 +3,10 @@ import os
 from bs4 import BeautifulSoup
 import re
 import progressbar
+import string
+from unidecode import unidecode
 
+PAREN_REGEX = re.compile(r'\([^)]*\)')
 table_regex = re.compile(r"Top\s*5")
 
 def parse_html(filename):
@@ -79,7 +82,7 @@ def parse_html(filename):
 
     # Convert sales columns to numeric
     for column in ["Week Sales", "Lag Week Sales", "Cumulative", "AVG"]:
-        if "AVG" not in df.columns:
+        if "AVG" not in df.columns and column == "AVG":
             continue
         df[column] = df[column].str.replace("$", "").str.replace(",", "").astype(pd.Int64Dtype())
 
@@ -91,9 +94,39 @@ def parse_html(filename):
 
     return df
 
+def clean_title(cleaned_name):
+
+    # Flag rereleases to change search year constraints
+    rerelease = False
+    if "RE" in cleaned_name or "rerelease" in cleaned_name.lower() or "reissue" in cleaned_name.lower() or "re-release" in cleaned_name.lower():
+        cleaned_name = cleaned_name.replace("RE", "")
+        cleaned_name = cleaned_name.replace("rerelease", "")
+        cleaned_name = cleaned_name.replace("reissue", "")
+        cleaned_name = cleaned_name.replace("re-release", "")
+        rerelease = True
+
+    # Remove anything in parentheses
+    cleaned_name = re.sub(PAREN_REGEX, '', cleaned_name)
+
+    # Remove anything after colon
+    cleaned_name = cleaned_name.split(":")[0]
+
+    # Remove accents
+    cleaned_name = unidecode(cleaned_name)
+
+    # Remove punctuation
+    cleaned_name = cleaned_name.translate(str.maketrans('', '', string.punctuation))
+
+    # Need to wait until after above to successfully flag RE
+    cleaned_name = cleaned_name.lower()
+    if "imax" in cleaned_name:
+        cleaned_name = cleaned_name.replace(" imax", " ")
+    cleaned_name = cleaned_name.replace(" pt", " Part")
+
+    return cleaned_name, rerelease
+
 # Read the data
 filenames = os.listdir("html")
-#filenames = ['070797.htm']
 dfs = []
 for filename in progressbar.progressbar(filenames):
     try:
[…]

### 2024-05-10 tmdb/code/get_tmdb_info.py

@@ -14,40 +14,10 @@ def get_movie_info(row):
     Function to apply to each row in the movie data
     Takes info from that row and returns the movie info from TMDB
     """
-    cleaned_name = row['Title']
-    year = row['year']
-
-    # Flag rereleases to change search year constraints
-    rerelease = False
-    if "RE" in cleaned_name or "rerelease" in cleaned_name.lower() or "reissue" in cleaned_name.lower() or "re-release" in cleaned_name.lower():
-        cleaned_name = cleaned_name.replace("RE", "")
-        cleaned_name = cleaned_name.replace("rerelease", "")
-        cleaned_name = cleaned_name.replace("reissue", "")
-        cleaned_name = cleaned_name.replace("re-release", "")
-        rerelease = True
-
-    # Remove anything in parentheses
-    cleaned_name = re.sub(PAREN_REGEX, '', cleaned_name)
-
-    # Remove anything after colon
-    cleaned_name = cleaned_name.split(":")[0]
-
-    # Remove accents
-    cleaned_name = unidecode(cleaned_name)
-
-    # Remove punctuation
-    cleaned_name = cleaned_name.translate(str.maketrans('', '', string.punctuation))
-
-    # Need to wait until after above to successfully flag RE
-    cleaned_name = cleaned_name.lower()
-    if "imax" in cleaned_name:
-        cleaned_name = cleaned_name.replace(" imax", " ")
-    cleaned_name = cleaned_name.replace(" pt", " Part")
-                                        
+    cleaned_name = row["Cleaned Title"]
+    year = row["year"]
+    rerelease = row["rerelease"]
     
-    
-    
-
     # Search for movie given name and year
     search = tmdb.Search()
     if rerelease:

### 2024-05-10 tmdb/code/embed_descriptions.py
Move back to high-dimensional embeddings

@@ -24,7 +24,6 @@ def get_embedding(description):
         response = client.embeddings.create(
             input=[description],
             model="text-embedding-3-small",
-            dimensions=6
         )
     except:
         print("Could not process description!")

### 2024-05-10 tmdb/code/plot_genres.py
Move Genre plots to Guru data

@@ -1,12 +1,44 @@
 import pickle
 import pandas as pd
 from sklearn.manifold import TSNE
+from sklearn.decomposition import PCA
 import numpy as np
 import matplotlib.pyplot as plt
+import re 
+
+GENRE_REGEX = re.compile(r"\[(\d+)")
 
 # Load in the embeddings
 with open('embeddings/movie_descriptions.pkl', 'rb') as f:
     movie_data = pickle.load(f)
+
+movie_data['genre'] = movie_data['tmdb_genres'].apply(lambda x: re.search(GENRE_REGEX, x).group(1) if re.search(GENRE_REGEX, x) else None)
+movie_data['genre'] = movie_data['genre'].astype(pd.Int64Dtype())
+
+# Label genre integers
+genre_labels = {
+    28: 'Action',
+    12: 'Adventure',
+    16: 'Animation',
+    35: 'Comedy',
+    80: 'Crime',
+    99: 'Documentary',
+    18: 'Drama',
+    10751: 'Family',
+    14: 'Fantasy',
+    36: 'History',
+    27: 'Horror',
+    10402: 'Music',
+    9648: 'Mystery',
+    10749: 'Romance',
+    878: 'Science Fiction',
+    10770: 'TV Movie',
+    53: 'Thriller',
+    10752: 'War',
+    37: 'Western'
+}
+
+movie_data['genre'] = movie_data['genre'].map(genre_labels)
 
 X = movie_data['embedding'].apply(lambda x: x.data[0].embedding).to_list()
 X = np.array(X)

### 2024-05-10 elasticity_estimate/code/estimate_elasticity.py
Start moving elasticity estiamates to guru data

@@ -4,14 +4,13 @@ import pickle
 from sklearn.decomposition import PCA
 from scipy.optimize import minimize
 import matplotlib.pyplot as plt
+import re
 
 # Load the embeddings
 embeddings = pickle.load(open('../tmdb/embeddings/movie_descriptions.pkl', 'rb'))
 
 # Rename movie_odid to movie_id
-embeddings.rename(columns={'movie_odid': 'movie_id'}, inplace=True)
-
-# Merge embedding data onto google tren
+embeddings.rename(columns={'tmdb_id': 'movie_id'}, inplace=True)
 
 raw_embeddings = embeddings['embedding'].apply(lambda x: x.data[0].embedding).to_list()
 
@@ -21,56 +20,93 @@ reduced_embeddings = PCA(n_components=6).fit_transform(raw_embeddings)
 for i in range(6):
     embeddings[f'pca_{i}'] = reduced_embeddings[:, i]
 
-# Load google trends
-google_trends = pd.read_csv('../google_trends/output/trends.csv')
+# Load box office info
+guru = pd.read_parquet('../guru/data/weekly_sales.parquet')
 
-# Filter only to dates with 100 interest
-premier_dates = google_trends[google_trends['interest'] == 100]
+guru = guru.reset_index(drop=False)
 
-# In the event of a tie, take the first date
-premier_dates = premier_dates.groupby('movie_id').first()
-premier_dates.reset_index(inplace=True)
+# Drop if missing theaters
+guru = guru.dropna(subset=['Theaters'])
 
-# Rename columns
-premier_dates.columns = ['movie_id', 'premier_date', 'interest']
-premier_dates = premier_dates[['movie_id', 'premier_date']]
+# Convert theaters and weeks to integers
+guru['Theaters'] = guru['Theaters'].astype(pd.Int64Dtype())
+guru['Weeks'] = guru['Weeks'].astype(pd.Int64Dtype())
 
-# Merge premier_dates onto trends
-google_trends = google_trends.merge(premier_dates, on='movie_id')
-
-# Keep if date is within three months of premier
-google_trends['date'] = pd.to_datetime(google_trends['date'])
-google_trends['premier_date'] = pd.to_datetime(google_trends['premier_date'])
-google_trends['days_from_premier'] = (google_trends['date'] - google_trends['premier_date']).dt.days
-google_trends = google_trends[google_trends['days_from_premier'].between(0, 90)]
-
-# Keep if date has positive interest
[…]

### 2024-05-10 elasticity_estimate/elasticity_estimate.sh
Set up shell scripts

@@ -1,0 +1,1 @@
+python3 code/estimate_elasticity.py

### 2024-05-10 guru/guru.sh

@@ -1,0 +1,2 @@
+# python3 code/guru_scraper.py
+python3 code/clean_data.py

### 2024-05-10 master.sh

@@ -1,0 +1,8 @@
+cd guru
+sh guru.sh
+
+cd ../tmdb
+sh tmdb.sh
+
+cd ../elasticity_estimate
+sh elasticity_estimate.sh

### 2024-05-10 tmdb/tmdb.sh

@@ -1,0 +1,3 @@
+python3 code/get_tmdb_info.py
+python3 code/embed_descriptions.py
+python3 code/plot_genres.py

### 2024-05-10 elasticity_estimate/elasticity_estimate.sh
Moved shell scripts off of python3 to python

@@ -1,1 +1,1 @@
-python3 code/estimate_elasticity.py
+python code/estimate_elasticity.py

### 2024-05-10 guru/guru.sh

@@ -1,2 +1,2 @@
 # python3 code/guru_scraper.py
-python3 code/clean_data.py
+python code/clean_data.py

### 2024-05-10 tmdb/tmdb.sh

@@ -1,3 +1,3 @@
-python3 code/get_tmdb_info.py
-python3 code/embed_descriptions.py
-python3 code/plot_genres.py
+python code/get_tmdb_info.py
+python code/embed_descriptions.py
+python code/plot_genres.py

### 2024-05-10 elasticity_estimate/code/estimate_elasticity.py
Introduced primary keys for movies

@@ -43,38 +43,6 @@ print(f"Missing log earnings: {guru['ln_earnings'].isnull().sum()}")
 guru = guru.dropna(subset=['Theaters', 'ln_earnings'])
 guru = guru.drop("Weeks", axis=1)
 
-# Disambiguate movies with the same name while computing movie spell lengths
-max_spell_length = 10000
-counter = 0
-while max_spell_length > 60:
-    print("Iteration: ", counter)
-    wide_release_dates = guru[guru['Theaters'] >= 600].groupby('Clean Title')['Date'].min()
-    # Rename the column
-    wide_release_dates = wide_release_dates.reset_index()
-    wide_release_dates.columns = ['Clean Title', 'Date_wide']
-
-    # Drop any showings before the wide release
-    guru = guru.merge(
-        wide_release_dates, 
-        on='Clean Title',
-    )
-    
-    guru = guru[guru['Date'] >= guru['Date_wide']]
-
-    # Compute weeks since wide release
-    guru['weeks_since_release'] = (guru['Date'] - guru['Date_wide']).dt.days // 7
-
-    # If a movie has been out for more than 50 weeks, it's likely a different movie with the same name
-    # Thus, flag title with counter
-    guru['Clean Title'] = np.where(guru['weeks_since_release'] > 60, guru['Clean Title'] + f'_{counter}', guru['Clean Title'])
-
-    max_spell_length = guru['weeks_since_release'].max()
-    counter += 1
-    
-    # Drop wide release date
-    print(guru.head())
-    print(guru.columns)
-    guru = guru.drop('Date_wide', axis=1)
 
 # Histogram of weeks since release
 plt.hist(guru['weeks_since_release'], bins=20)

### 2024-05-10 guru/code/clean_data.py

@@ -5,6 +5,7 @@ import re
 import progressbar
 import string
 from unidecode import unidecode
+import numpy as np
 
 PAREN_REGEX = re.compile(r'\([^)]*\)')
 table_regex = re.compile(r"Top\s*5")
@@ -142,20 +143,59 @@ df = pd.concat(dfs)
 df['Theaters'] = df['Theaters'].astype(pd.Int64Dtype())
 df['Weeks'] = df['Weeks'].astype(pd.Int64Dtype())
 
+cleaning_content = [clean_title(title) for title in df["Title"]]
+results = [content for content in cleaning_content]
+
+df['Clean Title'] = [content[0] for content in cleaning_content]
+df['Rerelease'] = [content[1] for content in cleaning_content]
+
+# Hash clean title into an ID
+df['movie_id'] = df['Clean Title'].apply(hash)
+
+# Disambiguate movies with the same name while computing movie spell lengths
+max_spell_length = 10000
+counter = 0
+while max_spell_length > 60:
+    print("Iteration: ", counter)
+    wide_release_dates = df[df['Theaters'] >= 600].groupby('movie_id')['Date'].min()
+    # Rename the column
+    wide_release_dates = wide_release_dates.reset_index()
+    wide_release_dates.columns = ['movie_id', 'Date_wide']
+
+    # Drop any showings before the wide release
+    df = df.merge(
+        wide_release_dates, 
+        on='movie_id',
+    )
+    
+    df = df[df['Date'] >= df['Date_wide']]
+
+    # Compute weeks since wide release
+    df['weeks_since_release'] = (df['Date'] - df['Date_wide']).dt.days // 7
+
+    # If a movie has been out for more than 50 weeks, it's likely a different movie with the same name
+    # Thus, flag title with counter
+    df['movie_id'] = np.where(df['weeks_since_release'] > 60, df['movie_id'] + 1, df['movie_id'])
+
+    max_spell_length = df['weeks_since_release'].max()
+    counter += 1
+    
+    if max_spell_length > 60:
+        df = df.drop('Date_wide', axis=1)
+
+# Rename Date_wide to Wide Release Date
+df = df.rename(columns={"Date_wide": "Wide Release Date"})
+
 # With the data cleaned, we can parse it into something normalized
-weekly_sales = df[["Title", "Date", "Week Sales", "Theaters", "Weeks"]]
-weekly_sales.set_index(["Title", "Date"], inplace=True)
[…]

### 2024-05-10 tmdb/code/embed_descriptions.py

@@ -33,7 +33,7 @@ def get_embedding(description):
     return response
 
 # Get descriptions
-movie_data = pd.read_csv('descriptions/movie_descriptions.csv', encoding="ISO-8859-1")
+movie_data = pd.read_parquet('descriptions/movie_descriptions.parquet')
 
 # Drop if descriptions are null
 movie_data = movie_data.dropna(subset=['tmdb_description'])

### 2024-05-10 tmdb/code/get_tmdb_info.py

@@ -2,7 +2,6 @@ import tmdbsimple as tmdb
 from unidecode import unidecode
 import pandas as pd
 import string
-import csv
 from tqdm import tqdm
 import re
 
@@ -14,7 +13,7 @@ def get_movie_info(row):
     Function to apply to each row in the movie data
     Takes info from that row and returns the movie info from TMDB
     """
-    cleaned_name = row["Cleaned Title"]
+    cleaned_name = row["Clean Title"]
     year = row["year"]
     rerelease = row["rerelease"]
     
@@ -46,16 +45,12 @@ with open('tmdb.secret', 'r') as f:
     api_key = f.read().strip()
     tmdb.API_KEY = api_key
 
-gurudata = pd.read_csv('../guru/data/weekly_sales.csv')
-
-# Convert Date to datetime
-gurudata['Date'] = pd.to_datetime(gurudata['Date'])
+movies = pd.read_parquet('../guru/data/movies.parquet')
+movies = movies.reset_index(drop=False)
 
 # Get the first date of each movie
-movies = gurudata.groupby('Title')['Date'].min().reset_index()
-movies['year'] = movies['Date'].dt.year
+movies['year'] = movies['Wide Release Date'].year
 
-# Merge premier onto movie data
 
 # Get movie info for each movie
 tqdm.pandas()
@@ -74,4 +69,4 @@ movies['tmdb_language'] = [query_result['original_language'] if query_result els
 movies['no_results'] = movies['tmdb_id'].isnull()
 
 # Write descriptions to file
-movies.to_csv('descriptions/movie_descriptions.csv', index=False)
+movies.to_parquet('descriptions/movie_descriptions.parquet', index=False)

### 2024-05-10 guru/code/clean_data.py
Hand correct a bunch of distributors to ensure uniqueness

@@ -6,6 +6,8 @@ import progressbar
 import string
 from unidecode import unidecode
 import numpy as np
+import datetime
+import hashlib
 
 PAREN_REGEX = re.compile(r'\([^)]*\)')
 table_regex = re.compile(r"Top\s*5")
@@ -110,7 +112,7 @@ def clean_title(cleaned_name):
     cleaned_name = re.sub(PAREN_REGEX, '', cleaned_name)
 
     # Remove anything after colon
-    cleaned_name = cleaned_name.split(":")[0]
+    #cleaned_name = cleaned_name.split(":")[0]
 
     # Remove accents
     cleaned_name = unidecode(cleaned_name)
@@ -126,22 +128,39 @@ def clean_title(cleaned_name):
 
     return cleaned_name, rerelease
 
-# Read the data
-filenames = os.listdir("html")
-dfs = []
-for filename in progressbar.progressbar(filenames):
+def parse_file(filename):
     try:
-        dfs.append(parse_html(filename))
+        return parse_html(filename)
+
     except ValueError:
         print(f"Failed to parse {filename}, No Tables")
+    except KeyboardInterrupt:
+        print("Interrupted")
+        raise
     except:
         print(f"Failed to parse {filename}")
         raise
+
+# Read the data
+filenames = os.listdir("html")
+
+#with ThreadPoolExecutor() as executor:
+#    dfs = list(executor.map(parse_file, filenames))
+dfs = []
+for filename in progressbar.progressbar(filenames):
+    dfs.append(parse_file(filename))    
 
 df = pd.concat(dfs)
 
 df['Theaters'] = df['Theaters'].astype(pd.Int64Dtype())
 df['Weeks'] = df['Weeks'].astype(pd.Int64Dtype())
+
+# Hand-correct a duplicate movie
+# I'm fairly sure this is right; the 1998-03-23 releases duplicate The Apostle, one on week 4 the other on week 14
+# Going to the previous week (http://www.boxofficeguru.com/031698.htm) we can see that the Apostle is on week 13, so the week 4
+# release is the error. Similarly, we can see there are 3 movies on week 3. 
+# Only one, Caught Up, has the same distributor as the week 4 release of The Apostle
+df.loc[(df['Title'] == 'The Apostle') & (df['Date'] == datetime.datetime.strptime('1998-03-23', "%Y-%m-%d")) & (df['Distributor'] == 'Live'), 'Title'] = "Caught Up"
 
[…]

### 2024-08-30 draft/draft.tex
Commiting old residual changes

--- 2024-04-29 draft/draft.tex
@@
 \usepackage{geometry}
+\usepackage{graphicx}
+\usepackage[style=authoryear]{biblatex}
+\usepackage{booktabs}
+
+% Double space
+\usepackage{setspace}
+\doublespacing
+
+% Read in bibliography
+\addbibresource{movies_bib.bib}
+
+\author{Gideon Moore}
+\title{Industrial Organization III: \\ Final Paper}
+\date{June 9, 2024}
 
@@
 
+\maketitle
+
+\begin{abstract}
+    I propose a novel method for estimating competition between products using text embeddings.
+Applying this method to the movie industry, I estimate impacts of competition which are reasonable in both sign and magnitude.
+Facing additional competitors lowers a film's revenue, and more similar competitors are more damaging.
+I find that if a movie could replace a competitor in the 90th percentile of similarity with one in the 10th percentile of similarity, its revenue would increase by 0.32\% to 0.44\%.
+\end{abstract}
+
+\section{Introduction}
+
+Demand estimation is a central pillar of the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
+However, demand estimation is often limited to markets with easily quantifiable characteristics, such as the storage in the market for hard drives or wattage in the market for light bulbs.
+We should not ignore markets with difficult-to-quantify characteristics simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a firm monopolizing comedy films is likely to be bad for consumers.
+Thus, it is important economists develop tools to estimate demand in markets for goods with difficult-to-quantify characteristics.
+
+I propose a new method for quantifying product characteristics using text descriptions.
+By embedding text descriptions of a product in characteristic space, we as econometricians can identify which products are most similar.
+Thus, we can predict which products compete most closely with others and estimate how the presence of one product impacts the sales of another.
+
+While this method has broad applicability, I use the film industry as a case study.
+Movies are a natural laboratory for several reasons.
+First, movies have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if you haven't seen \emph{Spider-Man}, you likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
+Second, movie characteristics are often amorphous: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a movie is--and harder still given the lack of clean data.
+Finally, the movie industry has specific market characteristics which make demand estimation straightforward conditional on my method.
+Uniform pricing shuts down the endogeneous pricing channel of competition, and long production pipelines force studios to choose characteristics without knowing anything about their competitors.
+
+Using data from BoxOfficeGuru and TheMovieDB, I estimate the impact of competitor characteristics on film revenue for each weekend from 1997-2019.
+I find substitution patterns which are plausible in both magnitude and direction, and which are robust to the choice of distance metric.
+As a movie becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
+I find that replacing a competitor in the 90th percentile of similarity with one in the 10th percentile of similarity would increase revenue by 0.32\% to 0.44\%.
+
+\section{Literature}
+
+This paper contributes to three strands of literature.
+First, I implement a novel use for text embeddings, innovating on the text-as-data literature in economics.
+Second, given my demand estimation method is based on position in latent characteristic space, I am also closely connected to existing work on spatial competition.
+Finally, given my film case study, I also contribute to the literature on the economics of the film industry.
+
+\subsection{Text as Data}
+
+This paper's primary contribution is to the literature on text as data, well summarized in \textcite{gentzkow2019EL}.
+The statistical analysis of text has a long history; the earliest example I have found is \textcite{mendenhall1887S} using word frequency to predict the author of a mysterious text.
+This is an early example of \emph{bag of words} methods, which treat text as a collection of words rather than a structured object.
+Other types of bag of words methods include term frequency-inverse document frequency (TF-IDF), which downweights words that appear in many documents, and Latent Dirichlet Allocation (LDA), which uses a generative model to infer topics from word concurrences.
+
+I would argue these models work well for long documents such as patents and books, but struggle when working with (what I will call) ``terse'' texts.
+If one paragraph includes the word ``wizardry'' while the other includes the word ``sorcery,'' a bag of words model will not capture the similarity between the two paragraphs.
+In large enough documents, there is some idea that the two words are likely to appear in similar contexts, but this is difficult to rely on when the sample text is only one or two sentences.
+
+For terse contexts such as mine, I prefer using \emph{text embedding methods} as pioneered by Google's word2vec architecture \parencite{mikolov2013}.
+Rather than counting individual words, word2vec identifies similar words using co-occurances in a training sample, then maps these words to a lower-dimensional space.
+This allows for the identification of similar texts even when they share few or no words, since it is able to identify that synonyms like ``wizardry'' and ``sorcery'' map to the same concept.
+
+word2vec, while an improvement on bag of words, still overlooks \emph{interactions} between words.
+While ``climate'' and ``change'' have meanings of their own, ``climate change'' is a largely distinct phenomenon.
+This is the motivation behind ``transformer'' architectures, pioneered again by Google in 2019's BERT sentence embeddings \parencite{devlin2019}.
+However, BERT limits itself to processing one sentence at a time and aggregating; thus, I prefer OpenAI's recent GPT-3 embeddings which can process entire paragraphs.
+
+This is an improvement upon much of the economics literature, which still relies on bag of words methods.
+For example, \textcite{baker2016QJE} uses counts of ``uncertain'' words to estimate economic policy uncertainty.
+Similarly, \textcite{ederer2022} uses counts of word frequences to construct a vector used as a proxy for product characteristics and estimate cross-price elasticities.
+
+The most similar paper to mine is \textcite{compiani2023} which uses text characteristics to estimate demand for technology products on Amazon.
+My paper innovates on theirs in multiple important ways.
+First, I have a much broader market; rather than focusing on fewer than 3,000 purchases in a specific Amazon retailer, I examine the entire United States movie market for multiple decades.
+Second, assumptions for demand estimation are much more plausible in the movie industry than in consumer electronics: the Kindle 8 is clearly designed endogeneously to the characteristics of the Kindle 7, whereas movies have long production pipelines which lock in product charactertics before competitor characteristics are known.
+Third, I use more advanced text embedding methods, pushing past word frequency and BERT-based methods to OpenAI's GPT-3 embeddings.
+Fourth, I (will...) benchmark my substitution patterns against ``real world'' user preference data from the MovieLens data, confirming my estimates capture actual substitution patterns.
+Finally, I estimate substitution patterns much more flexibly, allowing substitution to vary non-linearly and non-monotonically with distance between two products.
+
+\subsection{Spatial Competition}
+
+Though this paper does not focus on \emph{physical space}, I also contribute to the literature on spatial competition in the tradition of \textcite{hotelling1929E} and \textcite{salop1979BE}, where firms interact based on relative position in space.
+By projecting descriptions into characteristic space, I model high-dimensional competition between products.
+
+My closest predecessor studying competition in embedding space is \textcite{magnolfi2022} (hereafter MMS), who use t-Stochastic Triplet Embeddings to estimate competition in the cereal industry.
+I take heavy inspiration from their model, however substitute frontier text embedding methods from OpenAI for their triplet-based embeddings.
+Both they and I follow \textcite{pinkse2002E}, who first estimate product cross-elasticities as a function of distance in characteristic space.
+
+\subsection{Movies}
+
+The final strand of literature where I wish to highlight my contribution is on the economics of the film industry.
+
+Though I do not contribute to it, I benefit greatly from research documenting the uniform pricing puzzle \parencite{orbach2007IRLE, gil2009MS, ho2018MR}.
+This literature extensively documents movie theaters' use of uniform pricing across heterogeneous movies.
+While each paper suggests potential explanations for this phenomenon, my findings are agnostic to the true cause of uniform pricing; I simply use it as a convenient assumption to estimate demand.
+
+Much of my understanding of the economics of film literature comes from a survey by \textcite{mckenzie2012ES}.
+I contribute to the ongoing literature founded by \textcite{prag1994CE} of determinents of box office demand.
+The authors find well-rated and well-advertised movies generally sell more tickets, and conditional on those two factors other observables are generally insignificant.
+\textcite{elberse2003MS} model both the supply and the demand for movies, finding that theatres' decision to screen a movie at all is an important factor for revenue.
+\textcite{devany1996E, devany1997EI, devany1999JCE, devany2004EDC} examine many factors of box office revenue, including the impact of word-of-mouth, the hazard rate of movie retirement, and estimating the film production function.
+\textcite{ravid2004B} study the impact of film \emph{content} on performance, finding that violence increases film profitability while sexual content does not.
+
+To the best of my knowledge, I am the first paper to examine the \emph{interplay} between competing films: how much is my revenue impacted by the existence of box office competitors?
+
+\section{The Movie Industry}
+
+In the tradition of industrial organization, I use an industry case study of the film industry to highlight the value of my method.
+Text characteristics are distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{The Dark Knight} and \emph{Ant-Man} are big-budget superhero action movies, but even a quick glance at the movie descriptions makes clear they are not substitutable.
+This is particularly important given movies are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
+Recognizing this, we expect firms put more effort into the text descriptions of their movies than comparable firms might in more traditional goods markets, as these decisions are distinctively important to customer decisionmaking.
+Finally, the market for movies has two distinctive features which make it a strong candidate for demand estimation: long production pipelines and a tradition of uniform pricing.
+
+\subsection{Production Timelines}
+
+When estimating the impact of competitors' characteristics on demand for a movie, its characteristics must be exogeneous to the characteristics of its competitors.
+If characteristics are chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor.
+
+Fortunately, movie production timelines are both long and secretive.
+Since studios must purchase a script, hire a cast, and shoot all \emph{before} promoting a movie, it is difficult for a film to change its characteristics in response to its competitors' characteristics.
+I posit the existence of ``twin movies'' as evidence of this characteristic inelasticity.
+A pair of ``twin movies'' are two movies released in close proximity with very similar characteristics; the most famous examples, \emph{A Bug's Life} and \emph{Antz}, both released in fall of 1998 and feature early CGI animation of ants rebelling against oppression.
+
+Twin movies are a persistent feature of the film industry: the Wikipedia page for the phenomenon lists nearly 300 pairs of twin movies.
+Studios are generally inclined to avoid releasing a twin movie due to fear of excess competition.
+In 2016, French director Xavier Giannoli released his film \emph{Marguerite}, inspired by the true story of a New York Socialite turned failed opera singer, in the same year as 20th Century Fox's \emph{Florence Foster Jenkins} retelling the same story.
+Giannoli told the \emph{Independent}: ``For me, it was terrible...I work a lot as a writer to find completely original stories.
+I don't want the audience to have the feeling, `oh, I saw that!''' \parencite{mottram2016I}.
+
+Given filmmakers' expressed distaste for twin movies, why do they continue to exist?
+It suggests that studios are unable to change their movies' characteristics in response to their competitors' characteristics; at the point I learn of \emph{A Bug's Life}, it is too late for me to adjust \emph{Antz} to be more distinct.
+Thus, our demand estimation coefficients are identified--they do not capture the strategic interplay of positioning and substitutability, but instead the direct impact of competition on film sales.
+
+This should produce testable hypotheses: conditional on the characteristics of a movie, the characteristics of its competitors should ``look like'' the characteristics of movies releasing in other weekends.
+One complication to this is we think movies have seasonality, so we can't simply compare movies within the same year in different months.
+Thus, we'd likely want to compare movies released in July of one specific year against movies released in July of \emph{other} years.
+This will work less well if movie content shifts significantly over time (for example, if we think there are more superhero movies in the 2010s than in the 2000s).
+
+One threat to this strategy would be if firms lack flexibility on their films' \emph{content}, but can move the film's release date \emph{forward or backward in time}.
+If I discover I am releasing my children's movie opposite \emph{Frozen}, I may wish to postpone my release date.
+This would threaten the assumption of exogeneous characteristics among competitors.
+I think this is unlikely to be a major concern for a couple of reasons.
+First, there are certain prime release dates which are likely to be more profitable than others; for example, releasing a family movie in November is much better than doing so in January, as you can capture the holiday season.
+Thus, if a movie targeted a holiday release, moving its release date to avoid competition would potentially be quite costly since there no longer exists consumer demand.
+Second, the continued existence of twin movies further suggests this is not a margin firms seem to adjust significantly on: if studios could easily move release dates to avoid competition, we would expect to see fewer twin movies than we do.
+
+\subsection{Uniform Pricing}
+
+An ongoing puzzle within the film industry is the existence of a uniform pricing standard: within a timeslot, cinemas generally charge the same price for all movies regardless of excess demand.
+As a salient example, ``opening night'' showings of movies often sell out, yet theaters do not raise prices.
+Similarly, movies late in their run often have significant excess capacity, yet theathers do not lower prices.
+
+There are many explanations for this phenomenon; I am agnostic about which of these is the true explanation.
+What is important is that ticket price \emph{is} uniform across films.
+This pricing anomaly supports two simplifying assumptions which do not hold water in most other settings.
+
+First, we know \emph{movies do not compete on price.} Under more flexible pricing, we might believe two movies in close priximity would lower prices to compete for the same audience, raising the quantity sold overall.
+However, since prices are fixed we know that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
+
+Second, we can \emph{use revenue as a proxy for quantity.} As described in \textcite{bond2021JME}, this assumption is often perilous since a variable markup means that revenue may not reflect true production.
+However, since prices are fixed, we can understand revenue as simply the quantity of tickets sold multiplied by some scalar price.
+Thus, even without observing quantity, I am confident that observed revenue is a close substitute.
+
+\section{Model}
+
+\subsection{Log-Linear Demand}
+
+Following MMS, I estimate a log-linear demand model for movies.
+I start by including fixed effects for each movie and weekend.
+The distance between products $i$ and $j$ is given by some metric $d_{ij}$; a competitor's impact is given by via a flexible function of distance $f(d_{ij})$.
+The sum of individual competitors' impacts yields the total impact of competition on a given movie.
+Again following MMS, I estimate $f$ as a cubic polynomial in distance.
+
+The first benefit of the cubic polynomial functional form here is that this assumption is falsifiable; if $f$ is not monotonically attenuating, our spatial competition model is likely misspecified.
+The second benefit of this functional form is it is estimable linearly; by distributing the grand sum over each individual term and factoring out the $\gamma$ coefficients, demand becomes a simple linear equation in $d_{ij}$, $d_{ij}^2$, and $d_{ij}^3$.
+Thus, traditional matrix inversion techniques remain effective.
+The final benefit is that the cubic polynomial is flexible; it imposes no structure on sign, derivative, or magnitude.
+Thus, I can capture potentially quite rich substitution patterns, as well as evaluate my model fit against ``reasonable'' priors.
+
+Unlike cereals, demand for movies is heavily based on \emph{novelty}; that is, a movie will generally sell more the week of release than two months later.
+Thus, I include a fixed effect $\lambda_{t - r(j)}$ to control for time since release, where $r(j)$ is the release date of movie $j$.
+
+The final model of demand for movie $j$ in time $t$ is as follows:
+
+$$\ln(q_{jt}) = \alpha_j + \alpha_t + \sum_{k \neq j} f(d_{jk}; \gamma) + \lambda_{t - r(j)} + \varepsilon_{jt}$$
+
+$$f(d_{jk}; \gamma) = \gamma_0 + \gamma_1 d_{jk} + \gamma_2 d_{jk}^2 + \gamma_3 d_{jk^k}$$
+
+In the above specification, I purposely do not include price on the right hand side due to the uniform pricing assumption discussed above.
+We can interpret $f(d_{jk}; \gamma)$ as the cross-elasticity between movies $j$ and $k$; as movie $k$ becomes closer or further from movie $j$, $f$ maps how demand for movie $j$ changes.
+Thus, we generally expect to see $f$ monotonically attenuating in distance; as two movies become more distant, they should become increasingly insulated from each other.
+
+\subsection{Age Interaction and Estimation}
+
+The above model leaves something to be desired if we think \emph{new movies are more relevant for competitors as well}.
+For example, my Science Fiction movie is much more unhappy to release the same weekend as \emph{Star Wars} than to release two months after \emph{Star Wars}.
+To capture this, I need to account for the age of the competitors as well by scaling $f$ by some factor $\delta_{t - r(k)}$.
+To preserve power, I should topcode $\delta$ at some reasonable value based on what I see in the data; e.g., I may group movies with 10+ weeks.
+I can inform my topcoding threshold by looking for plateaus in the empirical distribution of sales as movies age.
+
+In this case, our new model is:
+
+$$\ln(q_{jt}) = \alpha_i + \alpha_t + \sum_{k \neq j} \delta_{t - r(k)} f(d_{jk}; \gamma) + \lambda_{t - r(j)} + \varepsilon_{jt}$$
+
+Note this model is no longer linearly estimable, as $\delta_{t - r(k)}$ interacts multiplicatively with the $\gamma$ coefficients in $f$.
+I could manually pull out the fixed effects, then optimize over the possible values of $\delta$, $\lambda$, and $\gamma$, however I believe I have a more clever solution.
+
+Note for any fixed $\delta$ vector, we know the model is linear in $\gamma$ and $\lambda$, which makes it easy to identify the least-squares solution using matrix inversion.
+Thus, I can perform a two-step procedure to identify the least-squares coefficients of my nonlinear model.
+
+First, I choose a $\delta$.
+Then, using that $\delta$, I compute the least squares solution for $\gamma$ and $\lambda$.
+This gives me a sum of squared errors for this set of coefficients.
+I then use this sum of squared errors as the minimization objective to optimize over \emph{$\delta$ alone}.
+Since OLS always minimizes least squares \emph{within} a $\delta$ estimate, the optimal model must be the one which minimizes this quantity \emph{across} $\delta$s.
+The psuedocode for this algorithm is something like:
+
+\begin{verbatim}
+def fit_model(delta, data):
+    for t in data:
+        weekend = data[t]
+        for j in weekend[movies]:
+            gamma_0_val = 0
+            gamma_1_val = 0
+            gamma_2_val = 0
+            gamma_3_val = 0
+            for k != j in weekend[movies]:
+                gamma_0_val += delta[movies[k][age]]
+                gamma_1_val += delta[movies[k][age]] * distance(j, k)
+                gamma_2_val += delta[movies[k][age]] * distance(j, k)^2
+                gamma_3_val += delta[movies[k][age]] * distance(j, k)^3
+
+    model = Panel(
+        y = ln(data[revenue]),
+        x = [
+            gamma_0_val,
+            gamma_1_val,
+            gamma_2_val,
+            gamma_3_val,
+            lambda[data[age]],
+        ],
+        fe = [data[movie], data[weekend]]
+    )
+
+    return model
+
+best_delta = minimize(
+    lambda d: fit_model(d).sse, 
+    data, 
+    np.zeros(LEN_DELTA)
+).x
+
+best_model = fit_model(best_delta)
+\end{verbatim}
+
+I have not yet estimated this model, but I believe this procedure should be consistent. Unfortunately analytic standard errors likely won't work, but I can always bootstrap them.
+
+\section{Data}
+
+\subsection{BoxOfficeGuru}
+
+I scrape weekend box office receipts from 1997-2024 from BoxOfficeGuru.com, a website maintained by Gitesh Pandya.
+Pandya is a film consultant specializing in releasing Indian films for the North American market.
+For each weekend, the website lists the top 10-20 movies by North American box office revenue.
+I have nearly the complete set of weekends over this 27 year span; however, when Gitesh in on vacation (a couple weekends a year) the website is not updated.
+An example of the data is visible in figure \ref{fig:guru}.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{images/guru.jpg}
+    \caption{BoxOfficeGuru Page for May 17-29, 2019}
+    \label{fig:guru}
+\end{figure}
+
+\subsection{TheMovieDB}
+
+I retrieve movie characteristics from the API of TheMovieDB, a user-generated database of movies.
+Specifically, I clean movie names and years retrieved from BoxOfficeGuru and query the search API for the top result released in the relevant year.
+For each movie, I can thus retrieve the description, genres, and original release language.
+Example movie descriptions are visible in table \ref{tab:tmdb_desc}.
+
+\begin{table}
+    \begin{tabular}{lp{11cm}}
+        \toprule 
+        \textbf{Movie} & \textbf{Description} \\
+        \midrule
+        \emph{Frozen} & Young princess Anna of Arendelle dreams about finding true love at her sister Elsa’s coronation.
+Fate takes her on a dangerous journey in an attempt to end the eternal winter that has fallen over the kingdom.
+She's accompanied by ice delivery man Kristoff, his reindeer Sven, and snowman Olaf.
+On an adventure where she will find out what friendship, courage, family, and true love really means.
+\\
+        \emph{The Notebook} & An epic love story centered around an older man who reads aloud to a woman with Alzheimer's.
+From a faded notebook, the old man's words bring to life the story about a couple who is separated by World War II, and is then passionately reunited, seven years later, after they have taken different paths.
+\\
+        \emph{Avengers: Endgame} & After the devastating events of Avengers: Infinity War, the universe is in ruins due to the efforts of the Mad Titan, Thanos.
+With the help of remaining allies, the Avengers must assemble once more in order to undo Thanos' actions and restore order to the universe once and for all, no matter what consequences may be in store.
+\\
+        \emph{A Bug's Life} & On behalf of ``oppressed bugs everywhere," an inventive ant named Flik hires a troupe of warrior bugs to defend his bustling colony from a horde of freeloading grasshoppers led by the evil-minded Hopper.
+\\
+        \bottomrule
+    \end{tabular} 
+    \caption{Example Movie Descriptions from TheMovieDB}
+    \label{tab:tmdb_desc}
+\end{table}
+
+\subsection{Sample Selection}
+
+To avoid the impact of the COVID-19 pandemic, I restrict my sample to movies released in 2019 or before.
+Similarly, I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
+
+Though I cannot find documentation for their methodology, genres on TMDB appear to be ordered; that is, the ``primary'' genre of a movie is listed first on the page.
+This is evident based on genre ordering reversals; for example, \emph{Godzilla x Kong: The New Empire} (a recent Kaiju movie) lists ``action'' before ``adventure;'' on the other hand, \emph{Kingdom of the Planet of the Apes} lists ``adventure'' before ``action.'' Thus, we know the ordering of genres is idiosyncratic by movie.
+When I must assign a genre to a movie, I use the first genre listed on TMDB.
+
+\section{Text Embeddings}
+
+I use OpenAI's embedding model \texttt{text-embedding-3-small} API to embed movie descriptions.
+This produces a 1,536-dimensional vector for each movie description.
+The OpenAI embedding procedure trains a transformer to predict adjacent texts based on a large corpus, such that vectors with similar cosine similarity are likely to have similar next words \parencite{neelakantan2022ao, kusupati2022ao}.
+
+I prefer the OpenAI embeddings to other text-as-data methods for a few reasons.
+First, movie descriptions are quite terse; most are around a paragraph, with a few as short as one or two sentences.
+Thus, traditional bag-of-words methods are likely to suffer, as word or bigram counts would be remarkably sparse.
+Because the OpenAI embedding model is trained on an enormous corpus of text, it can infer when descriptions are similar even when they share \emph{no} words--for example, a movie about ``wizardry'' would be similar to one about ``sorcery,'' which would not be true under a bag-of-words model.
+
+Secondly, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level!), while the OpenAI embeddings are generally more context-aware.
+Given the movie descriptions tend to be longer than a single sentence, I believe the OpenAI embeddings are likely to be more able to capture the full context of the movie description.
+Even so, I should likely also perform my procedure using BERT embeddings in my robustness checks.
+
+In the long run I will likely want to use \texttt{text-embedding-3-large} for better performance.
+For exploratory purposes, I use the smaller model due to cost; the large model costs roughly 10x as much as the small model, though produces a 3,072-dimensional embedding.
+
+As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering.
+I find that the embeddings do indeed cluster by genre, as visible in figure \ref{fig:tsne}.
+Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{../tmdb/plots/tsne_genres.png}
+    \caption{t-SNE of OpenAI Embeddings by Genre}
+    \label{fig:tsne}
+\end{figure}
+
+\subsection{Similarity Measures}
+
+I consider two methods of measuring embedding similarity.
+My first method is the most similar to MMS: I reduce the dimensionality of the 1,536-dimensional embedding to 6 dimensions using principal component analysis.
+From there, I can compute Euclidean distance between the PCA-reduced embeddings as a measure of similarity.
+
+While this is most similar to the MMS approach, my preferred specification uses cosine similarity of the full embeddings.
+This method uses all information contained in each embedding, rather than just the first six principal components.
+Moreover, OpenAI suggests the use of cosine similarity to compare embeddings in their documentation.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+    \begin{center}
+        \begin{tabular}{@{}lll@{}}
+        \toprule
+        \textbf{Baseline Movie}    & \textbf{Nearest Neighbor \#1}       & \textbf{Nearest Neighbor \#2}              \\ \midrule
+        \textit{Frozen}            & \textit{The Little Mermaid}         & \textit{The Princess Diaries 2}            \\
+        \textit{The Notebook}      & \textit{Collateral Beauty}          & \textit{I Still Believe}                   \\
+        \textit{Avengers: Endgame} & \textit{X-MEN: Days of Future Past} & \textit{Guardians of the Galaxy, Volume 3} \\
+        \textit{A Bug's Life}      & \textit{The Spongebob Movie}        & \textit{DC League of Super Pets}           \\ \bottomrule
+        \end{tabular}
+        \caption{Nearest Neighbors to Baseline Movies by Euclidean Distance}
+        \label{tab:euclidean}
+    \end{center}
+\end{table}
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+    \begin{center}
+        \begin{tabular}{@{}lll@{}}
+        \toprule
+        \textbf{Baseline Movie}    & \textbf{Nearest Neighbor \#1}   & \textbf{Nearest Neighbor \#2}       \\ \midrule
+        \textit{Frozen}            & \textit{Frozen II}              & \textit{Mirror Mirror}              \\
+        \textit{The Notebook}      & \textit{Message in a Bottle}    & \textit{The Longest Ride}           \\
+        \textit{Avengers: Endgame} & \textit{Avengers: Infinity War} & \textit{Captain America: Civil War} \\
+        \textit{A Bug's Life}      & \textit{The Ant Bully}          & \textit{Antz}                       \\ \bottomrule
+        \end{tabular}
+        \caption{Nearest Neighbors to Baseline Movies by Cosine Similarity}
+        \label{tab:cosine}
+    \end{center}
+\end{table}
+
+Nearest neighbors to some common movies under Euclidean and cosine distance are visible in tables \ref{tab:euclidean} and \ref{tab:cosine}, respectively.
+Euclidean distance appears to be capturing ``something''--the closest movies to \emph{Frozen} are both princess movies, and the closest movies to \emph{A Bug's Life} are both animated children's movies.
+However, I find the cosine-based neighbors to be more compelling: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
+Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War} respectively.
+This improvement in neighbor quality leads me to prefer cosine similarity for my analysis.
+
+\section{Elasticity Results}
+
+First I run the analysis using Euclidean distance in six-dimensional space, following MMS.
+The regression results are visible in table \ref{tab:results_euclid}.
+The joint regression is highly significant, though the individual coefficients are iffier; given only the coefficient on $\gamma_0$ is significant, it may not be \emph{distance} that matters in this specification, but instead just the \emph{number of competitors}.
+However, it is also possible the other coefficients are noisy because they covary.
+Since the results are difficult to interpret in numeric form, I plot the elasticity curve implied by the $\gamma$ coefficients in figure \ref{fig:cross_elasticity_euclid}.
+While the curve is not monotonically decreasing \emph{everywhere}, we do find that it is monotonically decreasing on the range containing the majority of the data.
+As hypothesized, we see that as a movie becomes more distant, it is less influential on the sales of its competitors.
+Down the line I will need to compute uncertainty in this curve.
+Given there is likely negative correlation in errors between $\gamma$ coefficients (e.g.
+overestimating $\gamma_0$ means we're probably underestimating $\gamma_1$), simply propogating the analytic uncertainty will drastically overstate the true variance.
+I can get around this by bootstrapping, though given the computational intensity of the model I haven't done this yet.
+
+Putting some magnitudes on these numbers, suppose Disney were releasing \emph{Frozen} opposite \emph{Pok\'emon: The Movie 2000}--a movie in the top decile of similarity.
+If Disney were instead facing the horror movie \emph{Premonition}--a movie in the bottom decile of similarity to \emph{Frozen}--they would expect a 0.44\% revenue increase.
+
+\begin{table}
+    \begin{center}
+    \input{../elasticity_estimate/output/euclid_regression.tex}
+
+    \caption{Regression Results, Euclidean Distance}
+    \label{tab:results_euclid}
+    \end{center}
+\end{table}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../elasticity_estimate/output/cross_elasticity_euclid.png}
+    \caption{Cross-Elasticity by Euclidean Distance}
+    \label{fig:cross_elasticity_euclid}
+    \end{center}
+\end{figure}
+
+Moving now to the specification using cosine distance, the regression results are visible in table \ref{tab:results_cosine}.
+Again, we see that the results are highly significant.
+Moreover, here the coefficients on \emph{distance} are also significant, even with the colinearity.
+The elasticity curve implied by the $\gamma$ coefficients is visible in figure \ref{fig:cross_elasticity}.
+While no longer monotonically decreasing, we see the pattern is still decreasing on net from the 10th percentile to the 90th percentile of the data.We also see decreasing behavior in the tails, as desired.
+
+Again putting some concrete meaning on these numbers, suppose Disney were releasing \emph{Frozen} opposite \emph{Inkheart}--a movie in the top decile of cosine similarity.
+If Disney were instead facing the gory historical action flick \emph{Medieval}--a movie in the bottom decile of cosine similarity to \emph{Frozen}--they would expect a 0.32\% revenue increase.
+
+\begin{table}
+    \begin{center}
+    \input{../elasticity_estimate/output/cosine_regression.tex}
+
+    \caption{Regression Results, Cosine Similarity}
+    \label{tab:results_cosine}
+    \end{center}
+\end{table}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../elasticity_estimate/output/cross_elasticity_zoom.png}
+
+    \caption{Cross-Elasticity by Distance}
+    \label{fig:cross_elasticity}
+    \end{center}
+\end{figure}
+
+Note the magnitudes of these elasticity estimates are both quite similar and precisely estimated!
+Moreover, the $R^2$ of the two regressions is also nearly identical.
+This suggests that the distance metric we use is not particularly critical for our results.
+
+It is reasonable to ask how well these trends are explained by movie genre alone.
+When I include a dummy for ``only film of genre'' showing in a given weekend, I find it does not materially change these results.
+Down the line I will likely want to do something more intensive--e.g., compute how similar genres are, then do something with ``genre distance'' alone rather than using the text information.
+Even if the genre \emph{can} explain much of the variation, I still believe this method is a contribution in that researchers are more likely to have raw, uncleaned text data than to have properly coded genre data.
+
+\section{MovieLens}
+
+While the use of text descriptions as a proxy for substitition patterns is appealing, we cannot know from only the price and quantity data we have whether filmgoers are \emph{actually} substituting between the movies we think they are.
+To address this, I intend to use the MovieLens review dataset to estimate movie similarity based on \emph{user preferences} rather than text descriptions.
+
+The fundamental assumption is that if many people rate both movie A and movie B distinctively highly, we can infer that the two movies are similar (and thus, substitutable).
+For example, if the majority of people who give \emph{Star Wars} five stars also give \emph{Star Trek} five stars (and vice versa), we can infer that the two movies are substitutable.
+
+Note this is not \emph{totally} correct.
+What we really want to do is find pairs of movies with high \emph{idiosyncratic} draws--everyone agrees that \emph{Citizen Kane} is a good movie, but if we can find movies most people think are three stars but the same people think are five stars, we can infer that the two movies are substitutable.
+
+There are many textbook recommendation algorithms I could use as benchmarks. The most promising is likely some form of topic-specific PageRank.
+
+Once I have my review-based measure of movie similarity, I can plot this measure of similarity against my text-based measure of similarity.
+If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
+
+\section{Conclusion}
+
+Based on my findings this quarter, I am increasingly confident there is ``something'' here: using boilerplate embeddings of unstructured text characteristics, I pick up significant substitution patterns in the movie industry.
+If my work is successful, I would be the first to estimate cross-movie elasticities, while also providing a new method for estimating demand in industries with unstructured product characteristic text.
+
+Over the summer, I hope to develop my paper in a few specific ways:
+\begin{itemize}
+    \item Enrich the model by interacting competitor age with cross-elasticity as discussed in section 4.2
+    \item Shore up arguments about substitution by comparing text-based similarity to review-based similarity as discussed in section 8
+    \item Confirm in the data that a movie's characteristics appear to be uncorrelated with the characteristics of its competitors as discussed in section 3.1
+\end{itemize}
+
+On a broader timeline, I also want to think about some way to address marketing.
+While the characteristics of the movies themselves may be set in advance, the characteristics \emph{emphasized in marketing} may be different.
+For example, if two romantic comedies come out in one weekend, both may benefit if one emphasizes the romantic half of the movie while the other emphasizes the comedic half.
+I am still thinking about clever ways to address this.
+TMDB contains some movie trailers, so one approach would be to see if similar \emph{movies} also have similar \emph{trailers}, even if they're released at roughly the same time.
+
+\printbibliography
+
 \end{document}

### 2024-08-30 archive/google_trends/code/pull_trends.py

@@ -9,7 +9,8 @@ from requests.exceptions import RetryError
 pytrends = TrendReq(
     hl='en-US', 
     tz=360,
-    retries=3,
+    retries=5,
+    backoff_factor = 1
 )
 
 # Load data from OpusData

### 2024-08-30 elasticity_estimate/code/elasticity_euclidean.py

@@ -1,0 +1,265 @@
+import pandas as pd
+import numpy as np
+import pickle
+from sklearn.decomposition import PCA
+from sklearn.manifold import TSNE
+from scipy.optimize import minimize
+import matplotlib.pyplot as plt
+import statsmodels.api as sm
+import re
+import linearmodels as lm
+import time
+from stargazer.stargazer import Stargazer
+
+N_DIMS = 6
+
+# Load the embeddings from parquet
+embeddings = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
+
+# Drop if embeddings are null
+embeddings = embeddings.dropna(subset=['embedding'])
+
+raw_embeddings = np.array(embeddings['embedding'].values)
+raw_embeddings = np.array([np.array(x) for x in raw_embeddings])
+
+start_time = time.time()
+# On CPU, recommended to PCA first
+# https://medium.com/rapids-ai/tsne-with-gpus-hours-to-seconds-9d9c17c941db
+#reduced_embeddings = PCA(n_components=50).fit_transform(raw_embeddings)
+#print("Starting t-SNE...")
+#reduced_embeddings = TSNE(n_components=N_DIMS, n_jobs=-1, method='exact', random_state=0).fit_transform(reduced_embeddings)
+#print("Finished t-SNE!")
+#print("Runtime: ", time.time() - start_time)
+
+reduced_embeddings = PCA(n_components=N_DIMS).fit_transform(raw_embeddings)
+
+
+for i in range(N_DIMS):
+    embeddings[f'pca_{i}'] = reduced_embeddings[:, i]
+
+# Load box office info
+guru = pd.read_parquet('../guru/data/weekly_sales.parquet')
+
+guru = guru.reset_index(drop=False)
+
+print(guru.dtypes)
+
+# Compute log interest
+guru['ln_earnings'] = np.log(guru['Week Sales'])
+
+# Drop if missing weeks
+guru = guru.dropna(subset=['Weeks'])
+
+# Histogram of weeks since release
+plt.hist(guru['Weeks'], bins=20)
+plt.xlabel('Weeks Since Wide Release')
+plt.ylabel('Frequency')
+plt.title('Histogram of Weeks Since Wide Release')
+plt.savefig('output/weeks_histogram.png')
+
+# Limit sample to movies with trends and embeddings
+movies_with_trends = guru['movie_id'].unique()
+movies_with_embeddings = embeddings['movie_id'].unique()
[…]

### 2024-08-30 elasticity_estimate/code/estimate_elasticity.py

@@ -11,8 +11,7 @@ import re
 import linearmodels as lm
 import time
 import statsmodels.formula.api as smf
-
-N_DIMS = 6
+from stargazer.stargazer import Stargazer
 
 # Load the embeddings from parquet
 tmdb = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
@@ -20,6 +19,46 @@ tmdb = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
 # Drop if embeddings are null
 tmdb = tmdb.dropna(subset=['embedding'])
 tmdb = tmdb.reset_index(drop=True)
+
+# Filter to primary genre
+tmdb['tmdb_genre'] = tmdb['tmdb_genres'].apply(lambda x: x[0] if len(x) > 0 else None)
+
+genre_labels = {
+    28: 'Action',
+    12: 'Adventure',
+    16: 'Animation',
+    35: 'Comedy',
+    80: 'Crime',
+    99: 'Documentary',
+    18: 'Drama',
+    10751: 'Family',
+    14: 'Fantasy',
+    36: 'History',
+    27: 'Horror',
+    10402: 'Music',
+    9648: 'Mystery',
+    10749: 'Romance',
+    878: 'Science Fiction',
+    10770: 'TV Movie',
+    53: 'Thriller',
+    10752: 'War',
+    37: 'Western'
+}
+
+# Replace genre with label
+tmdb['tmdb_genre'] = tmdb['tmdb_genre'].map(genre_labels)
+
+# Create bar chart of genres
+genre_counts = tmdb['tmdb_genre'].value_counts()
+genre_counts.plot(kind='bar')
+plt.xlabel('Genre')
+plt.ylabel('Count')
+plt.title('Count of Movies by Genre')
+
+# Label the genres
+plt.xticks(ticks=range(len(genre_counts)), rotation=45)
+
+plt.savefig('output/genre_counts.png')
 
 # Get array of movie ids
 movie_ids = tmdb['movie_id'].copy()
@@ -46,6 +85,15 @@ guru = pd.read_parquet('../guru/data/weekly_sales.parquet')
 guru = guru.reset_index(drop=False)
 
 print(guru.dtypes)
+
+# Merge on movie genre
+guru = guru.merge(tmdb[['movie_id', 'tmdb_genre']], on='movie_id', how='left')
+
+# Create dummy for whether another movie of that genre is showing within each week
[…]

### 2024-08-30 elasticity_estimate/code/estimate_elasticity.py
First attempt at nonlinear optimization

@@ -1,17 +1,12 @@
 import pandas as pd
 import numpy as np
-import pickle
-from sklearn.decomposition import PCA
-from sklearn.manifold import TSNE
 from scipy.optimize import minimize
-from scipy.spatial.distance import cosine
 import matplotlib.pyplot as plt
-import statsmodels.api as sm
-import re
 import linearmodels as lm
-import time
-import statsmodels.formula.api as smf
 from stargazer.stargazer import Stargazer
+
+# For weeks greater than this, we truncate to this
+WEEK_THRESHOLD = 5
 
 # Load the embeddings from parquet
 tmdb = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
@@ -72,13 +67,6 @@ distances = 1 - raw_embeddings @ raw_embeddings.T
 # Normalize distances
 distances = distances / np.max(distances)
 
-#reduced_embeddings = PCA(n_components=N_DIMS).fit_transform(raw_embeddings)
-
-#embedding_frame = pd.DataFrame(reduced_embeddings, columns=[f'pca_{i}' for i in range(N_DIMS)])
-
-#for i in range(N_DIMS):
-#    embeddings[f'pca_{i}'] = reduced_embeddings[:, i]
-
 # Load box office info
 guru = pd.read_parquet('../guru/data/weekly_sales.parquet')
 
@@ -107,6 +95,9 @@ plt.xlabel('Weeks Since Wide Release')
 plt.ylabel('Frequency')
 plt.title('Histogram of Weeks Since Wide Release')
 plt.savefig('output/weeks_histogram.png')
+
+# Truncate weeks to limit degrees of freedom
+guru['Weeks'] = np.minimum(guru['Weeks'], WEEK_THRESHOLD)
 
 # Limit sample to movies with trends and embeddings
 movies_with_trends = guru['movie_id'].unique()
@@ -188,65 +179,93 @@ for example in examples:
         print(f'{closest_movie_name}: {closest_distance}')
         distances_from_top_decile[closest_index] = np.inf
 
-# For each movie-date, compute the leave-one-out set of distances within the date
-gamma0 = []
-gamma1 = []
-gamma2 = []
-gamma3 = []
 
-for row in guru.itertuples():
-    movie_id = row.movie_id
-    date = row.Date
+def regress_given_lambda(age_coefficients, guru):
 
-    # Get index of movie_id in movie_ids
[…]

### 2024-08-31 elasticity_estimate/code/estimate_elasticity.py
Minor optimizations

@@ -4,6 +4,10 @@ from scipy.optimize import minimize
 import matplotlib.pyplot as plt
 import linearmodels as lm
 from stargazer.stargazer import Stargazer
+import warnings
+
+# Turn off un-actionable FutureWarning
+warnings.simplefilter(action='ignore', category=FutureWarning)
 
 # For weeks greater than this, we truncate to this
 WEEK_THRESHOLD = 5
@@ -58,6 +62,9 @@ plt.savefig('output/genre_counts.png')
 # Get array of movie ids
 movie_ids = tmdb['movie_id'].copy()
 
+# Create a map from movie_ids to indices
+movie_id_to_index = {movie_id: i for i, movie_id in enumerate(movie_ids)}
+
 raw_embeddings = np.array(tmdb['embedding'].values)
 raw_embeddings = np.array([np.array(x) for x in raw_embeddings])
 
@@ -98,6 +105,9 @@ plt.savefig('output/weeks_histogram.png')
 
 # Truncate weeks to limit degrees of freedom
 guru['Weeks'] = np.minimum(guru['Weeks'], WEEK_THRESHOLD)
+
+# Convert date to datetime
+guru['Date'] = pd.to_datetime(guru['Date'])
 
 # Limit sample to movies with trends and embeddings
 movies_with_trends = guru['movie_id'].unique()
@@ -180,6 +190,14 @@ for example in examples:
         distances_from_top_decile[closest_index] = np.inf
 
 
+# Per Nick: could include lambda in the regression and target that those coefficients are 1
+# For each date, pre-compute the values for each movie pair
+dates = guru['Date'].unique()
+date_movie_dict = {}
+for date in dates:
+    date_movies = guru[guru['Date'] == date]
+    date_movie_dict[date] = date_movies
+
 def regress_given_lambda(age_coefficients, guru):
 
     guru = guru.copy()
@@ -197,10 +215,15 @@ def regress_given_lambda(age_coefficients, guru):
         date = row.Date
 
         # Get index of movie_id in movie_ids
-        movie_index = np.where(movie_ids == movie_id)[0][0]
+        movie_index = movie_id_to_index[movie_id]
 
         # Get all movies on the date
-        date_movies = guru[guru['Date'] == date]
+        date_movies = date_movie_dict[date]
+
+
+
+
+
 
         competitor_distances = []
         for comparison_movie in date_movies.itertuples():
@@ -211,7 +234,7 @@ def regress_given_lambda(age_coefficients, guru):
[…]

### 2024-09-01 elasticity_estimate/code/estimate_elasticity.py
First working implementation of recommendation engine

@@ -5,12 +5,15 @@ import matplotlib.pyplot as plt
 import linearmodels as lm
 from stargazer.stargazer import Stargazer
 import warnings
+import cProfile
+import pstats
 
 # Turn off un-actionable FutureWarning
 warnings.simplefilter(action='ignore', category=FutureWarning)
 
 # For weeks greater than this, we truncate to this
-WEEK_THRESHOLD = 5
+# Note zero-indexing; thus, this is actually the *fifth* week
+WEEK_THRESHOLD = 4
 
 # Load the embeddings from parquet
 tmdb = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
@@ -102,6 +105,9 @@ plt.xlabel('Weeks Since Wide Release')
 plt.ylabel('Frequency')
 plt.title('Histogram of Weeks Since Wide Release')
 plt.savefig('output/weeks_histogram.png')
+
+# We want their weeks to also be zero-indexed so we can use it as an index
+guru['Weeks'] = guru['Weeks'] - 1
 
 # Truncate weeks to limit degrees of freedom
 guru['Weeks'] = np.minimum(guru['Weeks'], WEEK_THRESHOLD)
@@ -220,35 +226,18 @@ def regress_given_lambda(age_coefficients, guru):
         # Get all movies on the date
         date_movies = date_movie_dict[date]
 
-
-
-
-
-
-        competitor_distances = []
-        for comparison_movie in date_movies.itertuples():
-            # If looking at the same movie, skip
-            if movie_id == comparison_movie.movie_id:
-                continue
-
-            # Get the relevant age coefficient
-            competitor_age_coef = age_coefficients[comparison_movie.Weeks]
-
-            comparison_index = movie_id_to_index[comparison_movie.movie_id]
-
-            if np.isinf(distances[movie_index, comparison_index]):
-                print(f'Infinite distance between {movie_id} and {comparison_movie.movie_id}')
-
-            competitor_distances.append(competitor_age_coef * distances[movie_index, comparison_index])
-
-
-        competitor_distances = np.array(competitor_distances)
+        # Can ignore that self is included since distance(x, x) = 0
+        competitor_age_coefs = age_coefficients[date_movies['Weeks']]
+        competitor_indices = np.array([movie_id_to_index[movie_id] for movie_id in date_movies['movie_id']])
+        competitor_distances = distances[movie_index, competitor_indices]
[…]

### 2024-09-02 draft/draft.tex
Summarized collaborative filtering in draft

@@
 
-\section{MovieLens}
+\section{Benchmarking: Collaborative Filtering}
 
-While the use of text descriptions as a proxy for substitition patterns is appealing, we cannot know from only the price and quantity data we have whether filmgoers are \emph{actually} substituting between the movies we think they are.
-To address this, I intend to use the MovieLens review dataset to estimate movie similarity based on \emph{user preferences} rather than text descriptions.
+``Recommendation engines'' are a common tool in the tech world, used to predict what products a user will like.
+In some sense, we can think of this project as harnessing a new type of recommendation engine, identifying movies which are likely to be substitutable.
+Algorithms in this space are broadly divided into two categories: collaborative filtering and content-based filtering.
+The text embeddings used above are a form of content-based filtering: using the ``content'' of the movie (its text description) to predict its appeal.
+However, one way to benchmark the quality of this approach is to compare them to the results of a collaborative filtering algorithm.
+Unlike content-based filtering, collaborative filtering doesn't include any information about the content of a movie; instead, it uses the preferences of other users to predict what a user will like.
+For example, if I like \emph{The Dark Knight}, and most people who like \emph{The Dark Knight} also like \emph{Inception}, a collaborative filtering algorithm would predict I would like \emph{Inception}.
 
-The fundamental assumption is that if many people rate both movie A and movie B distinctively highly, we can infer that the two movies are similar (and thus, substitutable).
-For example, if the majority of people who give \emph{Star Wars} five stars also give \emph{Star Trek} five stars (and vice versa), we can infer that the two movies are substitutable.
+MovieLens produces a public-use movie review dataset; it contains more than 25 million (CHECK THIS NUMBER) 1-to-5 star reviews of movies by users.
+Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions.
+Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods.
+If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
 
-Note this is not \emph{totally} correct.
-What we really want to do is find pairs of movies with high \emph{idiosyncratic} draws--everyone agrees that \emph{Citizen Kane} is a good movie, but if we can find movies most people think are three stars but the same people think are five stars, we can infer that the two movies are substitutable.
+\subsection{UV Decomposition}
 
-There are many textbook recommendation algorithms I could use as benchmarks. The most promising is likely some form of topic-specific PageRank.
+For my collaborative filtering algorithm, I use UV-decomposition for matrix completion as described in Leskovec (CITE HERE).
+Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
+However, the matrix is generally sparse; most users have not reviewed most movies.
+UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix.
+This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
+By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.
 
-Once I have my review-based measure of movie similarity, I can plot this measure of similarity against my text-based measure of similarity.
-If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
+Note that UV decomposition relies on a contraction mapping: for a fixed $V$ matrix, we choose the $U$ matrix which minimizes the sum of squared errors.
+We then fix that $U$ matrix and choose an optimal $V$ matrix, iterating this process until convergence.
+This procedure is guaranteed to converge, though it may converge to a local minimum.
+Thus, we generally initialize the $U$ and $V$ matrices with random values, then run the algorithm multiple times to ensure we find something approximating the global minimum.
+
+With our decomposition in hand, it is simple to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
+If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
+The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
+
+\subsection{Validation}
+
+Rather than blindly trusting our decomposition, we can validate it by comparing predicted reviews against actual reviews.
+I hold out 5\% of the reviews as a test set, then train on the remaining 95\%.
+I then predict the held-out reviews using the trained UV decomposition and compare the predicted reviews to the actual reviews.
+If the predicted reviews are highly correlated with the actual reviews, we can be more confident in the UV decomposition.
+
+The relationship between predicted and actual reviews is visible in figure FIGURE.
+As you can see, there exists a strong positive correlation between the predicted reviews and the actual reviews.
+The strength of this correlation varies based on the number of reviews a movie has received in the data.
+Points are colored by the number of reviews their movie received; as is visible in the figure, movies with few reviews have a quite weak correlation between predicted and actual reviews.
+However, once we trim off the least popular 5\% of movies, the correlation between predicted and actual reviews becomes quite a bit stronger.
+This makes sense: our model is (as expected) better at predicting a movie's appeal when we have more data on that movie.
+
+These results affirm for me that our collaborative filtering results are likely to be quite predictive of a movie's appeal.
+Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
+
+\subsection{Comparison}
+
+LOREM IPSUM
 

### 2024-09-04 compare_text_cf/code/compare_text_cf.py
Test UV Decomposition

@@ -1,0 +1,30 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Load the text similarities
+text_similarities = pd.read_parquet('../movie_text_distances/output/movie_text_distances.parquet')
+text_similarities = text_similarities.rename(columns={'distance': 'text_distance'})
+text_similarities = text_similarities.sort_values(['tmdb_id_1', 'text_distance'], ascending=[True, True])
+print(text_similarities.head())
+
+collaborative_similarities = pd.read_csv('../movielens/output/UV_decomp_similarity.csv')
+collaborative_similarities['collaborative_distance'] = 1 - collaborative_similarities['similarity']
+collaborative_similarities = collaborative_similarities.rename(columns={'tmdbId_x': 'tmdb_id_1', 'tmdbId_y': 'tmdb_id_2'})
+collaborative_similarities = collaborative_similarities[['tmdb_id_1', 'tmdb_id_2', 'collaborative_distance']]
+collaborative_similarities = collaborative_similarities.sort_values(['tmdb_id_1', 'collaborative_distance'], ascending=[True, False])
+print(collaborative_similarities.head())
+
+# Merge the two similarity matrices
+similarity = text_similarities.merge(collaborative_similarities, on=['tmdb_id_1', 'tmdb_id_2'], how='inner')
+
+corr = similarity['text_distance'].corr(similarity['collaborative_distance'])
+
+# Plot the two distances against each other
+plt.scatter(similarity['text_distance'], similarity['collaborative_distance'], alpha=0.1)
+plt.xlabel('Text Distance')
+plt.ylabel('Collaborative Distance')
+plt.title(f'Collaborative vs Text Distance (Correlation: {corr:.2f})')
+plt.show()
+
+
+

### 2024-09-04 draft/draft.tex
Wrote Up Topic-Specific PageRank

--- 2024-08-30 draft/draft.tex
@@
 
-\section{MovieLens}
+\section{Benchmarking: Collaborative Filtering}
 
-While the use of text descriptions as a proxy for substitition patterns is appealing, we cannot know from only the price and quantity data we have whether filmgoers are \emph{actually} substituting between the movies we think they are.
-To address this, I intend to use the MovieLens review dataset to estimate movie similarity based on \emph{user preferences} rather than text descriptions.
+``Recommendation engines'' are a common tool in the tech world, used to predict what products a user will like.
+In some sense, we can think of this project as harnessing a new type of recommendation engine, identifying movies which are likely to be substitutable.
+Algorithms in this space are broadly divided into two categories: collaborative filtering and content-based filtering.
+The text embeddings used above are a form of content-based filtering: using the ``content'' of the movie (its text description) to predict its appeal.
+However, one way to benchmark the quality of this approach is to compare them to the results of a collaborative filtering algorithm.
+Unlike content-based filtering, collaborative filtering doesn't include any information about the content of a movie; instead, it uses the preferences of other users to predict what a user will like.
+For example, if I like \emph{The Dark Knight}, and most people who like \emph{The Dark Knight} also like \emph{Inception}, a collaborative filtering algorithm would predict I would like \emph{Inception}.
 
-The fundamental assumption is that if many people rate both movie A and movie B distinctively highly, we can infer that the two movies are similar (and thus, substitutable).
-For example, if the majority of people who give \emph{Star Wars} five stars also give \emph{Star Trek} five stars (and vice versa), we can infer that the two movies are substitutable.
+MovieLens produces a public-use movie review dataset; it contains more than 25 million (CHECK THIS NUMBER) 1-to-5 star reviews of movies by users.
+Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions.
+Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods.
+If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
 
-Note this is not \emph{totally} correct.
-What we really want to do is find pairs of movies with high \emph{idiosyncratic} draws--everyone agrees that \emph{Citizen Kane} is a good movie, but if we can find movies most people think are three stars but the same people think are five stars, we can infer that the two movies are substitutable.
+\subsection{Topic-Specific PageRank}
 
-There are many textbook recommendation algorithms I could use as benchmarks. The most promising is likely some form of topic-specific PageRank.
+Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies.
+PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages.
+The algorithm works by randomly walking through the network, with a probability of jumping to a random node at each step.
+The importance of a node is the proportion of time the random walker spends at that node.
+TSPR is an extension of the algorithm which always ``jumps'' to a specific set of nodes rather than a random node.
 
-Once I have my review-based measure of movie similarity, I can plot this measure of similarity against my text-based measure of similarity.
-If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
+For my application, I implement TSPR as follows. To find movies similar to specific movie $x$,:
+\begin{itemize}
+    \item The algorithm randomly walks over the bipartite network of users and movies. 
+    \item At each step, the algorithm has a 15\% chance of jumping back to $x$
+    \item If the algorithm does not jump to $x$, it takes a random step to a neighbor of its current node
+    \item Random draws are weighted by the number of stars in the review; e.g., a ``5-star'' review is five times as likely to be drawn as a ``1-star'' review
+\end{itemize}
+
+Running this algorithm until convergence provides a measure of how central any given movie $y$ is relative to movie $x$.
+However, we are still not quite done, as some movies are just \emph{inherently} central; if everyone likes \emph{Forrest Gump,} then it's not particularly informative that fans of movie $x$ \emph{also} like \emph{Forrest Gump}.
+Thus, I divide the TSPR score of $y$ by $y$'s \emph{non}-topic specific PageRank score.
+This gives a measure of how much more central $y$ is to $x$ than it is to the average movie.
+
+\subsection{TSPR vs. Text Embeddings}
+
+Once I have each movie's TSPR score for each other movie, I can compare these scores to the cosine similarity of the text embeddings.
+Note that the TSPR scores are not symmetric; that is, the TSPR score of movie $x$ to movie $y$ is not necessarily the same as the TSPR score of movie $y$ to movie $x$.
+Thus, for each pair of movies, I actually plot \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
+
+A common weakness of collaborative filtering algorithms like TSPR is that they require significant data before they become effective.
+Thus, I compute TSPR values only for movies which receive at least 10,000 reviews in the MovieLens data.
+This leaves me with 258 movies, or 66,564 pairs of movies.
+To assess the importance of this margin, I then plot the coefficient of regressing TSPR on text similarity for thresholds 10,000 to 50,000.
+I also plot the number of movie pairs included in the sample for each threshold value, to assess the trade-off of ``more ratings per movie'' vs.
+``more movies.'' The results of this plot are visible in figure \ref{fig:tspr_threshold}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../movielens/output/threshold_vs_coefficient.png}
+    \caption{Coefficient of TSPR on Text Similarity by Threshold}
+    \label{fig:tspr_threshold}
+    \end{center}
+\end{figure}
+
+This plot shows that the coefficient on text similarity for predicting TSPR is consistently and significantly positive for nearly all thresholds.
+Thus, the content of the text embeddings \emph{does} seem to have predictive power for which movies are similar based on revealed preference; that is, the patterns I see in the text-based model are not coincidental and reflect consumers' true preferences.
+
+Moreover, while the number of pairs included remains large enough to have useful confidence intervals, the coefficient is increasing in the threshold value.
+If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like our text-based cosine similarity measure.
+This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per movie to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.
+
+% \subsection{UV Decomposition}
+
+% For my collaborative filtering algorithm, I use UV-decomposition for matrix completion as described in Leskovec (CITE HERE).
+Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
+However, the matrix is generally sparse; most users have not reviewed most movies.
+UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix.
+This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
+By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.
+
+% Note that UV decomposition relies on a contraction mapping: for a fixed $V$ matrix, we choose the $U$ matrix which minimizes the sum of squared errors.
+We then fix that $U$ matrix and choose an optimal $V$ matrix, iterating this process until convergence.
+This procedure is guaranteed to converge, though it may converge to a local minimum.
+Thus, we generally initialize the $U$ and $V$ matrices with random values, then run the algorithm multiple times to ensure we find something approximating the global minimum.
+
+% With our decomposition in hand, it is simple to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
+If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
+The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
+
+% \subsection{Validation}
+
+% Rather than blindly trusting our decomposition, we can validate it by comparing predicted reviews against actual reviews.
+I hold out 5\% of the reviews as a test set, then train on the remaining 95\%.
+I then predict the held-out reviews using the trained UV decomposition and compare the predicted reviews to the actual reviews.
+If the predicted reviews are highly correlated with the actual reviews, we can be more confident in the UV decomposition.
+
+% The relationship between predicted and actual reviews is visible in figure FIGURE.
+As you can see, there exists a strong positive correlation between the predicted reviews and the actual reviews.
+The strength of this correlation varies based on the number of reviews a movie has received in the data.
+Points are colored by the number of reviews their movie received; as is visible in the figure, movies with few reviews have a quite weak correlation between predicted and actual reviews.
+However, once we trim off the least popular 5\% of movies, the correlation between predicted and actual reviews becomes quite a bit stronger.
+This makes sense: our model is (as expected) better at predicting a movie's appeal when we have more data on that movie.
+
+% These results affirm for me that our collaborative filtering results are likely to be quite predictive of a movie's appeal.
+Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
+
+\subsection{Comparison}
+
+LOREM IPSUM
 

### 2024-09-05 draft/draft.tex
Fixed UV decomposition

--- 2024-09-02 draft/draft.tex
@@
 If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
+I validate my text-based model by comparing it against two collaborative filtering algorithms: topic-specific PageRank and UV decomposition.
+
+\subsection{Topic-Specific PageRank}
+
+Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies, as implemented by CITE FAST PAGERANK HERE.
+PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages.
+The algorithm works by randomly walking through the network, with a probability of jumping to a random node at each step.
+The importance of a node is the proportion of time the random walker spends at that node.
+TSPR is an extension of the algorithm which always ``jumps'' to a specific set of nodes rather than a random node.
+
+For my application, I implement TSPR as follows. To find movies similar to specific movie $x$,:
+\begin{itemize}
+    \item The algorithm randomly walks over the bipartite network of users and movies. 
+    \item At each step, the algorithm has a 15\% chance of jumping back to $x$
+    \item If the algorithm does not jump to $x$, it takes a random step to a neighbor of its current node
+    \item Random draws are weighted by the number of stars in the review; e.g., a ``5-star'' review is five times as likely to be drawn as a ``1-star'' review
+\end{itemize}
+
+Running this algorithm until convergence provides a measure of how central any given movie $y$ is relative to movie $x$.
+However, we are still not quite done, as some movies are just \emph{inherently} central; if everyone likes \emph{Forrest Gump,} then it's not particularly informative that fans of movie $x$ \emph{also} like \emph{Forrest Gump}.
+Thus, I divide the TSPR score of $y$ by $y$'s \emph{non}-topic specific PageRank score.
+This gives a measure of how much more central $y$ is to $x$ than it is to the average movie.
+Given this is a ratio, it has propensity to explode if $y$'s centrality is very small; thus, I prefer to take the log of this ratio for my analysis.
+
+Once I have each movie's TSPR score for each other movie, I can compare these scores to the cosine similarity of the text embeddings.
+Note that the TSPR scores are not symmetric; that is, the TSPR score of movie $x$ to movie $y$ is not necessarily the same as the TSPR score of movie $y$ to movie $x$.
+Thus, for each pair of movies, I actually plot \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
+
+A common weakness of collaborative filtering algorithms like TSPR is that they require significant data before they become effective.
+Thus, I compute TSPR values only for movies which receive at least 10,000 reviews in the MovieLens data.
+This leaves me with 258 movies, or 66,564 pairs of movies.
+To assess the importance of this margin, I then plot the coefficient of regressing TSPR on text similarity for thresholds 10,000 to 50,000.
+I also plot the number of movie pairs included in the sample for each threshold value, to assess the trade-off of ``more ratings per movie'' vs.
+``more movies.'' The results of this plot are visible in figure \ref{fig:tspr_threshold}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../movielens/output/threshold_vs_coefficient.png}
+    \caption{Coefficient of log(TSPR) on Text Similarity by Threshold}
+    \label{fig:tspr_threshold}
+    \end{center}
+\end{figure}
+
+This plot shows that the coefficient on text similarity for predicting TSPR is consistently and significantly positive for nearly all thresholds.
+Thus, the content of the text embeddings \emph{does} seem to have predictive power for which movies are similar based on revealed preference; that is, the patterns I see in the text-based model are not coincidental and reflect consumers' true preferences.
+
+Moreover, while the number of pairs included remains large enough to have useful confidence intervals, the coefficient is increasing in the threshold value.
+If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like our text-based cosine similarity measure.
+This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per movie to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.
 
@@
 
-For my collaborative filtering algorithm, I use UV-decomposition for matrix completion as described in Leskovec (CITE HERE).
+I implement UV-decomposition for matrix completion as described in Leskovec (CITE HERE).
 Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
@@
 The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
-
-\subsection{Validation}
 
@@
 
-\subsection{Comparison}
-
-LOREM IPSUM
+I find XXXXXX
 

### 2024-09-05 draft/draft.tex
Wrote up UV decomp results

@@
 
-MovieLens produces a public-use movie review dataset; it contains more than 25 million (CHECK THIS NUMBER) 1-to-5 star reviews of movies by users.
+MovieLens produces a public-use movie review dataset; it contains more than 25 million 1-to-5 star reviews of movies by users.
 Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions.
@@
 If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
-I validate my text-based model by comparing it against two collaborative filtering algorithms: topic-specific PageRank and UV decomposition.
+I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: topic-specific PageRank and UV decomposition.
+
+\subsection{UV Decomposition}
+
+I implement UV-decomposition for matrix completion as described in Leskovec (CITE HERE).\footnote{Note the algorithm described in the textbook optimizes element-wise, which is quite slow.
+Based on my own vector calculus, I've implemented \emph{matrix-wide} optimization, which seems to work.
+However, I have not found a reference for this approach anywhere; I emailed Professor Leskovec to confirm the result is the same, but have not heard back.} Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
+However, the matrix is generally sparse; most users have not reviewed most movies.
+UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix.
+This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
+By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.
+
+Note that UV decomposition relies on a contraction mapping: for a fixed $V$ matrix, we choose the $U$ matrix which minimizes the sum of squared errors.
+We then fix that $U$ matrix and choose an optimal $V$ matrix, iterating this process until convergence.
+This procedure is guaranteed to converge, though it may converge to a local minimum.
+Thus, we generally initialize the $U$ and $V$ matrices with random values, then run the algorithm multiple times to ensure we find something approximating the global minimum.
+
+With our decomposition in hand, it is straightforward to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
+If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
+The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
+
+Rather than blindly trusting our decomposition, we can validate it by comparing predicted reviews against actual reviews.
+I hold out 5\% of the reviews as a test set, then train on the remaining 95\%.
+If the predicted reviews for this held-out set are highly correlated with the actual reviews, we can be more confident in the UV decomposition.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../movielens/output/uv_decomp_out_of_sample.png}
+    \caption{Actual vs. Predicted Reviews by Number of Reviews}
+    \label{fig:actual_vs_predicted}
+    \end{center}
+\end{figure}
+
+As we might expect, the UV decomposition is better at predicting reviews for movies with more reviews.
+To assess the size of this effect, I regress the true review on the predicted review for various thresholds of review count.
+If the threshold is low, we're including movies with fewer reviews in our prediction set; if the threshold is high, we're looking only at hits like \emph{Star Wars.} The results of these regressions are visible in figure \ref{fig:uv_threshold}.
+
+For any value of the review count threshold, the coefficient on predicted reviews is positive and significant.
+This suggests that the UV decomposition is quite predictive of a movie's appeal.
+Moreover, the coefficient is increasing in the review count threshold; that is, the UV decomposition is better at predicting reviews for movies with more reviews.
+This is as expected; the more data we have on a movie, the better we can predict its appeal.
+
+These results affirm for me that our collaborative filtering results are likely to be quite predictive of a movie's appeal.
+Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
+
+Assured that the UV decomposition results are valid, I attempt to predict the cosine similarity of the UV decomposition's predicted reviews using the cosine similarity of the text embeddings.
+Again we expect the quality of these predictions to vary with the volume of the data; thus, I plot the coefficient of this regression for various thresholds of review count as above.
+The results of these regressions are visible in figure \ref{fig:uv_threshold}.
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=\textwidth]{../movielens/output/UV_decomp_coef_vs_review_count.png}
+    \caption{Coefficient of UV Decomposition on Text Similarity by Threshold}
+    \label{fig:uv_threshold}
+    \end{center}
+\end{figure}
+
+For movies with relatively few reviews, the coefficient on text similarity is insignificant; this is to be expected, as for these movies UV decomposition struggles to capture anything of meaning.
+However, as the review count threshold increases, the coefficient on text similarity becomes increasingly positive and significant.
+This suggests that as UV decomposition becomes more accurate, it looks increasingly like our text-based cosine similarity measure.
+This is a strong validation of my text-based model; the patterns I see in the text embeddings are not coincidental, but reflect consumers' true preferences.
+The quality of the match begins to decline when the review threshold becomes too high; however, this is not a fault with my method but with the decomposition, as at these high standards the number of movies with sufficient views falls to less than 10, producing fewer than 100 pairs of movies.
 
@@
 
-\subsection{UV Decomposition}
-
-I implement UV-decomposition for matrix completion as described in Leskovec (CITE HERE).
-Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
-However, the matrix is generally sparse; most users have not reviewed most movies.
-UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix.
-This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
-By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.
-
-Note that UV decomposition relies on a contraction mapping: for a fixed $V$ matrix, we choose the $U$ matrix which minimizes the sum of squared errors.
-We then fix that $U$ matrix and choose an optimal $V$ matrix, iterating this process until convergence.
-This procedure is guaranteed to converge, though it may converge to a local minimum.
-Thus, we generally initialize the $U$ and $V$ matrices with random values, then run the algorithm multiple times to ensure we find something approximating the global minimum.
-
-With our decomposition in hand, it is simple to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
-If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
-The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
-
-Rather than blindly trusting our decomposition, we can validate it by comparing predicted reviews against actual reviews.
-I hold out 5\% of the reviews as a test set, then train on the remaining 95\%.
-I then predict the held-out reviews using the trained UV decomposition and compare the predicted reviews to the actual reviews.
-If the predicted reviews are highly correlated with the actual reviews, we can be more confident in the UV decomposition.
-
-The relationship between predicted and actual reviews is visible in figure FIGURE.
-As you can see, there exists a strong positive correlation between the predicted reviews and the actual reviews.
-The strength of this correlation varies based on the number of reviews a movie has received in the data.
-Points are colored by the number of reviews their movie received; as is visible in the figure, movies with few reviews have a quite weak correlation between predicted and actual reviews.
-However, once we trim off the least popular 5\% of movies, the correlation between predicted and actual reviews becomes quite a bit stronger.
-This makes sense: our model is (as expected) better at predicting a movie's appeal when we have more data on that movie.
-
-These results affirm for me that our collaborative filtering results are likely to be quite predictive of a movie's appeal.
-Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
-
-I find XXXXXX
-
 \section{Conclusion}

### 2024-09-10 elasticity_estimate/code/estimate_elasticity.py
Broke apart two types of age coefficients

@@ -1,6 +1,7 @@
 import pandas as pd
 import numpy as np
 from scipy.optimize import minimize
+from optimparallel import minimize_parallel
 import matplotlib.pyplot as plt
 import linearmodels as lm
 from stargazer.stargazer import Stargazer
@@ -62,6 +63,9 @@ plt.xticks(ticks=range(len(genre_counts)), rotation=45)
 
 plt.savefig('output/genre_counts.png')
 
+# Clear plot
+plt.clf()
+
 # Get array of movie ids
 movie_ids = tmdb['movie_id'].copy()
 
@@ -82,16 +86,11 @@ guru = pd.read_parquet('../guru/data/weekly_sales.parquet')
 
 guru = guru.reset_index(drop=False)
 
-print(guru.dtypes)
-
 # Merge on movie genre
 guru = guru.merge(tmdb[['movie_id', 'tmdb_genre']], on='movie_id', how='left')
 
 # Create dummy for whether another movie of that genre is showing within each week
-guru['genre_clash'] = guru.groupby(['Date', 'tmdb_genre'])['movie_id'].transform('count') > 1
-
-# Get summary of genre_clash
-print(guru['genre_clash'].value_counts())
+# guru['genre_clash'] = guru.groupby(['Date', 'tmdb_genre'])['movie_id'].transform('count') > 1
 
 # Compute log interest
 guru['ln_earnings'] = np.log(guru['Week Sales'])
@@ -99,11 +98,13 @@ guru['ln_earnings'] = np.log(guru['Week Sales'])
 # Drop if missing weeks
 guru = guru.dropna(subset=['Weeks'])
 
-# Histogram of weeks since release
+print(guru['Weeks'].describe())
+
+# Create histogram of weeks since release
 plt.hist(guru['Weeks'], bins=20)
-plt.xlabel('Weeks Since Wide Release')
-plt.ylabel('Frequency')
-plt.title('Histogram of Weeks Since Wide Release')
+plt.xlabel('Weeks Since Release')
+plt.ylabel('Count')
+plt.title('Histogram of Weeks Since Release')
 plt.savefig('output/weeks_histogram.png')
 
 # We want their weeks to also be zero-indexed so we can use it as an index
@@ -230,14 +231,14 @@ def regress_given_lambda(age_coefficients, guru):
         competitor_age_coefs = age_coefficients[date_movies['Weeks']]
[…]

### 2024-09-10 elasticity_estimate/code/estimate_elasticity.py
Further attempts at nonlinear estimation

@@ -16,196 +16,8 @@ warnings.simplefilter(action='ignore', category=FutureWarning)
 # Note zero-indexing; thus, this is actually the *fifth* week
 WEEK_THRESHOLD = 4
 
-# Load the embeddings from parquet
-tmdb = pd.read_parquet('../tmdb/embeddings/movie_descriptions.parquet')
-
-# Drop if embeddings are null
-tmdb = tmdb.dropna(subset=['embedding'])
-tmdb = tmdb.reset_index(drop=True)
-
-# Filter to primary genre
-tmdb['tmdb_genre'] = tmdb['tmdb_genres'].apply(lambda x: x[0] if len(x) > 0 else None)
-
-genre_labels = {
-    28: 'Action',
-    12: 'Adventure',
-    16: 'Animation',
-    35: 'Comedy',
-    80: 'Crime',
-    99: 'Documentary',
-    18: 'Drama',
-    10751: 'Family',
-    14: 'Fantasy',
-    36: 'History',
-    27: 'Horror',
-    10402: 'Music',
-    9648: 'Mystery',
-    10749: 'Romance',
-    878: 'Science Fiction',
-    10770: 'TV Movie',
-    53: 'Thriller',
-    10752: 'War',
-    37: 'Western'
-}
-
-# Replace genre with label
-tmdb['tmdb_genre'] = tmdb['tmdb_genre'].map(genre_labels)
-
-# Create bar chart of genres
-genre_counts = tmdb['tmdb_genre'].value_counts()
-genre_counts.plot(kind='bar')
-plt.xlabel('Genre')
-plt.ylabel('Count')
-plt.title('Count of Movies by Genre')
-
-# Label the genres
-plt.xticks(ticks=range(len(genre_counts)), rotation=45)
-
-plt.savefig('output/genre_counts.png')
-
-# Clear plot
-plt.clf()
-
-# Get array of movie ids
-movie_ids = tmdb['movie_id'].copy()
-
-# Create a map from movie_ids to indices
-movie_id_to_index = {movie_id: i for i, movie_id in enumerate(movie_ids)}
-
-raw_embeddings = np.array(tmdb['embedding'].values)
-raw_embeddings = np.array([np.array(x) for x in raw_embeddings])
-
-# Create distance matrix by taking dot product of embeddings
-distances = 1 - raw_embeddings @ raw_embeddings.T
-
-# Normalize distances
-distances = distances / np.max(distances)
-
[…]

### 2024-09-10 draft/draft.tex
Add citations for collaborative filtering section

@@
 
-MovieLens produces a public-use movie review dataset; it contains more than 25 million 1-to-5 star reviews of movies by users.
+\textcite{harper2016ATIIS} provide a public-use dataset of real movie reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of movies by users.
 Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions.
@@
 
-I implement UV-decomposition for matrix completion as described in Leskovec (CITE HERE).\footnote{Note the algorithm described in the textbook optimizes element-wise, which is quite slow.
+I implement UV-decomposition for matrix completion as described in \textcite{leskovec2020MMD}.\footnote{Note the algorithm described in the textbook optimizes element-wise, which is quite slow.
 Based on my own vector calculus, I've implemented \emph{matrix-wide} optimization, which seems to work.
@@
 
-Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies, as implemented by CITE FAST PAGERANK HERE.
+Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies, as implemented by \textcite{sajani2023}.
 PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages.

### 2024-09-16 elasticity_estimate/code/estimate_elasticity.py
Implemented old model

@@ -16,6 +16,13 @@ warnings.simplefilter(action='ignore', category=FutureWarning)
 # Note zero-indexing; thus, this is actually the *fifth* week
 WEEK_THRESHOLD = 4
 
+# Set equal to 1 if normalizing the first coefficient to 1
+PREFIX = []
+INITIAL_GUESS = [1, .8, .6, .4, .2]
+
+# Regression specification
+REG_FORMULA = 'ln_earnings_adjusted ~ 1 + gamma0 + gamma1 + gamma2 + gamma3 + EntityEffects + TimeEffects'
+
 def regress_given_lambda(age_coefficients, guru, movie_id_to_index, date_movie_dict, distances):
     """Function for nonlinear optimization of age coefficients"""
 
@@ -45,12 +52,12 @@ def regress_given_lambda(age_coefficients, guru, movie_id_to_index, date_movie_d
         competitor_distances = distances[movie_index, competitor_indices]
 
         # No need to weight by log price since log price is constant
-        # If using weighted distances, subtract off one for self
         # Don't want to raise age coefficient to power when computing distances
-        gamma0.append(len(competitor_distances) - 1)
-        gamma1.append(np.sum(competitor_age_coefs * competitor_distances))
-        gamma2.append(np.sum(competitor_age_coefs * (competitor_distances ** 2)))
-        gamma3.append(np.sum(competitor_age_coefs * (competitor_distances ** 3)))
+        # Need to subtract off the own-age coefficient
+        gamma0.append(np.sum((1 / competitor_age_coefs) * np.ones(len(competitor_distances))) - (1 / age_coefficients[row.Weeks]))
+        gamma1.append(np.sum((1 / competitor_age_coefs) * competitor_distances))
+        gamma2.append(np.sum((1 / competitor_age_coefs) * (competitor_distances ** 2)))
+        gamma3.append(np.sum((1 / competitor_age_coefs) * (competitor_distances ** 3)))
 
     # Add to the guru data
     guru['gamma0'] = gamma0
@@ -58,25 +65,37 @@ def regress_given_lambda(age_coefficients, guru, movie_id_to_index, date_movie_d
     guru['gamma2'] = gamma2
     guru['gamma3'] = gamma3
 
+    # Convert date to month and year
+    guru['month'] = guru['Date'].dt.month
[…]

### 2024-09-16 draft/draft.tex
Text description of new model

@@
 \usepackage{booktabs}
+\usepackage{amsmath}
 
@@
 
+\section{Revised Model: Unified Lambda and Fixed Effect}
+
+Per conversations in the spring (and with Liran this week), I want to include age in the cross-elasticity term as well.
+That is, if I'm releasing Star Trek, having Star Wars as my competitor is more important if Star Wars released recently than if it is old.
+Moreover, we estimate fixed effects for each movie; these fixed effects are likely important for competitors as well, not just for estimating own demand.
+Thus, I should include the $\alpha_i$ fixed effect in the cross-elasticity term as well.
+
+To do this, I've revamped my demand model as follows.
+Suppose we wish to estimate revenue for movie $i$ in time $t$.
+This is a function $\phi$ of the movie's quality $\delta_{it}$ and its competitors' qualities $\delta_{-it}$, as well as a fixed effect for the weekend $\alpha_t$ (e.g., many people go see movies on Memorial Day).
+This gives the following equation:
+
+$$\ln(q_{it}) = \phi(\delta_{it}, \delta_{-it}) + \alpha_t$$
+
+A movie's quality is given by a combination of a fixed effect $\alpha_i$, an age fixed effect $\lambda_{t - r(i)}$, and an idiosyncratic shock $\xi_{it}$:
+
+$$\delta_{it} = \alpha_i + \lambda_{t - r(i)} + \xi_{it}$$
+
+Thus, we can rewrite our specification above as:
+
+$$\ln(q_{it}) = \phi(\alpha_i + \lambda_{t - r(i)} + \xi_{it}, \alpha_{-i} + \lambda_{t - r(-i)} + \xi_{-it}) + \alpha_t$$
+
+We'll let the influence of competitors on movie $i$ be a function of their quality interacted with a flexible function of distance $f$.
+Intuitively $f$ should be negative and decrease in magnitude with distance (such that competition reduces quantity sold and more distant movies are less influential), but I don't impose this.
+The rest of our specification will be linear.
+Thus, we can rewrite our demand equation as:
+
+$$\ln(q_{it}) = \alpha_0 + \underbrace{\alpha_i + \lambda_{t - r(i)} + \xi_{it}}_{\delta_{it}} + \sum_{j \neq i} f(d_{ij}) \cdot \underbrace{(\alpha_j + \lambda_{t - r(j)} + \xi_{jt})}_{\delta_{jt}} + \alpha_t$$
+
+As it is written, this model is very difficult to estimate for multiple reasons.
+$\alpha_i$ appears not just in its own demand, but in competitors' as well; thus, we cannot use simple demeaning to absorb these fixed effects.
+Moreover, the $f$ function interacts multiplicatively with the $\delta_{jt}$ terms, making matrix inversion by itself useless.
+With some clever rearranging, however, I believe this model becomes more tractable:
+
+\begin{align*}
+    \ln(q_{it}) = \alpha_0 &+ \alpha_i + \sum_{j \neq i} f(d_{ij}) \alpha_j \\
+    &+ \lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(j)} \left(\sum_{j}^{r(j) = r(k)} f(d_{ij}) \right)\right) \\
+    &+ \alpha_t \\
+    &+ \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt}
+\end{align*}
+
+Thus, conditional on the values of $f(d_{ij})$, each movie's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend.
+Thus, the $\lambda$ and $\alpha$ coefficients are estimable using OLS.
+
+As an example, consider what our matrix will look like in the following market.
+Movie 1 released this week.
+Movie 2 released this week as well, while movie 3 released last week.
+I omit time fixed effects in this example since we focus on a single period; similarly, I omit time subscripts on the quantities.
+In this case, we will have the following data matrix where values in ``movie'' columns are used to estimate $\alpha$ values while ``age'' columns are used to estimate $\lambda$ values:
+
+\begin{center}
+
+\begin{tabular}{cccccc}
+    ln(Quantity) & Movie 1 & Movie 2 & Movie 3 & Age 0 & Age 1 \\
+    $\ln(q_{1})$ & 1 & $f(d_{12})$ & $f(d_{13})$ & $1 + f(d_{12})$ & $f(d_{13})$ \\
+    $\ln(q_{2})$ & $f(d_{12})$ & 1 & $f(d_{23})$ & $1 + f(d_{12})$ & $f(d_{23})$ \\
+    $\ln(q_{3})$ & $f(d_{13})$ & $f(d_{23})$ & 1 & $f(d_{13}) + f(d_{23})$ & $1$ 
+\end{tabular}
+
+\end{center}
+
+Regressing quantity on the given Xs identifies the coefficients for $\alpha_1$, $\alpha_2$, $\alpha_3$, $\lambda_0$, and $\lambda_1$ via the first through fifth columns, respectively.
+These results make some amount of intuitive sense; if $\ln(q_1)$ performs below expectations, one explanation is that $\alpha_2$ is large--represented by a large coefficient on the (negative) $f(d_{12})$.
+
+Note the above is all taking $f$ as given. Let $f$ be a cubic polynomial in distance; that is:
+
+$$f(d_{ij}) = \gamma_0 + \gamma_1 d_{ij} + \gamma_2 d_{ij}^2 + \gamma_3 d_{ij}^3$$
+
+This follows Magnolfi's specification, and allows for flexible sign, monotonicity, and concavity.
+Thus, if the final result \emph{does} exhibit those properties, we know it is coming from the data rather than baked into the model.
+
+Since the $\lambda$ and $\alpha$ coefficients in the model above interact multiplicatively with the $\gamma$ coefficients, we cannot estimate $\gamma$ linearly at the same time that we estimate $\lambda$ and $\alpha$ linearly.
+Instead, I will choose a $\gamma$ coefficient vector and estimate $\lambda$ and $\alpha$ given that $\gamma$ vector using OLS.
+We know the resulting model minimizes mean squared error \emph{in the set of models using this $\gamma$}.
+If I then optimize the $\gamma$ vector to minimize mean squared error of these optimized models, I search across the lower envelope and know my final model minimizes mean squared error over the whole $\gamma, \alpha, \lambda$ space.
+
 \section{Conclusion}

### 2024-09-16 elasticity_estimate/code/estimate_elasticity.py
Code mostly function s (but slowly)

@@ -20,85 +20,90 @@ WEEK_THRESHOLD = 4
 PREFIX = []
 INITIAL_GUESS = [1, .8, .6, .4, .2]
 
-# Regression specification
-REG_FORMULA = 'ln_earnings_adjusted ~ 1 + gamma0 + gamma1 + gamma2 + gamma3 + EntityEffects + TimeEffects'
-
-def regress_given_lambda(age_coefficients, guru, movie_id_to_index, date_movie_dict, distances):
+def regress_given_gamma(gamma, guru, movie_id_to_index, date_movie_dict, distances):
     """Function for nonlinear optimization of age coefficients"""
 
     guru = guru.copy()
 
-    # For each movie-date, compute the leave-one-out set of distances within the date
-    gamma0 = []
-    gamma1 = []
-    gamma2 = []
-    gamma3 = []
+    # Define cubic dependant on gamma
+    cubic = lambda x: gamma[0] + gamma[1] * x + gamma[2] * x ** 2 + gamma[3] * x ** 3
+
+    rows = []
+
+    print("Constructing Data...")
 
     for row in guru.itertuples():
-
-        # Can I avoid redundant computation within a date?
+        # Get the movie's date and id
         movie_id = row.movie_id
         date = row.Date
+        age = row.Weeks
+        ln_earnings = row.ln_earnings
 
         # Get index of movie_id in movie_ids
         movie_index = movie_id_to_index[movie_id]
 
+        # Create a lambda vector with length equal to the number of weeks
+        lambda_vals = np.zeros(WEEK_THRESHOLD)
+        lambda_vals[age] = 1
+
+        # Create an alpha vector with length equal to the number of movies
+        alpha_vals = np.zeros(len(movie_id_to_index))
+        alpha_vals[movie_index] = 1
+
         # Get all movies on the date
         date_movies = date_movie_dict[date]
 
-        # Can ignore that self is included since distance(x, x) = 0
-        competitor_age_coefs = age_coefficients[date_movies['Weeks']]
-        competitor_indices = np.array([movie_id_to_index[movie_id] for movie_id in date_movies['movie_id']])
-        competitor_distances = distances[movie_index, competitor_indices]
+        for competitor in date_movies.itertuples():
+            competitor_id = competitor.movie_id
+            competitor_age = competitor.Weeks
[…]

### 2024-09-16 elasticity_estimate/code/estimate_elasticity.py
Model estimation seems to be working

@@ -14,11 +14,66 @@ warnings.simplefilter(action='ignore', category=FutureWarning)
 
 # For weeks greater than this, we truncate to this
 # Note zero-indexing; thus, this is actually the *fifth* week
-WEEK_THRESHOLD = 4
+WEEK_THRESHOLD = 9
 
-# Set equal to 1 if normalizing the first coefficient to 1
-PREFIX = []
-INITIAL_GUESS = [1, .8, .6, .4, .2]
+# Initial guess for gamma
+INITIAL_GUESS = np.array([1, .1, .01, .001])
+
+
+def get_neighbors(example, distances, tmdb, movie_ids):
+    '''
+    Function to get characteristic examples for a given movie
+    '''
+    # Get index of the example
+    example_index = np.where(movie_ids == example)[0][0]
+
+    example_name = tmdb[tmdb['movie_id'] == example]['Clean Title'].values[0]
+
+    distances_to_example = distances[example_index, :].copy()
+
+    print("\n")
+
+    print(f'Closest movies to {example_name}:')
+    print("-------------------------")
+    for i in range(10):
+        closest_index = np.argmin(distances_to_example)
+        closest_movie = movie_ids[closest_index]
+        closest_distance = distances_to_example[closest_index]
+
+        closest_movie_name = tmdb['Clean Title'][closest_index]
+
+        print(f'{closest_movie_name}: {closest_distance}')
+        distances_to_example[closest_index] = np.inf
+    
+    print("\n")
+
+    distances_to_example = distances[example_index].copy()
+    print("Movies which are close:")
+    distances_from_bottom_decile = np.abs(distances_to_example - bottom_end)
+    for i in range(10):
+        closest_index = np.argmin(distances_from_bottom_decile)
+        closest_movie = movie_ids[closest_index]
+        closest_distance = distances_to_example[closest_index]
+
+        closest_movie_name = tmdb[tmdb['movie_id'] == closest_movie]['Clean Title'].values[0]
+
+        print(f'{closest_movie_name}: {closest_distance}')
+        distances_from_bottom_decile[closest_index] = np.inf
+
+    print("\n")
+
+    distances_to_example = distances[example_index].copy()
+    print("Movies which are far:")
+    distances_from_top_decile = np.abs(distances_to_example - top_end)
+    for i in range(10):
+        closest_index = np.argmin(distances_from_top_decile)
+        closest_movie = movie_ids[closest_index]
[…]

### 2024-09-16 elasticity_estimate/code/estimate_elasticity.py
Updated graphs

@@ -13,7 +13,6 @@ import pstats
 warnings.simplefilter(action='ignore', category=FutureWarning)
 
 # For weeks greater than this, we truncate to this
-# Note zero-indexing; thus, this is actually the *fifth* week
 WEEK_THRESHOLD = 9
 
 # Initial guess for gamma
@@ -356,7 +355,7 @@ def main():
 
     optimal_model = regress_given_gamma(gamma_star, guru, movie_id_to_index, date_movie_dict, distances)
 
-    print(optimal_model)
+    print(gamma_star)
 
     '''
     stargazer = Stargazer([optimal_model])
@@ -379,10 +378,6 @@ def main():
     bottom_index = np.argmin(np.abs(distance_grid - bottom_end))
     top_index = np.argmin(np.abs(distance_grid - top_end))
 
-
-    print(f'Cross-elasticity at bottom end of distance: {cross_elasticity[bottom_index]}')
-    print(f'Cross-elasticity at top end of distance: {cross_elasticity[top_index]}')
-
     # Flatten the distances
     distances = distances.flatten()
 
@@ -390,7 +385,7 @@ def main():
     fig, ax1 = plt.subplots()
 
     # Add title
-    ax1.set_title('Cross-Elasticity of Distance')
+    ax1.set_title('Impact over Distance')
     ax2 = ax1.twinx()
 
     # Plot the empirical distances
@@ -399,13 +394,13 @@ def main():
     ax2.set_ylabel('Density of Empirical Distance')
 
     ax1.plot(distance_grid, cross_elasticity, color='orange')
-    ax1.set_ylabel('Cross-Elasticity of Distance')
+    ax1.set_ylabel('Competition Function')
 
     ax2.axvline(x=top_end, color='red', linestyle='--')
     ax2.axvline(x=bottom_end, color='red', linestyle='--')
 
     fig.tight_layout()
-    plt.savefig('output/cross_elasticity.png')
+    plt.savefig('output/gamma.png')
 
     # Zoom in on the cross-elasticity in the support of the data
 
@@ -430,10 +425,10 @@ def main():
     ax1.axhline(y=0, color='grey', linestyle='--')
 
     # Add title
-    ax1.set_title('Cross-Elasticity of Distance')
+    ax1.set_title('Impact over Distance')
 
     ax1.plot(distance_grid, cross_elasticity, color='orange')
-    ax1.set_ylabel('Cross-Elasticity of Distance')
[…]

### 2024-09-16 elasticity_estimate/code/estimate_elasticity.py
Add information about the average alpha

@@ -125,7 +125,7 @@ def regress_given_gamma(gamma, guru, movie_id_to_index, date_movie_dict, distanc
     regression_df = pd.DataFrame(rows)
 
     # Rename columns
-    regression_df.columns = [f'alpha_{i}' for i in range(len(movie_id_to_index))] + [f'lambda_{i}' for i in range(WEEK_THRESHOLD)] + ['date', 'ln_earnings', 'movie_id']
+    regression_df.columns = [f'alpha_{index}' for index in range(len(movie_id_to_index))] + [f'lambda_{i}' for i in range(WEEK_THRESHOLD)] + ['date', 'ln_earnings', 'movie_id']
 
     # Set the index to the date and movie_id
     regression_df['date'] = pd.to_datetime(regression_df['date'])
@@ -331,7 +331,7 @@ def main():
     # The minimiziation process is expensive; thus, 
     # we get "close" to the optimum with a small sample, then
     # use that as the initial guess for the full sample
-
+    '''
     initial_minimization = minimize(
         fun = compute_model_error, 
         x0 = INITIAL_GUESS,
@@ -340,7 +340,8 @@ def main():
 
     # Get the optimal age coefficients
     gamma_star = initial_minimization.x
-    '''
+
+
 
     final_minimization = minimize(
         fun = compute_model_error,
@@ -350,6 +351,7 @@ def main():
 
     gamma_star = final_minimization.x
     '''
+    gamma_star = [0.332, -1.903, 3.298, -1.796]
 
     print("Initial Minimiziation Complete!")
 
@@ -444,20 +446,6 @@ def main():
     fig.tight_layout()
     plt.savefig('output/gamma_zoomed.png')
 
-
-    # Get the alpha coefficients
-    alpha_cols = [colname for colname in optimal_model.params.index if 'alpha' in colname]
-    alpha_values = optimal_model.params[alpha_cols]
-
-    # Plot a distribution of the alpha coefficients
-    plt.clf()
-    plt.hist(alpha_values, bins=20)
-    plt.xlabel('Alpha Coefficient')
-    plt.ylabel('Count')
-    plt.title('Distribution of Alpha Coefficients')
-    plt.savefig('output/alpha_coefficients.png')
-
-
     # Get the age coefficients and errors
     lambda_cols = [colname for colname in optimal_model.params.index if 'lambda' in colname]
     lambda_values = optimal_model.params[lambda_cols]
[…]

### 2024-09-17 draft/draft.tex
Updated draft with new model

--- 2024-09-10 draft/draft.tex
@@
 
-\subsection{Log-Linear Demand}
+\subsection{Construction}
 
-Following MMS, I estimate a log-linear demand model for movies.
-I start by including fixed effects for each movie and weekend.
-The distance between products $i$ and $j$ is given by some metric $d_{ij}$; a competitor's impact is given by via a flexible function of distance $f(d_{ij})$.
-The sum of individual competitors' impacts yields the total impact of competition on a given movie.
-Again following MMS, I estimate $f$ as a cubic polynomial in distance.
+Movie consumers face a discrete choice problem: which of the movies on offer this weekend should they see?
+Let a movie $i$'s attractiveness at time $t$ be given by some $\delta_{it}$.
+Consumers receive idiosyncratic demand shocks for each movie, and choose the movie which maximizes their utility.
+Due to the uniform pricing scheme discussed above, price \emph{does not} enter into the consumer's decision; choices are driven entirely by movie quality and idiosyncratic taste.
 
-The first benefit of the cubic polynomial functional form here is that this assumption is falsifiable; if $f$ is not monotonically attenuating, our spatial competition model is likely misspecified.
-The second benefit of this functional form is it is estimable linearly; by distributing the grand sum over each individual term and factoring out the $\gamma$ coefficients, demand becomes a simple linear equation in $d_{ij}$, $d_{ij}^2$, and $d_{ij}^3$.
-Thus, traditional matrix inversion techniques remain effective.
-The final benefit is that the cubic polynomial is flexible; it imposes no structure on sign, derivative, or magnitude.
-Thus, I can capture potentially quite rich substitution patterns, as well as evaluate my model fit against ``reasonable'' priors.
+Aggregating over all consumers, we can express the demand for movie $i$ in week $t$ as a function $\phi$ of the movie's appeal $\delta_{it}$ and the appeal of its competitors $\delta_{-it}$.
+Again due to the uniform pricing scheme, we can use revenues as a proxy for quantity, since quantity is simply a constant scalar transformation of revenue.
+Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Memorial Day than on a random weekend in July.
+Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation.
+Thus, I can write the demand for movie $i$ in week $t$ as:
 
-Unlike cereals, demand for movies is heavily based on \emph{novelty}; that is, a movie will generally sell more the week of release than two months later.
-Thus, I include a fixed effect $\lambda_{t - r(j)}$ to control for time since release, where $r(j)$ is the release date of movie $j$.
+$$\ln(q_{it}) = \phi(\delta_{it}, \delta_{-it}) + \alpha_t$$
 
-The final model of demand for movie $j$ in time $t$ is as follows:
+A movie's appeal is composed of two parts.
+First, a fixed effect $\alpha_i$ captures the movie's quality and characteristics.
+For example, we would expect \emph{Star Wars} to have a high $\alpha_i$ to capture the fact it is a popular movie in a vacuum separate from its competitors.
+Second, movies have an age fixed effect $\lambda_{t - r(i)}$.
+As an experience good, demand for a movie is driven heavily by novelty; thus, the same movie will experience much less demand in its second week than in its first.
+While we intuitively expect these $\lambdas$ to decrease with time, I do not impose this in the model; this will be a test of model fit later on.
+Finally, I include an idiosyncratic shock $\xi_{it}$ to capture unobserved demand drivers.
+Thus, the movie's appeal is given by:
 
-$$\ln(q_{jt}) = \alpha_j + \alpha_t + \sum_{k \neq j} f(d_{jk}; \gamma) + \lambda_{t - r(j)} + \varepsilon_{jt}$$
+$$\delta_{it} = \alpha_i + \lambda_{t - r(i)} + \xi_{it}$$
 
-$$f(d_{jk}; \gamma) = \gamma_0 + \gamma_1 d_{jk} + \gamma_2 d_{jk}^2 + \gamma_3 d_{jk^k}$$
+Putting these terms into our demand equation, we have:
 
-In the above specification, I purposely do not include price on the right hand side due to the uniform pricing assumption discussed above.
-We can interpret $f(d_{jk}; \gamma)$ as the cross-elasticity between movies $j$ and $k$; as movie $k$ becomes closer or further from movie $j$, $f$ maps how demand for movie $j$ changes.
-Thus, we generally expect to see $f$ monotonically attenuating in distance; as two movies become more distant, they should become increasingly insulated from each other.
+$$\ln(q_{it}) = \phi(\alpha_i + \lambda_{t - r(i)} + \xi_{it}, \alpha_{-i} + \lambda_{t - r(-i)} + \xi_{-it}) + \alpha_t$$
 
-\subsection{Age Interaction and Estimation}
+How should the demand for movie $i$ depend on its competitors? I put forward three competitor traits which should impact demand for movie $i$:
+\begin{itemize}
+    \item \emph{Competitor Quality}: If competitor $j$ has a high $\alpha_j$, it should draw more demand away from movie $i$.
+For example, \emph{The Lord of the Rings} is broadly a ``better'' movie than \emph{Willow,} despite being similar on paper; thus, $i$ would prefer to compete against \emph{Willow} rather than \emph{The Lord of the Rings}.
+    \item \emph{Competitor Age}: If competitor $j$ is more recent, it should draw more demand away from movie $i$.
+That is, $i$ should broadly prefer to release against movies in their fifth week of showing than in their first week.
+    \item \emph{Competitor Similarity}: If competitor $j$ is more similar to movie $i$, it should draw away more demand.
+For example, if $i$ is \emph{Star Wars}, it is much more enthusiastic to release against \emph{The Notebook} than against \emph{Star Trek}.
+\end{itemize}
 
-The above model leaves something to be desired if we think \emph{new movies are more relevant for competitors as well}.
-For example, my Science Fiction movie is much more unhappy to release the same weekend as \emph{Star Wars} than to release two months after \emph{Star Wars}.
-To capture this, I need to account for the age of the competitors as well by scaling $f$ by some factor $\delta_{t - r(k)}$.
-To preserve power, I should topcode $\delta$ at some reasonable value based on what I see in the data; e.g., I may group movies with 10+ weeks.
-I can inform my topcoding threshold by looking for plateaus in the empirical distribution of sales as movies age.
+Let $d_{ij}$ be a measure of similarity between movies $i$ and $j$.
+In this exercise I use the cosine distance of the embedded text descriptions; however, this could be any measure of similarity.
+There is no reason to expect similarity to enter linearly; instead, I consider some flexible function of distance $f$.
 
-In this case, our new model is:
+To capture the three attributes above, I propose the following functional form:
 
-$$\ln(q_{jt}) = \alpha_i + \alpha_t + \sum_{k \neq j} \delta_{t - r(k)} f(d_{jk}; \gamma) + \lambda_{t - r(j)} + \varepsilon_{jt}$$
+$$\ln(q_{it}) = \underbrace{\alpha_i + \lambda_{t - r(i)} + \xi_{it}}_{\delta_{it}} + \sum_{j \neq i} f(d_{ij}) \cdot (\underbrace{\alpha_j + \lambda_{t - r(j)} + \xi_{jt}}_{\delta_{jt}}) + \alpha_t$$
 
-Note this model is no longer linearly estimable, as $\delta_{t - r(k)}$ interacts multiplicatively with the $\gamma$ coefficients in $f$.
-I could manually pull out the fixed effects, then optimize over the possible values of $\delta$, $\lambda$, and $\gamma$, however I believe I have a more clever solution.
+Thus, demand for $i$ is a linear function of its own appeal $\delta_{it}$, a weekend fixed effect $\alpha_t$, and its competitors' appeals $\delta_{jt}$ weighted by some function $f$ of distance $d_{ij}$.
 
-Note for any fixed $\delta$ vector, we know the model is linear in $\gamma$ and $\lambda$, which makes it easy to identify the least-squares solution using matrix inversion.
-Thus, I can perform a two-step procedure to identify the least-squares coefficients of my nonlinear model.
+\subsection{Estimation}
 
-First, I choose a $\delta$.
-Then, using that $\delta$, I compute the least squares solution for $\gamma$ and $\lambda$.
-This gives me a sum of squared errors for this set of coefficients.
-I then use this sum of squared errors as the minimization objective to optimize over \emph{$\delta$ alone}.
-Since OLS always minimizes least squares \emph{within} a $\delta$ estimate, the optimal model must be the one which minimizes this quantity \emph{across} $\delta$s.
-The psuedocode for this algorithm is something like:
+As written, the model above is difficult to estimate.
+$\alpha_i$ appears not only in its own demand, but in competitors' as well; thus, we cannot use simple demeaning to absorb these fixed effects.
+Moreover, the $f$ function interacts multiplicatively with the $\delta_{jt}$ terms, making matrix inversion ineffective.
+With some clever rearranging, however, this model becomes more tractable:
 
-\begin{verbatim}
-def fit_model(delta, data):
-    for t in data:
-        weekend = data[t]
-        for j in weekend[movies]:
-            gamma_0_val = 0
-            gamma_1_val = 0
-            gamma_2_val = 0
-            gamma_3_val = 0
-            for k != j in weekend[movies]:
-                gamma_0_val += delta[movies[k][age]]
-                gamma_1_val += delta[movies[k][age]] * distance(j, k)
-                gamma_2_val += delta[movies[k][age]] * distance(j, k)^2
-                gamma_3_val += delta[movies[k][age]] * distance(j, k)^3
+\begin{align*}
+    \ln(q_{it}) = \alpha_i &+ \sum_{j \neq i} f(d_{ij}) \alpha_j \\
+    &+ \lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(j)} \left(\sum_{j}^{r(j) = r(k)} f(d_{ij}) \right)\right) \\
+    &+ \alpha_t \\
+    &+ \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt}
+\end{align*}
 
-    model = Panel(
-        y = ln(data[revenue]),
-        x = [
-            gamma_0_val,
-            gamma_1_val,
-            gamma_2_val,
-            gamma_3_val,
-            lambda[data[age]],
-        ],
-        fe = [data[movie], data[weekend]]
-    )
+Thus, conditional on the values of $f(d_{ij})$, each movie's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend.
+Thus, the $\lambda$ and $\alpha$ coefficients are estimable using OLS.
 
-    return model
+As an example, consider what our matrix will look like in the following market.
+Movie 1 released this week.
+Movie 2 released this week as well, while movie 3 released last week.
+I omit time fixed effects in this example since we focus on a single period; similarly, I omit time subscripts on the quantities.
+In this case, we will have the following data matrix where values in ``movie'' columns are used to estimate $\alpha$ values while ``age'' columns are used to estimate $\lambda$ values:
 
-best_delta = minimize(
-    lambda d: fit_model(d).sse, 
-    data, 
-    np.zeros(LEN_DELTA)
-).x
+\begin{center}
 
-best_model = fit_model(best_delta)
-\end{verbatim}
+\begin{tabular}{cccccc}
+    ln(Quantity) & Movie 1 & Movie 2 & Movie 3 & Age 0 & Age 1 \\
+    $\ln(q_{1})$ & 1 & $f(d_{12})$ & $f(d_{13})$ & $1 + f(d_{12})$ & $f(d_{13})$ \\
+    $\ln(q_{2})$ & $f(d_{12})$ & 1 & $f(d_{23})$ & $1 + f(d_{12})$ & $f(d_{23})$ \\
+    $\ln(q_{3})$ & $f(d_{13})$ & $f(d_{23})$ & 1 & $f(d_{13}) + f(d_{23})$ & $1$ 
+\end{tabular}
 
-I have not yet estimated this model, but I believe this procedure should be consistent. Unfortunately analytic standard errors likely won't work, but I can always bootstrap them.
+\end{center}
+
+Regressing quantity on the given Xs identifies the coefficients for $\alpha_1$, $\alpha_2$, $\alpha_3$, $\lambda_0$, and $\lambda_1$ via the first through fifth columns, respectively.
+These results make some amount of intuitive sense; if $\ln(q_1)$ performs below expectations, one explanation is that $\alpha_2$ is large--represented by a large coefficient on the (negative) $f(d_{12})$.
+
+Note the above is all taking $f$ as given. Following \textcite{magnolfi2022}, let $f$ be a cubic polynomial in distance; that is:
+
+$$f(d_{ij}) = \gamma_0 + \gamma_1 d_{ij} + \gamma_2 d_{ij}^2 + \gamma_3 d_{ij}^3$$
+
+This functional form makes few assumptions about the relationship between distance and substitutability, allowing for flexible sign, monotonicity, and concavity.
+Thus, if the final result \emph{does} exhibit those properties, we know it is coming from the data rather than baked into the model.
+
+Since the $\lambda$ and $\alpha$ coefficients in the model above interact multiplicatively with the $\gamma$ coefficients in $f$, we cannot estimate $\gamma$ at the same time that we estimate $\lambda$ and $\alpha$ linearly.
+Instead, I will choose a $\gamma$ coefficient vector and estimate $\lambda$ and $\alpha$ conditional on that $\gamma$ vector using OLS.
+We know the resulting model minimizes mean squared error \emph{in the set of models using this $\gamma$}.
+By adjusting the $\gamma$ vector to minimize mean squared error of these optimized models, I search across the lower envelope and know my final model minimizes mean squared error over the whole $\gamma, \alpha, \lambda$ space.
+
+Note this optimization is very expensive, as it requires estimating thousands of $\alpha$ coefficients for each interation.
+To speed estimation, I first optimize $\gamma$ using data from 2000 alone (the first ``quality'' year of my data).
+Conditional on this optimized $\gamma$, I then estimate $\alpha$ and $\lambda$ on all releases from 2000-2019 (omitting 2020 onward due to the COVID-19 pandemic).
+This allows me to estimate the full model in a reasonable amount of time.
+In the future, I could potentially run the optimization over the whole dataset using the research computing cluster; however, optimizing on my personal computer is already quite time-consuming even on the small sample.
 
@@
 
-To avoid the impact of the COVID-19 pandemic, I restrict my sample to movies released in 2019 or before.
-Similarly, I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
+While BoxOfficeGuru is getting its footing in the late 1990s, the data is reported less consistently; thus, I begin my sample in 2000.
+To avoid the impact of the COVID-19 pandemic, I end my sample in 2019.
+The administrator of BoxOfficeGuru is on vacation a few weekends a year, so I have a few missing weekends in my sample.
+I drop these weekends from the sample, as I do not have a good way to impute the missing data.
+This yields a grand total of 901 weekends in my sample.
 
-Though I cannot find documentation for their methodology, genres on TMDB appear to be ordered; that is, the ``primary'' genre of a movie is listed first on the page.
-This is evident based on genre ordering reversals; for example, \emph{Godzilla x Kong: The New Empire} (a recent Kaiju movie) lists ``action'' before ``adventure;'' on the other hand, \emph{Kingdom of the Planet of the Apes} lists ``adventure'' before ``action.'' Thus, we know the ordering of genres is idiosyncratic by movie.
-When I must assign a genre to a movie, I use the first genre listed on TMDB.
+I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
+Thus, I observe 2,970 movies over the 20 year span.
 
@@
 
-\begin{table}
-    \begin{center}
-    \input{../elasticity_estimate/output/euclid_regression.tex}
-
-    \caption{Regression Results, Euclidean Distance}
-    \label{tab:results_euclid}
-    \end{center}
-\end{table}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\textwidth]{../elasticity_estimate/output/cross_elasticity_euclid.png}
-    \caption{Cross-Elasticity by Euclidean Distance}
-    \label{fig:cross_elasticity_euclid}
-    \end{center}
-\end{figure}
-
 Moving now to the specification using cosine distance, the regression results are visible in table \ref{tab:results_cosine}.
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{../elasticity_estimate/output/cross_elasticity_zoom.png}
+    \includegraphics[width=\textwidth]{../elasticity_estimate/output/gamma_zoomed.png}
 

### 2024-09-17 elasticity_estimate/code/estimate_elasticity.py

@@ -264,7 +264,7 @@ def main():
     guru['Date'] = pd.to_datetime(guru['Date'])
 
     # Keep only movies in 2019 or earlier
-    guru = guru[guru['Date'].dt.year <= 2019]
+    guru = guru[(2000 <= guru['Date'].dt.year) & (guru['Date'].dt.year <= 2019)]
 
     # Limit sample to movies with trends and embeddings
     movies_with_trends = guru['movie_id'].unique()
@@ -298,7 +298,6 @@ def main():
     #    get_neighbors(example, distances, tmdb, movie_ids)
 
 
-    # Per Nick: could include lambda in the regression and target that those coefficients are 1
     # For each date, pre-compute the values for each movie pair
     dates = guru['Date'].unique()
     date_movie_dict = {}
@@ -306,6 +305,7 @@ def main():
         date_movies = guru[guru['Date'] == date]
         date_movie_dict[date] = date_movies
 
+    print(f"Number of dates: {len(dates)}")
 
     # Profile the objective function
     #cProfile.run('compute_model_error(INITIAL_GUESS, guru, movie_id_to_index, date_movie_dict, distances)', 'output/profile.prof')

### 2024-09-17 draft/draft.tex
Revised results section to reflect new model

@@
 \usepackage{amsmath}
+\usepackage{appendix}
 
@@
 \author{Gideon Moore}
-\title{Industrial Organization III: \\ Final Paper}
-\date{June 9, 2024}
+\title{Estimating Substitution Using Text Embeddings: \\ Evidence from the Movie Industry}
 
@@
 \begin{abstract}
-    I propose a novel method for estimating competition between products using text embeddings.
-Applying this method to the movie industry, I estimate impacts of competition which are reasonable in both sign and magnitude.
-Facing additional competitors lowers a film's revenue, and more similar competitors are more damaging.
-I find that if a movie could replace a competitor in the 90th percentile of similarity with one in the 10th percentile of similarity, its revenue would increase by 0.32\% to 0.44\%.
+    REWRITE
 \end{abstract}
@@
 
-Using data from BoxOfficeGuru and TheMovieDB, I estimate the impact of competitor characteristics on film revenue for each weekend from 1997-2019.
+Using data from BoxOfficeGuru and TheMovieDB, I estimate the impact of competitor characteristics on film revenue for each weekend from 2000-2019.
 I find substitution patterns which are plausible in both magnitude and direction, and which are robust to the choice of distance metric.
 As a movie becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
-I find that replacing a competitor in the 90th percentile of similarity with one in the 10th percentile of similarity would increase revenue by 0.32\% to 0.44\%.
 
@@
 As an experience good, demand for a movie is driven heavily by novelty; thus, the same movie will experience much less demand in its second week than in its first.
-While we intuitively expect these $\lambdas$ to decrease with time, I do not impose this in the model; this will be a test of model fit later on.
+While we intuitively expect these $\lambda$ to decrease with time, I do not impose this in the model; this will be a test of model fit later on.
 Finally, I include an idiosyncratic shock $\xi_{it}$ to capture unobserved demand drivers.
@@
 In the future, I could potentially run the optimization over the whole dataset using the research computing cluster; however, optimizing on my personal computer is already quite time-consuming even on the small sample.
+
+Traditional standard error methods will struggle here.
+Since the analytic standard errors on the $\alpha$ and $\lambda$ coefficients do not factor in the uncertainty from the $\gamma$ estimation, they will understate the true uncertainty.
+I should use bootstrapping to estimate the true standard errors on all three types of coefficients; however, again this is computationally infeasible on my consumer laptop.
+If I were to run this on the research computing cluster, I could potentially estimate uncertainty on all of these coefficients.
 
@@
 
-I consider two methods of measuring embedding similarity.
-My first method is the most similar to MMS: I reduce the dimensionality of the 1,536-dimensional embedding to 6 dimensions using principal component analysis.
-From there, I can compute Euclidean distance between the PCA-reduced embeddings as a measure of similarity.
-
-While this is most similar to the MMS approach, my preferred specification uses cosine similarity of the full embeddings.
-This method uses all information contained in each embedding, rather than just the first six principal components.
-Moreover, OpenAI suggests the use of cosine similarity to compare embeddings in their documentation.
-
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
-\begin{table}[]
-    \begin{center}
-        \begin{tabular}{@{}lll@{}}
-        \toprule
-        \textbf{Baseline Movie}    & \textbf{Nearest Neighbor \#1}       & \textbf{Nearest Neighbor \#2}              \\ \midrule
-        \textit{Frozen}            & \textit{The Little Mermaid}         & \textit{The Princess Diaries 2}            \\
-        \textit{The Notebook}      & \textit{Collateral Beauty}          & \textit{I Still Believe}                   \\
-        \textit{Avengers: Endgame} & \textit{X-MEN: Days of Future Past} & \textit{Guardians of the Galaxy, Volume 3} \\
-        \textit{A Bug's Life}      & \textit{The Spongebob Movie}        & \textit{DC League of Super Pets}           \\ \bottomrule
-        \end{tabular}
-        \caption{Nearest Neighbors to Baseline Movies by Euclidean Distance}
-        \label{tab:euclidean}
-    \end{center}
-\end{table}
-
 % Please add the following required packages to your document preamble:
@@
 
-Nearest neighbors to some common movies under Euclidean and cosine distance are visible in tables \ref{tab:euclidean} and \ref{tab:cosine}, respectively.
-Euclidean distance appears to be capturing ``something''--the closest movies to \emph{Frozen} are both princess movies, and the closest movies to \emph{A Bug's Life} are both animated children's movies.
-However, I find the cosine-based neighbors to be more compelling: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
+Nearest neighbors to some common movies are visible in \ref{tab:cosine}.
+Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
 Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War} respectively.
-This improvement in neighbor quality leads me to prefer cosine similarity for my analysis.
+Thus, I feel comfortable using these embeddings to capture similarity between movies.
 
-\section{Elasticity Results}
+\section{Results}
 
-First I run the analysis using Euclidean distance in six-dimensional space, following MMS.
-The regression results are visible in table \ref{tab:results_euclid}.
-The joint regression is highly significant, though the individual coefficients are iffier; given only the coefficient on $\gamma_0$ is significant, it may not be \emph{distance} that matters in this specification, but instead just the \emph{number of competitors}.
-However, it is also possible the other coefficients are noisy because they covary.
-Since the results are difficult to interpret in numeric form, I plot the elasticity curve implied by the $\gamma$ coefficients in figure \ref{fig:cross_elasticity_euclid}.
-While the curve is not monotonically decreasing \emph{everywhere}, we do find that it is monotonically decreasing on the range containing the majority of the data.
-As hypothesized, we see that as a movie becomes more distant, it is less influential on the sales of its competitors.
-Down the line I will need to compute uncertainty in this curve.
-Given there is likely negative correlation in errors between $\gamma$ coefficients (e.g.
-overestimating $\gamma_0$ means we're probably underestimating $\gamma_1$), simply propogating the analytic uncertainty will drastically overstate the true variance.
-I can get around this by bootstrapping, though given the computational intensity of the model I haven't done this yet.
+\subsection{Age}
 
-Putting some magnitudes on these numbers, suppose Disney were releasing \emph{Frozen} opposite \emph{Pok\'emon: The Movie 2000}--a movie in the top decile of similarity.
-If Disney were instead facing the horror movie \emph{Premonition}--a movie in the bottom decile of similarity to \emph{Frozen}--they would expect a 0.44\% revenue increase.
+First, let us check the age coefficients.
+The movie in my sample with the longest run is \emph{My Big Fat Greek Wedding,} which ran for 43 weeks.
+Estimating a weekly fixed effect for all 43 possible ages would absorb much of my power and create colinearity issues, as few movies reach this age.
+Instead, I topcode age.
+Initially I ran the model with a topcode of 26 weeks; however, point estimates converged after only 9 while the standard errors become quite large.
+Given this inflection point, I impose a topcode of 9 weeks.
 
-Moving now to the specification using cosine distance, the regression results are visible in table \ref{tab:results_cosine}.
-Again, we see that the results are highly significant.
-Moreover, here the coefficients on \emph{distance} are also significant, even with the colinearity.
-The elasticity curve implied by the $\gamma$ coefficients is visible in figure \ref{fig:cross_elasticity}.
-While no longer monotonically decreasing, we see the pattern is still decreasing on net from the 10th percentile to the 90th percentile of the data.We also see decreasing behavior in the tails, as desired.
+\begin{center}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../elasticity_estimate/output/lambda_coefficients.png}
+        \caption{Age Coefficients}
+        \label{fig:age_coefs}
+    \end{figure}
+\end{center}
 
-Again putting some concrete meaning on these numbers, suppose Disney were releasing \emph{Frozen} opposite \emph{Inkheart}--a movie in the top decile of cosine similarity.
-If Disney were instead facing the gory historical action flick \emph{Medieval}--a movie in the bottom decile of cosine similarity to \emph{Frozen}--they would expect a 0.32\% revenue increase.
+The estimated $\lambda$ coefficients are visible in figure \ref{fig:age_coefs}.
+As hypothesized, the age fixed effects decline with time such that new movies are more popular than old ones.
+Recall that the dependent variable is \emph{log} quantity; thus, the linear decline we observe in the $\lambda$ coefficients implies an exponential decline in sales, as is expected in the film industry.
+Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are negative and decreasing suggests our model is well specified.
 
-\begin{table}
+\subsection{Quality Distribution}
+
 \begin{center}
-    \input{../elasticity_estimate/output/cosine_regression.tex}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../elasticity_estimate/output/alpha_coefficients.png}
+        \caption{Distribution of $\alpha$ Coefficients}
+        \label{fig:alpha_dist}
+    \end{figure}
+\end{center}
 
-    \caption{Regression Results, Cosine Similarity}
-    \label{tab:results_cosine}
+Now we can check the $\alpha$ coefficients.
+The distribution of $\alpha$ coefficients is visible in figure \ref{fig:alpha_dist}.
+The distribution is roughly symmetrical with a mean of 0.5.
+Again recall the log scale of the dependent variable; thus, this distribution implies that the distribution of quality in \emph{levels} is quite skew, again as expected in the film industry.
+
+\subsection{Elasticity}
+
+\begin{center}
+    \begin{figure}
+        \includegraphics[width=\textwidth]{../elasticity_estimate/output/gamma_zoomed.png}
+        \caption{Influence by Distance}
+        \label{fig:gamma_zoomed}
+    \end{figure}
 \end{center}
-\end{table}
 
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\textwidth]{../elasticity_estimate/output/gamma_zoomed.png}
+The impact of competition $f$ is visible in figure \ref{fig:gamma_zoomed}.
+The curve is plotted over the density of the data; thus, we believe the area over the top of the histogram is most precisely estimated, while the edges are less precise.
+I have included the 10th and 90th percentiles of the data for reference; I would suggest focusing on the patter within this range.
 
-    \caption{Cross-Elasticity by Distance}
-    \label{fig:cross_elasticity}
-    \end{center}
-\end{figure}
+This curve exhibits a number of desirable characteristics on the support of the data:
+\begin{itemize}
+    \item It is negative for similar movies, suggesting that these movies are substitutes
+    \item It is (roughly) monotonic and attenuating with distance, suggesting that as movies become less similar their competitive impact declines
+    \item It approaches zero as distance becomes large, suggesting that movies which are very dissimilar have no impact on each other
+    \item It is concave, suggesting that competition is most important when movies are similar but less so when they are dissimilar
+\end{itemize}
 
-Note the magnitudes of these elasticity estimates are both quite similar and precisely estimated!
-Moreover, the $R^2$ of the two regressions is also nearly identical.
-This suggests that the distance metric we use is not particularly critical for our results.
+The curve becomes positive for a portion of the domain.
+This may be an artifact of trying to fit a cubic to a function which is actually flat in this region since competition has ceased to matter (and thus this is actually 0).
+However, I can think of a couple reasons this phenomenon might be real:
+\begin{itemize}
+    \item If two movies are dissimilar enough, they may be complements rather than substutites.
+For example, if a teenage child wants to see \emph{Friday the 13th}, parents may drop off the child and take their younger sibling to see \emph{Frozen}, when they would not have seen \emph{Frozen} otherwise.
+    \item Theater crowding may push people to see movies they would not have otherwise seen.
+If a theater is showing \emph{Friday the 13th,} this may crowd out an extra showing of \emph{The Lego Movie}.
+This reduced supply of \emph{The Lego Movie} drives people to see \emph{Frozen} instead.
+\end{itemize}
 
-It is reasonable to ask how well these trends are explained by movie genre alone.
-When I include a dummy for ``only film of genre'' showing in a given weekend, I find it does not materially change these results.
-Down the line I will likely want to do something more intensive--e.g., compute how similar genres are, then do something with ``genre distance'' alone rather than using the text information.
-Even if the genre \emph{can} explain much of the variation, I still believe this method is a contribution in that researchers are more likely to have raw, uncleaned text data than to have properly coded genre data.
+It is difficult to interpret the value of this function in a vacuum since it always interacts with $\delta_{it}$.
+However, using the $\alpha$ and $\lambda$ estimates above, I can compute some examples to help benchmark the magnitude of these coefficients.
+
+We know the average movie's $\alpha_i$ is roughly 0.5.
+Similarly, the $\lambda_{r(i) - t}$ in the week of a movie's premier is roughly 2.5.
+Thus, the average movie's appeal $\delta_{it}$ in its first week is 3.
+
+Suppose movie $i$ is attempting to estimate the impact of competitor $j$'s premier, anticipating $j$ is an average movie.
+If $j$ is in the 90th percentile of similarity to $i$, then $i$'s log quantity would be reduced by $f(d_{ij}) \cdot \delta_{jt} = -0.01 \cdot (0.5 + 2.5) = -0.03$, or roughly 3\%.
+On the other hand, if $j$ is in the 10th percentile of similarity, $i$'s log quantity would instead by \emph{increased} by $f(d_{ij}) \cdot \delta_{jt} = 0.001 \cdot (0.5 + 2.5) = 0.003$, or roughly 0.3\%.
+Thus, the presence of the wrong competitor relative to the right one could reasonably lower films' box office revenues by 3.3\%.
+
+To put this magnitude in context, \textcite{follows2016SF} estimates that box office revenue makes up 42\% of a movie's revenue.
+Further, he argues that the average film earns a profit of roughly 3.7\%.
+If a film experiences a 3.3\% decline in 42\% of its revenue, its total revenue falls by 1.3\%.
+This implies the profit margin in the presence of the additional competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
+
+To be even more concrete, let us consider specific movies.
+Ignoring the $\alpha$ coefficients for each of the following movies, suppose Disney were releasing \emph{Frozen} opposite gory historical action flick \emph{Medieval}: a movie in the bottom decile of cosine similarity to \emph{Frozen}.
+If Disney were instead facing children's fantasy romp \emph{Inkheart}--a movie in the top decile of similarity to \emph{Frozen}--they would expect a reduction in revenue similar to the one computed above as \emph{Inkheart} pulls away customers \emph{Medieval} would not.
+
+% Do I want to include ``genre clash'' here?
 
@@
 
-``Recommendation engines'' are a common tool in the tech world, used to predict what products a user will like.
-In some sense, we can think of this project as harnessing a new type of recommendation engine, identifying movies which are likely to be substitutable.
+A skeptic might argue the patterns I see are merely coincidental: why would we believe movies with similar descriptions necessarily appeal to the same audience?
+In this section, I use real movie reviews to validate my text-based model, showing that movies with similar descriptions indeed appeal to similar consumers.
+
+``Recommendation engines'' are a common tool in the tech world used to predict what products a user will like.
+We can think of this project as harnessing a new type of recommendation engine, identifying movies which are likely to be substitutable based on their descriptions.
 Algorithms in this space are broadly divided into two categories: collaborative filtering and content-based filtering.
 The text embeddings used above are a form of content-based filtering: using the ``content'' of the movie (its text description) to predict its appeal.
-However, one way to benchmark the quality of this approach is to compare them to the results of a collaborative filtering algorithm.
+However, one way to benchmark the quality of this approach is to compare against a \emph{collaborative} filtering algorithm.
 Unlike content-based filtering, collaborative filtering doesn't include any information about the content of a movie; instead, it uses the preferences of other users to predict what a user will like.
@@
 
-I implement UV-decomposition for matrix completion as described in \textcite{leskovec2020MMD}.\footnote{Note the algorithm described in the textbook optimizes element-wise, which is quite slow.
-Based on my own vector calculus, I've implemented \emph{matrix-wide} optimization, which seems to work.
+I implement UV-decomposition for matrix completion as described in \textcite{leskovec2020MMD}.\footnote{The algorithm described in the textbook optimizes element-wise, which is quite slow.
+Based on my own vector calculus visible in appendix \ref{app:uv}, I've implemented \emph{matrix-wide} optimization, which seems to work.
 However, I have not found a reference for this approach anywhere; I emailed Professor Leskovec to confirm the result is the same, but have not heard back.} Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
@@
 
+\begin{appendices}
+
+\section{Matrix-Wise Identification of UV Decomposition}
+\label{app:uv}
+
+Consider the following optimization problem:
+
+\end{appendices}
+
 \end{document}

### 2024-09-17 elasticity_estimate/code/estimate_elasticity.py

@@ -242,6 +242,12 @@ def main():
     # Compute log interest
     guru['ln_earnings'] = np.log(guru['Week Sales'])
 
+    # Convert date to datetime
+    guru['Date'] = pd.to_datetime(guru['Date'])
+
+    # Keep only movies in 2019 or earlier
+    guru = guru[(2000 <= guru['Date'].dt.year) & (guru['Date'].dt.year <= 2019)]
+
     # Drop if missing weeks
     guru = guru.dropna(subset=['Weeks'])
 
@@ -254,17 +260,20 @@ def main():
     plt.title('Histogram of Weeks Since Release')
     plt.savefig('output/weeks_histogram.png')
 
+    # Get the movie with the longest run
+    longest_run = guru[guru['Weeks'] == guru['Weeks'].max()]
+    print(longest_run)
+    kill
+
+
+
     # We want their weeks to also be zero-indexed so we can use it as an index
     guru['Weeks'] = guru['Weeks'] - 1
 
     # Truncate weeks to limit degrees of freedom
     guru['Weeks'] = np.minimum(guru['Weeks'], WEEK_THRESHOLD - 1)
 
-    # Convert date to datetime
-    guru['Date'] = pd.to_datetime(guru['Date'])
 
-    # Keep only movies in 2019 or earlier
-    guru = guru[(2000 <= guru['Date'].dt.year) & (guru['Date'].dt.year <= 2019)]
 
     # Limit sample to movies with trends and embeddings
     movies_with_trends = guru['movie_id'].unique()

### 2024-09-17 draft/draft.tex
Added UV decomposition appendix

@@
 To put this magnitude in context, \textcite{follows2016SF} estimates that box office revenue makes up 42\% of a movie's revenue.
-Further, he argues that the average film earns a profit of roughly 3.7\%.
+Further (acknowledging the difficulties of managing so-called ``Hollywood accounting''), he argues that the average film earns a profit of roughly 3.7\%.
 If a film experiences a 3.3\% decline in 42\% of its revenue, its total revenue falls by 1.3\%.
@@
 However, the matrix is generally sparse; most users have not reviewed most movies.
-UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV^T$ approximates the original matrix.
+UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV'$ approximates the original matrix.
 This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
@@
 
-Consider the following optimization problem:
+Let $M$ be an $m \times n$ matrix where most observations are missing.
+We wish to identify matrices $U$ and $V$, of dimension $m \times d$ and $n \times d$ respectively, such that $UV'$ approximates the observed entries of $M$, minimizing mean squared error.
+
+I've found it's easiest to start in the \emph{vector} case, then stack to get to the matrix case.
+Let us start by adjusting one $1 \times d$ row $u$ of $U$ at a time to minimize mean squared error.
+Let $m$ represent the corresponding row of $M$ impacted by changes in $u$.
+We wish to solve the following optimization problem:
+$$\min_{u} \{ (m - uV')(m - uV')' \} = \min_{u} \{mm' - mV'u' - uVm' + uVV'u' \} $$
+
+The dimensions of this objective imply the result is a scalar; thus, we can take the gradient with respect to $u$ and set it equal to zero to find the optimal $u$. The gradient is:
+$$\nabla_u = \vec{0} = -2mV' + 2uVV'$$
+
+Solving for the optimal $u^*$, this yields the following first-order condition:
+$$u^* = mV'(VV')^{-1}$$
+
+Entirely analogously, we can optimize a row of $V$ with respect to the existing $U$ to find:
+$$v^* = (U'U)^{-1}U'm$$
+
+From here, it is not hard to stack the vectors to get the matrix-wide optimization:
+
+$$U^* = MV'(VV')^{-1} \qquad V^* = (U'U)^{-1}U'M$$
+
+By generating random starting $U$ and $V$, we can iterate this process until convergence to find (locally) optimal $U$ and $V$ matrices.
+
+These matrix forms are particularly appealing since they highlight the procedure's implicit relation to linear regression.
+Philosophically, we are searching for $d$ latent characteristics of each movie (corresponding to rows of $V$) and the analogous coefficients on these characteristics which define each user's taste (corresponding to rows of $U$) such that the coefficient-characteristic pair justify the existing reviews.
+From there, we can extrapolate to missing movies by simply multiplying the latent characteristics by the coefficients for each missing user.
+Note since the problem is symmetric we could just as easily say the latent values for users are the ``characteristics'' and the values for movies are ``coefficients,'' but I prefer to think of movies as more easily defined by a finite set of characteristics than users.
 

### 2024-09-17 draft/draft.tex
Rewrote conclusion

@@
 
-Based on my findings this quarter, I am increasingly confident there is ``something'' here: using boilerplate embeddings of unstructured text characteristics, I pick up significant substitution patterns in the movie industry.
-If my work is successful, I would be the first to estimate cross-movie elasticities, while also providing a new method for estimating demand in industries with unstructured product characteristic text.
+In summary, I have developed a new method for estimating competition in characteristic space.
+By developing a new model of competition in the film industry, I can harness boilerplate OpenAI embeddings to estimate impact of competition on a movie's box office revenue.
 
-Over the summer, I hope to develop my paper in a few specific ways:
-\begin{itemize}
-    \item Enrich the model by interacting competitor age with cross-elasticity as discussed in section 4.2
-    \item Shore up arguments about substitution by comparing text-based similarity to review-based similarity as discussed in section 8
-    \item Confirm in the data that a movie's characteristics appear to be uncorrelated with the characteristics of its competitors as discussed in section 3.1
-\end{itemize}
+I find that movies with similar descriptions are indeed substitutes; moreover, substitutability falls as distance in characteristic space increases.
+My magnitudes are large but reasonable; changing one competitor movie from the bottom percentile of similarity to the top percentile can reduce profitability by 47\%.
 
-On a broader timeline, I also want to think about some way to address marketing.
-While the characteristics of the movies themselves may be set in advance, the characteristics \emph{emphasized in marketing} may be different.
-For example, if two romantic comedies come out in one weekend, both may benefit if one emphasizes the romantic half of the movie while the other emphasizes the comedic half.
-I am still thinking about clever ways to address this.
-TMDB contains some movie trailers, so one approach would be to see if similar \emph{movies} also have similar \emph{trailers}, even if they're released at roughly the same time.
+Two frontier collaborative filtering algorithms, UV decomposition and topic-specific PageRank, confirm the validity of my text similarity measure.
+I find that the patterns I see in the text-based model are not coincidental, but reflect consumers' true revealed preferences.
 

### 2024-09-18 draft/draft.tex
Revised draft per Shosh's old comments

@@
 \begin{abstract}
-    REWRITE
+    Using text descriptions of movies in conjunction with weekly box office receipts, I develop a novel model of characteristic-space competition in the movie industry.
+By exploiting plausably exogeneous variation in movie release windows, I identify the impact of competitor characteristics on movie revenue.
+As movies become more similar, the impact of competition increases.
+Due to the film industry's thin profit margins and high fixed costs, replacing a competitor in the 10th percentile of similarity with one in the 90th percentile can reduce profit by as much as 47\%.
 \end{abstract}
@@
 Demand estimation is a central pillar of the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
-However, demand estimation is often limited to markets with easily quantifiable characteristics, such as the storage in the market for hard drives or wattage in the market for light bulbs.
-We should not ignore markets with difficult-to-quantify characteristics simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a firm monopolizing comedy films is likely to be bad for consumers.
+However, these methods are often limited to markets in which goods have quantifiable characteristics such as storage in the market for hard drives or wattage in the market for light bulbs.
+Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a firm monopolizing ``funny'' films is likely to be bad for consumers.
 Thus, it is important economists develop tools to estimate demand in markets for goods with difficult-to-quantify characteristics.
+
+Product-level demand systems such as the Almost Ideal Demand System \parencite{deaton1980AER} can estimate rich substitution patterns without any characteristic data by looking only at substitution patterns under variation; that is, the substitution patterns of good $i$ are uninformative about the patterns of good $j$, even if $i$ and $j$ are actually very similar.
+These methods often require onerous quantities of data on consumer behavior.
+For example, estimation of AIDS requires sufficient variation to estimate parameters quadratic in the number of goods.
+A corollary of this is that these methods require goods to have \emph{already entered the market}.
+If a firm wishes to estimate substitution for a new product, these methods will have little to offer them.
 
@@
 Thus, we can predict which products compete most closely with others and estimate how the presence of one product impacts the sales of another.
+Unlike product-based demand estimation methods, this method can inform research on products which have not yet entered the market, and can fit rich substitution patterns while tuning relatively few parameters.
 
@@
 First, movies have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if you haven't seen \emph{Spider-Man}, you likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
-Second, movie characteristics are often amorphous: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a movie is--and harder still given the lack of clean data.
+Second, movie characteristics are often difficult to fit into a traditional econometric framework: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a movie is--and harder still given the lack of clean data.
 Finally, the movie industry has specific market characteristics which make demand estimation straightforward conditional on my method.
-Uniform pricing shuts down the endogeneous pricing channel of competition, and long production pipelines force studios to choose characteristics without knowing anything about their competitors.
+Uniform pricing shuts down the endogeneous pricing channel of competition, and long production pipelines force studios to choose characteristics without perfect knowledge of their competitors.
 
@@
 First, I implement a novel use for text embeddings, innovating on the text-as-data literature in economics.
-Second, given my demand estimation method is based on position in latent characteristic space, I am also closely connected to existing work on spatial competition.
+Second, given that my demand estimation method is based on position in latent characteristic space, I am also closely connected to existing work on spatial competition.
 Finally, given my film case study, I also contribute to the literature on the economics of the film industry.
@@
 
-I would argue these models work well for long documents such as patents and books, but struggle when working with (what I will call) ``terse'' texts.
+As argued in \textcite{kenter2015P2AICIKM}, bag-of-words and string comparison models struggle when working with short texts.
 If one paragraph includes the word ``wizardry'' while the other includes the word ``sorcery,'' a bag of words model will not capture the similarity between the two paragraphs.
-In large enough documents, there is some idea that the two words are likely to appear in similar contexts, but this is difficult to rely on when the sample text is only one or two sentences.
+In large enough documents, it is more likely the two words will appear in similar contexts, but this is difficult to rely on when the sample text is only one or two sentences.
 
-For terse contexts such as mine, I prefer using \emph{text embedding methods} as pioneered by Google's word2vec architecture \parencite{mikolov2013}.
+When working with short snippets of text like movie descriptions, it is more profitable to pursue \emph{text embedding methods} as pioneered by Google's word2vec architecture \parencite{mikolov2013}.
 Rather than counting individual words, word2vec identifies similar words using co-occurances in a training sample, then maps these words to a lower-dimensional space.
 This allows for the identification of similar texts even when they share few or no words, since it is able to identify that synonyms like ``wizardry'' and ``sorcery'' map to the same concept.
+I use OpenAI's frontier GPT-3 embeddings, which embeds not only words but entire paragraphs in order to capture a broader context.
 
-word2vec, while an improvement on bag of words, still overlooks \emph{interactions} between words.
+% word2vec, while an improvement on bag of words, still overlooks \emph{interactions} between words.
 While ``climate'' and ``change'' have meanings of their own, ``climate change'' is a largely distinct phenomenon.
@@
 
-This is an improvement upon much of the economics literature, which still relies on bag of words methods.
+% This is an improvement upon much of the economics literature, which still relies on bag of words methods.
 For example, \textcite{baker2016QJE} uses counts of ``uncertain'' words to estimate economic policy uncertainty.
@@
 The most similar paper to mine is \textcite{compiani2023} which uses text characteristics to estimate demand for technology products on Amazon.
-My paper innovates on theirs in multiple important ways.
-First, I have a much broader market; rather than focusing on fewer than 3,000 purchases in a specific Amazon retailer, I examine the entire United States movie market for multiple decades.
-Second, assumptions for demand estimation are much more plausible in the movie industry than in consumer electronics: the Kindle 8 is clearly designed endogeneously to the characteristics of the Kindle 7, whereas movies have long production pipelines which lock in product charactertics before competitor characteristics are known.
-Third, I use more advanced text embedding methods, pushing past word frequency and BERT-based methods to OpenAI's GPT-3 embeddings.
-Fourth, I (will...) benchmark my substitution patterns against ``real world'' user preference data from the MovieLens data, confirming my estimates capture actual substitution patterns.
-Finally, I estimate substitution patterns much more flexibly, allowing substitution to vary non-linearly and non-monotonically with distance between two products.
+However, rather than using more traditional logit substitution, my model can estimate substition patterns non-linearly and non-monotonically with distance between two products.
+I also benchmark my substitution patterns against ``real world'' user preference data from the MovieLens data, confirming my estimates capture actual substitution patterns.
+Finally, rather than Euclidean distance, my similarity measure is based on the cosine between the embedding vectors; this measure hews more closely to the industry standard, is more efficient to compute at scale, and avoids curse-of-dimensionality issues which often arise when taking Euclidean distances in high-dimensional space.
 
@@
 
-My closest predecessor studying competition in embedding space is \textcite{magnolfi2022} (hereafter MMS), who use t-Stochastic Triplet Embeddings to estimate competition in the cereal industry.
+My closest predecessor studying competition in embedding space is \textcite{magnolfi2022}, who use t-Stochastic Triplet Embeddings to estimate competition in the cereal industry.
 I take heavy inspiration from their model, however substitute frontier text embedding methods from OpenAI for their triplet-based embeddings.
@@
 
-Though I do not contribute to it, I benefit greatly from research documenting the uniform pricing puzzle \parencite{orbach2007IRLE, gil2009MS, ho2018MR}.
+I benefit greatly from research documenting the uniform pricing puzzle \parencite{orbach2007IRLE, gil2009MS, ho2018MR}.
 This literature extensively documents movie theaters' use of uniform pricing across heterogeneous movies.
@@
 
-Much of my understanding of the economics of film literature comes from a survey by \textcite{mckenzie2012ES}.
 I contribute to the ongoing literature founded by \textcite{prag1994CE} of determinents of box office demand.
@@
 \textcite{elberse2003MS} model both the supply and the demand for movies, finding that theatres' decision to screen a movie at all is an important factor for revenue.
-\textcite{devany1996E, devany1997EI, devany1999JCE, devany2004EDC} examine many factors of box office revenue, including the impact of word-of-mouth, the hazard rate of movie retirement, and estimating the film production function.
-\textcite{ravid2004B} study the impact of film \emph{content} on performance, finding that violence increases film profitability while sexual content does not.
+\textcite{devany1996E, devany1997EI, devany1999JCE, devany2004EDC} study several factors in box office revenue, including word-of-mouth, movie retirement, and the film production function.
+\textcite{ravid2004B} examine the impact of film \emph{content} on performance, finding that violence increases film profitability while sexual content does not.
 
-To the best of my knowledge, I am the first paper to examine the \emph{interplay} between competing films: how much is my revenue impacted by the existence of box office competitors?
+To the best of my knowledge, this paper is the first to examine the \emph{interplay} between competing films: how much is my revenue impacted by the existence of box office competitors?
 
@@
 
-In the tradition of industrial organization, I use an industry case study of the film industry to highlight the value of my method.
-Text characteristics are distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{The Dark Knight} and \emph{Ant-Man} are big-budget superhero action movies, but even a quick glance at the movie descriptions makes clear they are not substitutable.
-This is particularly important given movies are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
+I use an industry case study of the film industry to highlight the value of my method.
+Text is distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{Batman Begins} and \emph{Ant-Man} are big-budget superhero action movies, but even a quick glance at the movie descriptions makes clear they are not substitutable.\footnote{\emph{Batman Begins}: ``Driven by tragedy, billionaire Bruce Wayne dedicates his life to uncovering and defeating the corruption that plagues his home, Gotham City.
+Unable to work within the system, he instead creates a new identity, a symbol of fear for the criminal underworld.'' \emph{Ant-Man}: ``Armed with the astonishing ability to shrink in scale but increase in strength, master thief Scott Lang must embrace his inner-hero and help his mentor, Doctor Hank Pym, protect the secret behind his spectacular Ant-Man suit from a new generation of towering threats''} This is particularly important given movies are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
 Recognizing this, we expect firms put more effort into the text descriptions of their movies than comparable firms might in more traditional goods markets, as these decisions are distinctively important to customer decisionmaking.
@@
 
-When estimating the impact of competitors' characteristics on demand for a movie, its characteristics must be exogeneous to the characteristics of its competitors.
-If characteristics are chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor.
-
-Fortunately, movie production timelines are both long and secretive.
+Movie production timelines are both long and secretive.
 Since studios must purchase a script, hire a cast, and shoot all \emph{before} promoting a movie, it is difficult for a film to change its characteristics in response to its competitors' characteristics.
-I posit the existence of ``twin movies'' as evidence of this characteristic inelasticity.
-A pair of ``twin movies'' are two movies released in close proximity with very similar characteristics; the most famous examples, \emph{A Bug's Life} and \emph{Antz}, both released in fall of 1998 and feature early CGI animation of ants rebelling against oppression.
+To provide suggestive evidence of this inelasticity, I highlight the existance of ``twin movies:'' pairs of movies released in close proximity with very similar characteristics.
+The most famous examples, \emph{A Bug's Life} and \emph{Antz}, both released in fall of 1998 and feature early CGI animation of ants rebelling against oppression.
 
@@
 Given filmmakers' expressed distaste for twin movies, why do they continue to exist?
-It suggests that studios are unable to change their movies' characteristics in response to their competitors' characteristics; at the point I learn of \emph{A Bug's Life}, it is too late for me to adjust \emph{Antz} to be more distinct.
+It suggests that studios are unable to change their movies' characteristics in response to their competitors' characteristics; at the point Universal learns of \emph{A Bug's Life}, it is too late for them to adjust \emph{Antz} to be more distinct.
 Thus, our demand estimation coefficients are identified--they do not capture the strategic interplay of positioning and substitutability, but instead the direct impact of competition on film sales.
 
-This should produce testable hypotheses: conditional on the characteristics of a movie, the characteristics of its competitors should ``look like'' the characteristics of movies releasing in other weekends.
+Given the above,our demand model assumes characteristics are exogeneous.
+If instead they were chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor.
+However, the existance of twin movies suggests this is not a major concern.
+
+% This should produce testable hypotheses: conditional on the characteristics of a movie, the characteristics of its competitors should ``look like'' the characteristics of movies releasing in other weekends.
 One complication to this is we think movies have seasonality, so we can't simply compare movies within the same year in different months.
@@
 First, we know \emph{movies do not compete on price.} Under more flexible pricing, we might believe two movies in close priximity would lower prices to compete for the same audience, raising the quantity sold overall.
-However, since prices are fixed we know that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
+However, since prices are fixed we assume that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
 
@@
 However, since prices are fixed, we can understand revenue as simply the quantity of tickets sold multiplied by some scalar price.
-Thus, even without observing quantity, I am confident that observed revenue is a close substitute.
 
@@
 Again due to the uniform pricing scheme, we can use revenues as a proxy for quantity, since quantity is simply a constant scalar transformation of revenue.
-Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Memorial Day than on a random weekend in July.
+Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Christmas Day than on a random weekend in sunny July.
 Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation.
@@
 A movie's appeal is composed of two parts.
-First, a fixed effect $\alpha_i$ captures the movie's quality and characteristics.
+First, a fixed effect $\alpha_i$ captures the movie's quality.
 For example, we would expect \emph{Star Wars} to have a high $\alpha_i$ to capture the fact it is a popular movie in a vacuum separate from its competitors.
@@
 
-To capture the three attributes above, I propose the following functional form:
+To capture the three attributes above while also letting $d_{ij}$ enter flexibly, I propose the following functional form:
 
@@
 
-Thus, demand for $i$ is a linear function of its own appeal $\delta_{it}$, a weekend fixed effect $\alpha_t$, and its competitors' appeals $\delta_{jt}$ weighted by some function $f$ of distance $d_{ij}$.
+Demand for $i$ is thus a linear function of its own appeal $\delta_{it}$, a weekend fixed effect $\alpha_t$, and its competitors' appeals $\delta_{jt}$ weighted by some function $f$ of distance $d_{ij}$.
 
@@
 Moreover, the $f$ function interacts multiplicatively with the $\delta_{jt}$ terms, making matrix inversion ineffective.
-With some clever rearranging, however, this model becomes more tractable:
-
+With some clever rearranging, however, this model becomes more tractable:\footnote{Gory algebraic detail of this rearrangement is available in appendix \ref{app:rearrangement}.}
 \begin{align*}
     \ln(q_{it}) = \alpha_i &+ \sum_{j \neq i} f(d_{ij}) \alpha_j \\
-    &+ \lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(j)} \left(\sum_{j}^{r(j) = r(k)} f(d_{ij}) \right)\right) \\
+    &+ \lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(k)} \left(\sum_{j}^{r(j) = r(k)} f(d_{ij}) \right)\right) \\
     &+ \alpha_t \\
-    &+ \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt}
+    &+ \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} \qquad \bigg\} \text{ Mean 0 given $f(d_{ij}) \perp \xi_{jt}$}
 \end{align*}
 
-Thus, conditional on the values of $f(d_{ij})$, each movie's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend.
+Thus, conditional on the values of $f(d_{ij})$, each movie's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend plus mean-zero noise.
 Thus, the $\lambda$ and $\alpha$ coefficients are estimable using OLS.
@@
 
-Note this optimization is very expensive, as it requires estimating thousands of $\alpha$ coefficients for each interation.
+Note this optimization is very expensive, as it requires estimating thousands of $\alpha$ coefficients for each interation as well as computing potentially tens of millions of values of $f(d_{ij})$.
 To speed estimation, I first optimize $\gamma$ using data from 2000 alone (the first ``quality'' year of my data).
@@
 I should use bootstrapping to estimate the true standard errors on all three types of coefficients; however, again this is computationally infeasible on my consumer laptop.
-If I were to run this on the research computing cluster, I could potentially estimate uncertainty on all of these coefficients.
+If I were to run this on the research computing cluster, I could potentially estimate uncertainty on all of these coefficients via bootstrapping.
 
@@
 For each weekend, the website lists the top 10-20 movies by North American box office revenue.
-I have nearly the complete set of weekends over this 27 year span; however, when Gitesh in on vacation (a couple weekends a year) the website is not updated.
-An example of the data is visible in figure \ref{fig:guru}.
+
+While BoxOfficeGuru is getting its footing in the late 1990s, the data is reported less consistently; thus, I begin my sample in 2000.
+To avoid the impact of the COVID-19 pandemic, I end my sample in 2019.
+When Pandya is on vacation the website is not updated, so I have a few missing weekends in my sample interspersed thoughout my period of interest.
+I drop these weekends from the sample, as I do not have a good way to impute the missing data.
+This yields a grand total of 901 weekends for analysis.
 
@@
     \includegraphics[width=\textwidth]{images/guru.jpg}
-    \caption{BoxOfficeGuru Page for May 17-29, 2019}
+    \caption{BoxOfficeGuru Page for May 17-19, 2019}
     \label{fig:guru}
@@
 Example movie descriptions are visible in table \ref{tab:tmdb_desc}.
+
+TMDB also lists the primary language of each movie.
+I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
+In total,my sample contains 2,970 movies over the 20 year span matched between BoxOfficeGuru and TheMovieDB.
 
@@
 
-\subsection{Sample Selection}
-
-While BoxOfficeGuru is getting its footing in the late 1990s, the data is reported less consistently; thus, I begin my sample in 2000.
-To avoid the impact of the COVID-19 pandemic, I end my sample in 2019.
-The administrator of BoxOfficeGuru is on vacation a few weekends a year, so I have a few missing weekends in my sample.
-I drop these weekends from the sample, as I do not have a good way to impute the missing data.
-This yields a grand total of 901 weekends in my sample.
-
-I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
-Thus, I observe 2,970 movies over the 20 year span.
-
 \section{Text Embeddings}
@@
 Thus, traditional bag-of-words methods are likely to suffer, as word or bigram counts would be remarkably sparse.
-Because the OpenAI embedding model is trained on an enormous corpus of text, it can infer when descriptions are similar even when they share \emph{no} words--for example, a movie about ``wizardry'' would be similar to one about ``sorcery,'' which would not be true under a bag-of-words model.
+Because the OpenAI embedding model is trained on an enormous corpus of text, it can infer when descriptions are similar even when they share \emph{no} words--for example, a movie about ``wizardry'' would be similar to one about ``sorcery,'' which would not be true under a bag-of-words model \parencite{brown2020a}.
 
-Secondly, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level!), while the OpenAI embeddings are generally more context-aware.
+Secondly, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level), while the OpenAI embeddings are generally more context-aware.
 Given the movie descriptions tend to be longer than a single sentence, I believe the OpenAI embeddings are likely to be more able to capture the full context of the movie description.
-Even so, I should likely also perform my procedure using BERT embeddings in my robustness checks.
 
-In the long run I will likely want to use \texttt{text-embedding-3-large} for better performance.
+% In the long run I will likely want to use \texttt{text-embedding-3-large} for better performance.
 For exploratory purposes, I use the smaller model due to cost; the large model costs roughly 10x as much as the small model, though produces a 3,072-dimensional embedding.
 
+\subsection{Quality Check: Genre}
+
 As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering.
-I find that the embeddings do indeed cluster by genre, as visible in figure \ref{fig:tsne}.
+For each movie, I identify the first genre listed on the TMDB page.Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}.
 Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.
@@
 
-\subsection{Similarity Measures}
+\subsection{Quality Check: Nearest Neighbors}
 
@@
 
-Nearest neighbors to some common movies are visible in \ref{tab:cosine}.
+Nearest neighbors to some common movies are visible in table \ref{tab:cosine}.
 Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
@@
 
-First, let us check the age coefficients.
+First, let us check the age coefficients $\lambda$.
 The movie in my sample with the longest run is \emph{My Big Fat Greek Wedding,} which ran for 43 weeks.
@@
 Instead, I topcode age.
-Initially I ran the model with a topcode of 26 weeks; however, point estimates converged after only 9 while the standard errors become quite large.
+Initially I ran the model with a topcode of 26 weeks; however, point estimates converged to be constant after only 9 while the standard errors become quite large.
 Given this inflection point, I impose a topcode of 9 weeks.
@@
 Recall that the dependent variable is \emph{log} quantity; thus, the linear decline we observe in the $\lambda$ coefficients implies an exponential decline in sales, as is expected in the film industry.
-Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are negative and decreasing suggests our model is well specified.
+Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are monotonically decreasing suggests our model is well specified.
 
@@
 The impact of competition $f$ is visible in figure \ref{fig:gamma_zoomed}.
-The curve is plotted over the density of the data; thus, we believe the area over the top of the histogram is most precisely estimated, while the edges are less precise.
-I have included the 10th and 90th percentiles of the data for reference; I would suggest focusing on the patter within this range.
+The curve is plotted over the density of the data; thus, the area over the top of the histogram should be most precisely estimated, while the edges are less precise.
+I have included the 10th and 90th percentiles of the data for reference; I would suggest focusing on the pattern within this range.
 
@@
 \begin{itemize}
-    \item It is negative for similar movies, suggesting that these movies are substitutes
-    \item It is (roughly) monotonic and attenuating with distance, suggesting that as movies become less similar their competitive impact declines
+    \item It is negative for similar movies, suggesting that additional competition reduces a movie's revenue
+    \item It is (roughly) monotonic and attenuating with distance, suggesting that as movies become more distant their competitive impact declines
     \item It approaches zero as distance becomes large, suggesting that movies which are very dissimilar have no impact on each other
@@
     \item If two movies are dissimilar enough, they may be complements rather than substutites.
-For example, if a teenage child wants to see \emph{Friday the 13th}, parents may drop off the child and take their younger sibling to see \emph{Frozen}, when they would not have seen \emph{Frozen} otherwise.
+For example, if a teenage child wants to see \emph{Friday the 13th}, parents may drop off the child and take the younger sibling to see \emph{Frozen}, when they would not have seen \emph{Frozen} otherwise.
     \item Theater crowding may push people to see movies they would not have otherwise seen.
@@
 
-To put this magnitude in context, \textcite{follows2016SF} estimates that box office revenue makes up 42\% of a movie's revenue.
+To put this magnitude in context, \textcite{follows2016SF} estimates that box office receipt makes up 42\% of a movie's revenue.
 Further (acknowledging the difficulties of managing so-called ``Hollywood accounting''), he argues that the average film earns a profit of roughly 3.7\%.
 If a film experiences a 3.3\% decline in 42\% of its revenue, its total revenue falls by 1.3\%.
-This implies the profit margin in the presence of the additional competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
+The author documents that the vast majority of a movie's cost is fixed rather than variable; thus, we can assume cost does not change given the presence of competitors.
+These figures together imply the profit margin in the presence of the additional competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
 
@@
 I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: topic-specific PageRank and UV decomposition.
+
+\pagebreak
 
@@
 
-I find that movies with similar descriptions are indeed substitutes; moreover, substitutability falls as distance in characteristic space increases.
-My magnitudes are large but reasonable; changing one competitor movie from the bottom percentile of similarity to the top percentile can reduce profitability by 47\%.
-
 Two frontier collaborative filtering algorithms, UV decomposition and topic-specific PageRank, confirm the validity of my text similarity measure.
@@
 
+Movies with similar descriptions are indeed substitutes; moreover, substitutability falls as distance in characteristic space increases.
+My magnitudes are large but reasonable; changing one competitor movie from the bottom decile of similarity to the top decile can reduce revenue by 3.3\%, decreasing total profitability by 47\%.
+
 \printbibliography
 
+\pagebreak
+
 \begin{appendices}
+
+\section{Rearranging the Model}
+\label{app:rearrangement}
+
+Let us start with the expression of the model as given in the model construction section: $$\ln(q_{it}) = \alpha_i + \lambda_{t - r(i)} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) (\alpha_j + \lambda_{t - r(j)} + \xi_{jt}) + \alpha_{t}$$
+
+From there, we can rearrange the right side one small step at a time:
+\begin{align*}
+    &= \alpha_i + \lambda_{t - r(i)} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) (\alpha_j + \lambda_{t - r(j)} + \xi_{jt}) + \alpha_{t} \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \sum_{j \neq i} f(d_{ij}) \lambda_{t - r(j)} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \underbrace{\sum_{r(k)} \left(\sum_{j \neq i}^{r(j) = r(k)} f(d_{ij}) \lambda_{t - r(k)} \right)}_{\text{Group competitors of same age}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \sum_{r(k)} \underbrace{\left(\lambda_{t - r(k)} \sum_{j \neq i}^{r(j) = r(k)} f(d_{ij}) \right)}_{\text{$\lambda_{t - r(k)}$ constant within group}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \underbrace{\lambda_{t - r(i)} \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij}) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(k)} \sum_j^{r(j) = r(k)} f(d_{ij}) \right)}_{\text{Break the term for $r(i)$, movies of age equal to $i$'s, out of sum}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \underbrace{\lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right)}_{\text{Group matching $\lambda_{t - r(i)}$}} + \sum_{r(k) \neq r(i)}\left(\lambda_{t - r(k)} \sum_j^{r(j) = r(k)} f(d_{ij}) \right) + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t
+\end{align*}
+
+Describing the same steps above in words:
+\begin{itemize}
+    \item First, we break up the sum and rearrange the equation so the $\alpha$ terms are together, the $\lambda$ terms are together, and the $\xi$ terms are together
+    \item Next we break down the sum over competitor $\lambda$ terms into two sums; the outer sum is over the age of the competitor, while the inner sum is over the competitors of that age
+    \item From there we pull out the element of the outer sum corresponding movies of the same age as movie $i$
+    \item Finally we can pull $\lambda_{t - r(i)}$ out of the combined sum of movie $i$'s age coefficient and its competitors' weighted age coefficients
+\end{itemize}
+
+Thus, in the resulting equation, each $\alpha$ and $\lambda$ parameter appears only once, multiplying an element that is constructed entirely from data conditional on $f$.
 

### 2024-09-18 elasticity_estimate/code/estimate_elasticity.py

@@ -112,6 +112,7 @@ def regress_given_gamma(gamma, guru, movie_id_to_index, date_movie_dict, distanc
         competitor_ages = date_movies['Weeks'].values
         competitor_indices = np.array([movie_id_to_index[competitor_id] for competitor_id in competitor_ids])
 
+        # CAN I PRE-COMPUTE THIS ONCE AT THE START OF THE FUNCTION, RATHER THAN RE-DOING IT FOR EACH ROW?
         fs_of_distances = cubic(distances[movie_index, competitor_indices])
         
         alpha_vals[competitor_indices] += fs_of_distances

### 2024-09-18 draft/draft.tex
Minor draft polish

@@
 
-Given the above,our demand model assumes characteristics are exogeneous.
+Given the above, our demand model assumes characteristics are exogeneous.
 If instead they were chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor.
@@
 Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation.
-Thus, I can write the demand for movie $i$ in week $t$ as:
+I can therefore write the demand for movie $i$ in week $t$ as:
 
@@
 As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering.
-For each movie, I identify the first genre listed on the TMDB page.Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}.
+For each movie, I identify the first genre listed on the TMDB page.
+Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}.
 Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.
@@
 
-Nearest neighbors to some common movies are visible in table \ref{tab:cosine}.
+Nearest neighbors to some popular movies are visible in table \ref{tab:cosine}.
 Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
@@
 If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
-I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: topic-specific PageRank and UV decomposition.
+I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: UV decomposition and topic-specific PageRank.
 
@@
 
+% \section*{Acknowledgments}
+% Thank you to Liran Einav and Shoshana Vasserman for their support both in and after IO III.
+Thanks also to Arun Chandrasekhar, Nick Bloom, Janet Stefanov, Nick Scott-Hearn, and Lauren Harris for their feedback on this project, as well as the rest of the 2024 IO III class.
+
 \printbibliography

### 2024-09-18 draft/draft.tex
Move all twin-x axes to stacked

@@
 Note that the TSPR scores are not symmetric; that is, the TSPR score of movie $x$ to movie $y$ is not necessarily the same as the TSPR score of movie $y$ to movie $x$.
-Thus, for each pair of movies, I actually plot \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
+Thus, for each pair of movies, I actually observe \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
 

### 2024-09-18 draft/draft.tex
Fixed genre plot

@@
 
-As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and highlight clustering.
+As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and emphasize clustering.
 For each movie, I identify the first genre listed on the TMDB page.
+When running t-SNE I limit to only the four most popular genres (action, drama, comedy, and horror) to keep the figure legible.
 Embeddings do cluster by genre, as visible in figure \ref{fig:tsne}.
-Generally, we see action and adventure movies grouped in the top-right quadrant, science fiction and fantasy in the bottom-right quadrant, and romance and comedies on the left; that is, it seems as through the embeddings are not just clustering \emph{within} a genre, but also grouping \emph{similar genres}.
+Generally we see comedies visible on the left, while actions are on the right and horrors are at the top (admittedly, the nebulous ``drama'' genre is more diffuse).
 

### 2024-09-18 tmdb/code/plot_genres.py

@@ -8,12 +8,16 @@ import re
 
 GENRE_REGEX = re.compile(r"\[(\d+)")
 
-# Load in the embeddings
-with open('embeddings/movie_descriptions.pkl', 'rb') as f:
-    movie_data = pickle.load(f)
+GENRE_LIST = ['Action', 'Comedy', 'Drama', 'Horror']
 
-movie_data['genre'] = movie_data['tmdb_genres'].apply(lambda x: re.search(GENRE_REGEX, x).group(1) if re.search(GENRE_REGEX, x) else None)
-movie_data['genre'] = movie_data['genre'].astype(pd.Int64Dtype())
+# Load in the embeddings from parquet
+movie_data = pd.read_parquet('embeddings/movie_descriptions.parquet')
+
+# movie_data['genre'] = movie_data['tmdb_genres'].apply(lambda x: re.search(GENRE_REGEX, x).group(1) if re.search(GENRE_REGEX, x) else None)
+
+# Drop observations with no genres
+movie_data = movie_data[movie_data['tmdb_genres'].apply(lambda x: x.size > 0)]
+movie_data['genre'] = movie_data['tmdb_genres'].apply(lambda x: x[0])
 
 # Label genre integers
 genre_labels = {
@@ -40,7 +44,18 @@ genre_labels = {
 
 movie_data['genre'] = movie_data['genre'].map(genre_labels)
 
-X = movie_data['embedding'].apply(lambda x: x.data[0].embedding).to_list()
+# Count movies by genre
+genre_counts = movie_data['genre'].value_counts()
+print(genre_counts)
+
+# Keep only the genres we want to plot
+movie_data = movie_data[movie_data['genre'].isin(GENRE_LIST)]
+
+# Filter out movies with no embeddings
+movie_data = movie_data.dropna(subset=['embedding'])
+
+
+X = movie_data['embedding'].tolist()
 X = np.array(X)
 
 # Create t-SNE model

### 2024-09-20 draft/draft.tex
Integrated Lauren's comments

@@
 However, these methods are often limited to markets in which goods have quantifiable characteristics such as storage in the market for hard drives or wattage in the market for light bulbs.
-Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a firm monopolizing ``funny'' films is likely to be bad for consumers.
+Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a studio monopolizing ``funny'' films is likely to be bad for consumers.
 Thus, it is important economists develop tools to estimate demand in markets for goods with difficult-to-quantify characteristics.
@@
 Movies are a natural laboratory for several reasons.
-First, movies have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if you haven't seen \emph{Spider-Man}, you likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
+First, movies have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if you someone hasn't seen \emph{Spider-Man}, they likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
 Second, movie characteristics are often difficult to fit into a traditional econometric framework: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a movie is--and harder still given the lack of clean data.
@@
 As a movie becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
+Replacing a competitor in the 10th percentile of similarity with one in the 90th percentile reduces box-office revenue by 3.3\%.
+Given films' thin profit margins, this is roughly a 47\% decline in profit.
 
@@
 
-This paper's primary contribution is to the literature on text as data, well summarized in \textcite{gentzkow2019EL}.
-The statistical analysis of text has a long history; the earliest example I have found is \textcite{mendenhall1887S} using word frequency to predict the author of a mysterious text.
+This paper's primary contribution is to the literature on text as data, which is well summarized in \textcite{gentzkow2019EL}.
+The statistical analysis of text has a long history; to my knowledge, the earlist example is \textcite{mendenhall1887S} using word frequency to predict the author of a mysterious text.
 This is an early example of \emph{bag of words} methods, which treat text as a collection of words rather than a structured object.
@@
 
-\subsection{Movies}
+\subsection{Economics of Film}
 
@@
 
-I use an industry case study of the film industry to highlight the value of my method.
-Text is distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{Batman Begins} and \emph{Ant-Man} are big-budget superhero action movies, but even a quick glance at the movie descriptions makes clear they are not substitutable.\footnote{\emph{Batman Begins}: ``Driven by tragedy, billionaire Bruce Wayne dedicates his life to uncovering and defeating the corruption that plagues his home, Gotham City.
+I use a case study of the film industry to highlight the value of my method.
+Text is distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{Batman Begins} and \emph{Ant-Man} are big-budget action movies about animal-based superheros, but even a quick glance at the movie descriptions makes clear they are not close substitutes.\footnote{\emph{Batman Begins}: ``Driven by tragedy, billionaire Bruce Wayne dedicates his life to uncovering and defeating the corruption that plagues his home, Gotham City.
 Unable to work within the system, he instead creates a new identity, a symbol of fear for the criminal underworld.'' \emph{Ant-Man}: ``Armed with the astonishing ability to shrink in scale but increase in strength, master thief Scott Lang must embrace his inner-hero and help his mentor, Doctor Hank Pym, protect the secret behind his spectacular Ant-Man suit from a new generation of towering threats''} This is particularly important given movies are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
@@
 Studios are generally inclined to avoid releasing a twin movie due to fear of excess competition.
-In 2016, French director Xavier Giannoli released his film \emph{Marguerite}, inspired by the true story of a New York Socialite turned failed opera singer, in the same year as 20th Century Fox's \emph{Florence Foster Jenkins} retelling the same story.
+In 2016, French director Xavier Giannoli released his film \emph{Marguerite}, inspired by the true story of a New York socialite turned failed opera singer, in the same year as 20th Century Fox's \emph{Florence Foster Jenkins} retelling the same story.
 Giannoli told the \emph{Independent}: ``For me, it was terrible...I work a lot as a writer to find completely original stories.
-I don't want the audience to have the feeling, `oh, I saw that!''' \parencite{mottram2016I}.
+I don't want the audience to have the feeling, `oh, I saw that!'" \parencite{mottram2016I}.
 
 Given filmmakers' expressed distaste for twin movies, why do they continue to exist?
-It suggests that studios are unable to change their movies' characteristics in response to their competitors' characteristics; at the point Universal learns of \emph{A Bug's Life}, it is too late for them to adjust \emph{Antz} to be more distinct.
-Thus, our demand estimation coefficients are identified--they do not capture the strategic interplay of positioning and substitutability, but instead the direct impact of competition on film sales.
+Studios are seemingly unable to change their movies' characteristics in response to their competitors' characteristics; at the point DreamWorks learns of \emph{A Bug's Life}, it is too late for them to adjust \emph{Antz} to be more distinct.
+Thus, the demand estimation coefficients are identified--they do not capture the strategic interplay of positioning and substitutability, but instead the direct impact of competition on film sales.
 
-Given the above, our demand model assumes characteristics are exogeneous.
+Given the above, my demand model assumes characteristics are exogeneous.
 If instead they were chosen as equilibrium objects, then the coefficients on proximity are no longer causal: the impact of locating near a close competitor becomes confounded with latent demand characteristics driving the \emph{choice} to locate near a close competitor.
@@
 
-First, we know \emph{movies do not compete on price.} Under more flexible pricing, we might believe two movies in close priximity would lower prices to compete for the same audience, raising the quantity sold overall.
-However, since prices are fixed we assume that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
+First, I argue \emph{movies do not compete on price.} Under more flexible pricing, it is possible two movies in close priximity would lower prices to compete for the same audience, raising the quantity sold overall.
+However, since prices are fixed I assume that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
 
-Second, we can \emph{use revenue as a proxy for quantity.} As described in \textcite{bond2021JME}, this assumption is often perilous since a variable markup means that revenue may not reflect true production.
-However, since prices are fixed, we can understand revenue as simply the quantity of tickets sold multiplied by some scalar price.
+Second, I \emph{use revenue as a proxy for quantity.} As described in \textcite{bond2021JME}, this assumption is often perilous since a variable markup means that revenue may not reflect true production.
+However, since prices are fixed, I can interpret revenue as simply the quantity of tickets sold multiplied by some scalar price.
 
@@
 Movie consumers face a discrete choice problem: which of the movies on offer this weekend should they see?
-Let a movie $i$'s attractiveness at time $t$ be given by some $\delta_{it}$.
+Let movie $i$'s attractiveness at time $t$ be given by some $\delta_{it}$.
 Consumers receive idiosyncratic demand shocks for each movie, and choose the movie which maximizes their utility.
@@
 For example, we would expect \emph{Star Wars} to have a high $\alpha_i$ to capture the fact it is a popular movie in a vacuum separate from its competitors.
-Second, movies have an age fixed effect $\lambda_{t - r(i)}$.
+Second, movies have an age fixed effect $\lambda_{t - r(i)}$, where $r(i)$ represents movie $i$'s release date.
 As an experience good, demand for a movie is driven heavily by novelty; thus, the same movie will experience much less demand in its second week than in its first.
-While we intuitively expect these $\lambda$ to decrease with time, I do not impose this in the model; this will be a test of model fit later on.
+While it is reasonable to expect these $\lambda$ to decrease with time, I do not impose this; instead, verifying this is true will be a test of model fit later on.
 Finally, I include an idiosyncratic shock $\xi_{it}$ to capture unobserved demand drivers.
@@
 
-Putting these terms into our demand equation, we have:
+Putting these terms into the demand equation, we have:
 
@@
 
-As an example, consider what our matrix will look like in the following market.
+As an example, consider what the data matrix will look like in the following market.
 Movie 1 released this week.
@@
 If I were to run this on the research computing cluster, I could potentially estimate uncertainty on all of these coefficients via bootstrapping.
+
+\subsection{Application}
+
+In some ways this model is a strange fit for ``characteristic-space'' competition, in that it includes movie fixed effects $\alpha_i$.
+One of the common virtues of using a characteristic-space model is that it allows us to estimate demand for goods which are not yet in the market by arguing based on similar characteristics.
+However, the econometrician cannot estimate $\alpha_i$ for a movie which has not yet released.
+
+While this is a limitation, I do not believe it to be as costly as it initially appears.
+First, as part of estimating the model, I will identify the distribution of $\alpha$ values.
+Thus, even with no information on a specific $\alpha_i$, it is possible for me to estimate the distribution of outcomes given the possible $\alpha_i$ values, computing both expected values and the variance around these values.
+
+Moreover, the \emph{additive separability} of the $\alpha$ and $\lambda$ terms goes grants certain counterfactuals extra relevance.
+Note in the model above $\alpha$ and $\lambda$ never interact multiplicatively.
+Thus, given competitor similarity $d_{ij}$ and ages $\lambda_{t - r(k)}$, it is possible to perform comparative statics agnostic to the $\alpha$ values.
+For example, I can ask how demand for movie $i$ will compare if it releases in week 1 rather than week 2 conditional on its competitors.
+Since the terms containing $\alpha_i$ will be identical in these two scenarios, they drop out and leave a concrete estimate of the change in quantity which is not dependent on the $\alpha_i$ values.
 
@@
 Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
-Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War} respectively.
+Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War}, respectively.
 Thus, I feel comfortable using these embeddings to capture similarity between movies.
@@
 
+\begin{figure}
     \begin{center}
-    \begin{figure}
         \includegraphics[width=\textwidth]{../elasticity_estimate/output/lambda_coefficients.png}
@@
         \label{fig:age_coefs}
+    \end{center}
 \end{figure}
-\end{center}
 
@@
 Recall that the dependent variable is \emph{log} quantity; thus, the linear decline we observe in the $\lambda$ coefficients implies an exponential decline in sales, as is expected in the film industry.
-Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are monotonically decreasing suggests our model is well specified.
+Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are monotonically decreasing suggests my model is well specified.
 
@@
 
+\begin{figure}
     \begin{center}
-    \begin{figure}
         \includegraphics[width=\textwidth]{../elasticity_estimate/output/alpha_coefficients.png}
@@
         \label{fig:alpha_dist}
+    \end{center}
 \end{figure}
-\end{center}
 
@@
 
+\begin{figure}
     \begin{center}
-    \begin{figure}
-        \includegraphics[width=\textwidth]{../elasticity_estimate/output/gamma_zoomed.png}
+        \includegraphics[width=\textwidth]{../elasticity_estimate/output/gamma_zoom_stacked.png}
         \caption{Influence by Distance}
         \label{fig:gamma_zoomed}
+    \end{center}
 \end{figure}
-\end{center}
 
@@
 
-To put this magnitude in context, \textcite{follows2016SF} estimates that box office receipt makes up 42\% of a movie's revenue.
+To put this magnitude in context, \textcite{follows2016SF,follows2016SFa} estimates that box office receipt makes up 42\% of a movie's revenue.
 Further (acknowledging the difficulties of managing so-called ``Hollywood accounting''), he argues that the average film earns a profit of roughly 3.7\%.
@@
 The author documents that the vast majority of a movie's cost is fixed rather than variable; thus, we can assume cost does not change given the presence of competitors.
-These figures together imply the profit margin in the presence of the additional competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
+These figures together imply the profit margin in the presence of the closer competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
 
@@
 
-With our decomposition in hand, it is straightforward to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
+With this decomposition in hand, it is straightforward to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
 If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
@@
 
-Rather than blindly trusting our decomposition, we can validate it by comparing predicted reviews against actual reviews.
+Rather than blindly trusting the decomposition, we can validate it by comparing predicted reviews against actual reviews.
 I hold out 5\% of the reviews as a test set, then train on the remaining 95\%.
@@
 
-These results affirm for me that our collaborative filtering results are likely to be quite predictive of a movie's appeal.
+These results affirm for me that the collaborative filtering results are likely to be quite predictive of a movie's appeal.
 Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
@@
 However, as the review count threshold increases, the coefficient on text similarity becomes increasingly positive and significant.
-This suggests that as UV decomposition becomes more accurate, it looks increasingly like our text-based cosine similarity measure.
+This suggests that as UV decomposition becomes more accurate, it looks increasingly like text-based cosine similarity.
 This is a strong validation of my text-based model; the patterns I see in the text embeddings are not coincidental, but reflect consumers' true preferences.
@@
 Moreover, while the number of pairs included remains large enough to have useful confidence intervals, the coefficient is increasing in the threshold value.
-If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like our text-based cosine similarity measure.
+If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like text-based cosine similarity.
 This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per movie to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.

### 2024-09-20 elasticity_estimate/code/estimate_elasticity.py

@@ -212,7 +212,7 @@ def main():
     # Label the genres
     plt.xticks(ticks=range(len(genre_counts)), rotation=45)
 
-    plt.savefig('output/genre_counts.png')
+    plt.savefig('output/genre_counts.png', bbox_inches='tight')
 
     # Clear plot
     plt.clf()
@@ -259,14 +259,11 @@ def main():
     plt.xlabel('Weeks Since Release')
     plt.ylabel('Count')
     plt.title('Histogram of Weeks Since Release')
-    plt.savefig('output/weeks_histogram.png')
+    plt.savefig('output/weeks_histogram.png', bbox_inches='tight')
 
     # Get the movie with the longest run
     longest_run = guru[guru['Weeks'] == guru['Weeks'].max()]
     print(longest_run)
-    kill
-
-
 
     # We want their weeks to also be zero-indexed so we can use it as an index
     guru['Weeks'] = guru['Weeks'] - 1
@@ -398,11 +395,11 @@ def main():
 
     # Add title
     ax1.set_title('Impact over Distance')
+    ax1.set_xlabel('Distance')
     ax2 = ax1.twinx()
 
     # Plot the empirical distances
     ax2.hist(distances, bins=20, alpha=0.5, density=True)
-    ax2.set_xlabel('Distance')
     ax2.set_ylabel('Density of Empirical Distance')
 
     ax1.plot(distance_grid, cross_elasticity, color='orange')
@@ -412,7 +409,7 @@ def main():
     ax2.axvline(x=bottom_end, color='red', linestyle='--')
 
     fig.tight_layout()
-    plt.savefig('output/gamma.png')
+    plt.savefig('output/gamma.png', bbox_inches='tight')
 
     # Zoom in on the cross-elasticity in the support of the data
 
@@ -440,13 +437,13 @@ def main():
     ax1.set_title('Impact over Distance')
 
     ax1.plot(distance_grid, cross_elasticity, color='orange')
+    ax1.set_xlabel('Distance')
     ax1.set_ylabel('Competition Function')
 
     ax2 = ax1.twinx()
 
     # Plot the empirical distances
     ax2.hist(distances, bins=20, alpha=0.5, density=True)
-    ax2.set_xlabel('Distance')
     ax2.set_ylabel('Density of Empirical Distance')
 
     # Add vertical lines at the top and bottom ends
@@ -454,7 +451,38 @@ def main():
[…]

### 2024-09-22 draft/draft.tex
Include grammar fixes from mom

@@
 \author{Gideon Moore}
-\title{Estimating Substitution Using Text Embeddings: \\ Evidence from the Movie Industry}
+\title{Estimating Substitution Using Text Embeddings: \\ Evidence from the Film Industry}
 
@@
 \begin{abstract}
-    Using text descriptions of movies in conjunction with weekly box office receipts, I develop a novel model of characteristic-space competition in the movie industry.
-By exploiting plausably exogeneous variation in movie release windows, I identify the impact of competitor characteristics on movie revenue.
-As movies become more similar, the impact of competition increases.
+    Using text descriptions of films in conjunction with weekly box office receipts, I develop a novel model of characteristic-space competition in the film industry.
+By exploiting plausably exogeneous variation in film release windows, I identify the impact of competitor characteristics on film revenue.
+As films become more similar, the impact of competition increases.
 Due to the film industry's thin profit margins and high fixed costs, replacing a competitor in the 10th percentile of similarity with one in the 90th percentile can reduce profit by as much as 47\%.
@@
 
-Demand estimation is a central pillar of the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
-However, these methods are often limited to markets in which goods have quantifiable characteristics such as storage in the market for hard drives or wattage in the market for light bulbs.
-Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult to quantify; though it is hard to put a number how funny a movie is, we understand that a studio monopolizing ``funny'' films is likely to be bad for consumers.
-Thus, it is important economists develop tools to estimate demand in markets for goods with difficult-to-quantify characteristics.
+Demand estimation methods are central pillars of the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
+However, these approaches are often limited to markets in which goods have quantifiable characteristics such as storage in the market for hard drives or wattage in the market for lightbulbs.
+Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult-to-quantify; though it is hard to put a number on how funny a film is, we understand that a studio monopolizing ``funny'' films is likely to be bad for consumers.
+Thus, it is important economists develop tools to estimate demand in markets where traditional methods struggle to perform well.
 
-Product-level demand systems such as the Almost Ideal Demand System \parencite{deaton1980AER} can estimate rich substitution patterns without any characteristic data by looking only at substitution patterns under variation; that is, the substitution patterns of good $i$ are uninformative about the patterns of good $j$, even if $i$ and $j$ are actually very similar.
+Product-level demand systems such as the Almost Ideal Demand System (AIDS) can estimate rich substitution patterns without any characteristic data by looking only at substitution patterns under variation; that is, the substitution patterns of good $i$ are uninformative about the patterns of good $j$, even if $i$ and $j$ are actually very similar \parencite{deaton1980AER}.
 These methods often require onerous quantities of data on consumer behavior.
@@
 I propose a new method for quantifying product characteristics using text descriptions.
-By embedding text descriptions of a product in characteristic space, we as econometricians can identify which products are most similar.
-Thus, we can predict which products compete most closely with others and estimate how the presence of one product impacts the sales of another.
+By embedding text descriptions of a product in characteristic space, econometricians can identify which products are most similar.
+Thus, they can predict which products compete most closely with others and estimate how the presence of one product impacts the sales of another.
 Unlike product-based demand estimation methods, this method can inform research on products which have not yet entered the market, and can fit rich substitution patterns while tuning relatively few parameters.
@@
 While this method has broad applicability, I use the film industry as a case study.
-Movies are a natural laboratory for several reasons.
-First, movies have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if you someone hasn't seen \emph{Spider-Man}, they likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
-Second, movie characteristics are often difficult to fit into a traditional econometric framework: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a movie is--and harder still given the lack of clean data.
-Finally, the movie industry has specific market characteristics which make demand estimation straightforward conditional on my method.
+Film is a natural laboratory for several reasons.
+First, films have a broad consumer base, such that readers have an intuitive sense whether my method ``makes sense''--even if someone hasn't seen \emph{Spider-Man}, they likely have a sense whether it is more similar to \emph{Man of Steel} or \emph{The Notebook}, and can use this lens to judge my results.
+Second, film characteristics are often difficult to fit into a traditional econometric framework: while we can easily measure the wattage of a lightbulb, it is much harder to put a number on how funny or violent a film is--and harder still given the lack of clean data.
+Finally, the film industry has specific market characteristics which make demand estimation using my method straightforward.
 Uniform pricing shuts down the endogeneous pricing channel of competition, and long production pipelines force studios to choose characteristics without perfect knowledge of their competitors.
@@
 I find substitution patterns which are plausible in both magnitude and direction, and which are robust to the choice of distance metric.
-As a movie becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
+As a film becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
 Replacing a competitor in the 10th percentile of similarity with one in the 90th percentile reduces box-office revenue by 3.3\%.
@@
 
-As argued in \textcite{kenter2015P2AICIKM}, bag-of-words and string comparison models struggle when working with short texts.
-If one paragraph includes the word ``wizardry'' while the other includes the word ``sorcery,'' a bag of words model will not capture the similarity between the two paragraphs.
+As argued in \textcite{kenter2015P2AICIKM}, bag of words and string comparison models struggle when working with short texts.
+If one paragraph includes the word ``wizardry'' while another includes the word ``sorcery,'' a bag of words model will not capture the similarity between the two paragraphs.
 In large enough documents, it is more likely the two words will appear in similar contexts, but this is difficult to rely on when the sample text is only one or two sentences.
 
-When working with short snippets of text like movie descriptions, it is more profitable to pursue \emph{text embedding methods} as pioneered by Google's word2vec architecture \parencite{mikolov2013}.
+When working with short snippets of text like film descriptions, it is more profitable to pursue \emph{text embedding methods} as pioneered by Google's word2vec architecture \parencite{mikolov2013}.
 Rather than counting individual words, word2vec identifies similar words using co-occurances in a training sample, then maps these words to a lower-dimensional space.
@@
 The most similar paper to mine is \textcite{compiani2023} which uses text characteristics to estimate demand for technology products on Amazon.
-However, rather than using more traditional logit substitution, my model can estimate substition patterns non-linearly and non-monotonically with distance between two products.
+However, rather than using more traditional logit substitution, my model can estimate substitution patterns non-linearly and non-monotonically with distance between two products.
 I also benchmark my substitution patterns against ``real world'' user preference data from the MovieLens data, confirming my estimates capture actual substitution patterns.
@@
 
-My closest predecessor studying competition in embedding space is \textcite{magnolfi2022}, who use t-Stochastic Triplet Embeddings to estimate competition in the cereal industry.
+My closest predecessor studying competition in embedding space is \textcite{magnolfi2024AM}, who use t-Stochastic Triplet Embeddings to estimate competition in the cereal industry.
 I take heavy inspiration from their model, however substitute frontier text embedding methods from OpenAI for their triplet-based embeddings.
@@
 I benefit greatly from research documenting the uniform pricing puzzle \parencite{orbach2007IRLE, gil2009MS, ho2018MR}.
-This literature extensively documents movie theaters' use of uniform pricing across heterogeneous movies.
+This literature extensively documents movie theaters' use of uniform pricing across heterogeneous films.
 While each paper suggests potential explanations for this phenomenon, my findings are agnostic to the true cause of uniform pricing; I simply use it as a convenient assumption to estimate demand.
@@
 I contribute to the ongoing literature founded by \textcite{prag1994CE} of determinents of box office demand.
-The authors find well-rated and well-advertised movies generally sell more tickets, and conditional on those two factors other observables are generally insignificant.
-\textcite{elberse2003MS} model both the supply and the demand for movies, finding that theatres' decision to screen a movie at all is an important factor for revenue.
-\textcite{devany1996E, devany1997EI, devany1999JCE, devany2004EDC} study several factors in box office revenue, including word-of-mouth, movie retirement, and the film production function.
+The authors find well-rated and well-advertised films generally sell more tickets, and conditional on those two factors other observables are generally insignificant.
+\textcite{elberse2003MS} model both the supply and the demand for films, finding that theatres' decision to screen a film at all is an important factor for revenue.
+\textcite{devany1996E, devany1997EI, devany1999JCE, devany2004EDC} study several factors in box office revenue, including word-of-mouth, film retirement, and the film production function.
 \textcite{ravid2004B} examine the impact of film \emph{content} on performance, finding that violence increases film profitability while sexual content does not.
@@
 
-\section{The Movie Industry}
+\section{The Film Industry}
 
 I use a case study of the film industry to highlight the value of my method.
-Text is distinctively important for the movie industry because movie characteristics are otherwise difficult to quantify; for example, both \emph{Batman Begins} and \emph{Ant-Man} are big-budget action movies about animal-based superheros, but even a quick glance at the movie descriptions makes clear they are not close substitutes.\footnote{\emph{Batman Begins}: ``Driven by tragedy, billionaire Bruce Wayne dedicates his life to uncovering and defeating the corruption that plagues his home, Gotham City.
-Unable to work within the system, he instead creates a new identity, a symbol of fear for the criminal underworld.'' \emph{Ant-Man}: ``Armed with the astonishing ability to shrink in scale but increase in strength, master thief Scott Lang must embrace his inner-hero and help his mentor, Doctor Hank Pym, protect the secret behind his spectacular Ant-Man suit from a new generation of towering threats''} This is particularly important given movies are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
-Recognizing this, we expect firms put more effort into the text descriptions of their movies than comparable firms might in more traditional goods markets, as these decisions are distinctively important to customer decisionmaking.
-Finally, the market for movies has two distinctive features which make it a strong candidate for demand estimation: long production pipelines and a tradition of uniform pricing.
+Text is distinctively important for the film industry because film characteristics are otherwise difficult to quantify; for example, both \emph{Batman Begins} and \emph{Ant-Man} are big-budget action films about animal-based superheros, but even a quick glance at the descriptions makes clear they are not close substitutes.\footnote{\emph{Batman Begins}: ``Driven by tragedy, billionaire Bruce Wayne dedicates his life to uncovering and defeating the corruption that plagues his home, Gotham City.
+Unable to work within the system, he instead creates a new identity, a symbol of fear for the criminal underworld.'' \emph{Ant-Man}: ``Armed with the astonishing ability to shrink in scale but increase in strength, master thief Scott Lang must embrace his inner-hero and help his mentor, Doctor Hank Pym, protect the secret behind his spectacular Ant-Man suit from a new generation of towering threats.''} This is particularly important given films are an ``experience good''--people rely on the description to inform their purchasing decision, as most people will not form opinions through repeat purchases.
+Recognizing this, we expect firms put more effort into the text descriptions of their films than comparable firms might in more traditional goods markets, as these decisions are distinctively important to customer decision-making.
+Finally, the market for films has two distinctive features which make it a strong candidate for demand estimation: long production pipelines and a tradition of uniform pricing.
 
@@
 
-Movie production timelines are both long and secretive.
-Since studios must purchase a script, hire a cast, and shoot all \emph{before} promoting a movie, it is difficult for a film to change its characteristics in response to its competitors' characteristics.
-To provide suggestive evidence of this inelasticity, I highlight the existance of ``twin movies:'' pairs of movies released in close proximity with very similar characteristics.
+Film production timelines are both long and secretive.
+Since studios must purchase a script, hire a cast, and shoot all \emph{before} promoting a film, it is difficult for a film to change its characteristics in response to its competitors' characteristics.
+To provide suggestive evidence of this inelasticity, I highlight the existance of ``twin movies'': pairs of films released in close proximity with very similar characteristics.
 The most famous examples, \emph{A Bug's Life} and \emph{Antz}, both released in fall of 1998 and feature early CGI animation of ants rebelling against oppression.
@@
 Given filmmakers' expressed distaste for twin movies, why do they continue to exist?
-Studios are seemingly unable to change their movies' characteristics in response to their competitors' characteristics; at the point DreamWorks learns of \emph{A Bug's Life}, it is too late for them to adjust \emph{Antz} to be more distinct.
+Studios are seemingly unable to change their films' characteristics in response to their competitors' characteristics; at the point DreamWorks learns of \emph{A Bug's Life}, it is too late for them to adjust \emph{Antz} to be more distinct.
 Thus, the demand estimation coefficients are identified--they do not capture the strategic interplay of positioning and substitutability, but instead the direct impact of competition on film sales.
@@
 I think this is unlikely to be a major concern for a couple of reasons.
-First, there are certain prime release dates which are likely to be more profitable than others; for example, releasing a family movie in November is much better than doing so in January, as you can capture the holiday season.
-Thus, if a movie targeted a holiday release, moving its release date to avoid competition would potentially be quite costly since there no longer exists consumer demand.
+First, there are certain prime release dates which are likely to be more profitable than others; for example, releasing a family film in November is much better than doing so in January, as you can capture the holiday season.
+Thus, if a film targeted a holiday release, moving its release date to avoid competition would potentially be quite costly since there no longer exists consumer demand.
 Second, the continued existence of twin movies further suggests this is not a margin firms seem to adjust significantly on: if studios could easily move release dates to avoid competition, we would expect to see fewer twin movies than we do.
@@
 
-An ongoing puzzle within the film industry is the existence of a uniform pricing standard: within a timeslot, cinemas generally charge the same price for all movies regardless of excess demand.
-As a salient example, ``opening night'' showings of movies often sell out, yet theaters do not raise prices.
-Similarly, movies late in their run often have significant excess capacity, yet theathers do not lower prices.
+An ongoing puzzle within the film industry is the existence of a uniform pricing standard: within a timeslot, cinemas generally charge the same price for all films regardless of excess demand.
+As a salient example, ``opening night'' showings of films often sell out, yet theaters do not raise prices.
+Similarly, films late in their run often have significant excess capacity, yet theaters do not lower prices.
 
@@
 
-First, I argue \emph{movies do not compete on price.} Under more flexible pricing, it is possible two movies in close priximity would lower prices to compete for the same audience, raising the quantity sold overall.
+First, I argue \emph{films do not compete on price.} Under more flexible pricing, it is possible two films in close proximity would lower prices to compete for the same audience, raising the quantity sold overall.
 However, since prices are fixed I assume that consumer decisions are driven entirely by product characteristics and idiosyncratic taste.
@@
 
-Movie consumers face a discrete choice problem: which of the movies on offer this weekend should they see?
-Let movie $i$'s attractiveness at time $t$ be given by some $\delta_{it}$.
-Consumers receive idiosyncratic demand shocks for each movie, and choose the movie which maximizes their utility.
-Due to the uniform pricing scheme discussed above, price \emph{does not} enter into the consumer's decision; choices are driven entirely by movie quality and idiosyncratic taste.
+Consumers face a discrete choice problem: which of the films on offer this weekend should they see?
+Let film $i$'s attractiveness at time $t$ be given by some $\delta_{it}$.
+Consumers receive idiosyncratic demand shocks for each film, and choose the option which maximizes their utility.
+Due to the uniform pricing scheme discussed above, price \emph{does not} enter into the consumer's decision; choices are driven entirely by film quality and idiosyncratic taste.
 
-Aggregating over all consumers, we can express the demand for movie $i$ in week $t$ as a function $\phi$ of the movie's appeal $\delta_{it}$ and the appeal of its competitors $\delta_{-it}$.
+Aggregating over all consumers, we can express the demand for film $i$ in week $t$ as a function $\phi$ of the film's appeal $\delta_{it}$ and the appeal of its competitors $\delta_{-it}$.
 Again due to the uniform pricing scheme, we can use revenues as a proxy for quantity, since quantity is simply a constant scalar transformation of revenue.
-Finally, demand for movies varies idiosyncratically by weekend; for example, there is much more demand for movies on Christmas Day than on a random weekend in sunny July.
+Finally, demand for films varies idiosyncratically by weekend; for example, there is much more demand for films on Christmas Day than on a random weekend in sunny July.
 Thus, I include a fixed effect $\alpha_t$ to capture week-by-week variation.
-I can therefore write the demand for movie $i$ in week $t$ as:
+I can therefore write the demand for film $i$ in week $t$ as:
 
@@
 
-A movie's appeal is composed of two parts.
-First, a fixed effect $\alpha_i$ captures the movie's quality.
-For example, we would expect \emph{Star Wars} to have a high $\alpha_i$ to capture the fact it is a popular movie in a vacuum separate from its competitors.
-Second, movies have an age fixed effect $\lambda_{t - r(i)}$, where $r(i)$ represents movie $i$'s release date.
-As an experience good, demand for a movie is driven heavily by novelty; thus, the same movie will experience much less demand in its second week than in its first.
+A film's appeal is composed of two parts.
+First, a fixed effect $\alpha_i$ captures the film's quality.
+For example, we would expect \emph{Star Wars} to have a high $\alpha_i$ to capture the fact it is a popular film in a vacuum separate from its competitors.
+Second, films have an age fixed effect $\lambda_{t - r(i)}$, where $r(i)$ represents film $i$'s release date.
+As an experience good, demand for a film is driven heavily by novelty; thus, the same film will experience much less demand in its second week than in its first.
 While it is reasonable to expect these $\lambda$ to decrease with time, I do not impose this; instead, verifying this is true will be a test of model fit later on.
 Finally, I include an idiosyncratic shock $\xi_{it}$ to capture unobserved demand drivers.
-Thus, the movie's appeal is given by:
+Thus, the film's appeal is given by:
 
@@
 
-How should the demand for movie $i$ depend on its competitors? I put forward three competitor traits which should impact demand for movie $i$:
+How should the demand for film $i$ depend on its competitors? I put forward three competitor traits which should impact demand for film $i$:
 \begin{itemize}
-    \item \emph{Competitor Quality}: If competitor $j$ has a high $\alpha_j$, it should draw more demand away from movie $i$.
-For example, \emph{The Lord of the Rings} is broadly a ``better'' movie than \emph{Willow,} despite being similar on paper; thus, $i$ would prefer to compete against \emph{Willow} rather than \emph{The Lord of the Rings}.
-    \item \emph{Competitor Age}: If competitor $j$ is more recent, it should draw more demand away from movie $i$.
-That is, $i$ should broadly prefer to release against movies in their fifth week of showing than in their first week.
-    \item \emph{Competitor Similarity}: If competitor $j$ is more similar to movie $i$, it should draw away more demand.
+    \item \emph{Competitor Quality}: If competitor $j$ has a high $\alpha_j$, it should draw more demand away from film $i$.
+For example, \emph{The Lord of the Rings} is broadly a ``better'' film than \emph{Willow,} despite being similar on paper; thus, $i$ would prefer to compete against \emph{Willow} rather than \emph{The Lord of the Rings}.
+    \item \emph{Competitor Age}: If competitor $j$ is more recent, it should draw more demand away from film $i$.
+That is, $i$ should broadly prefer to release against films in their fifth week of showing than in their first week.
+    \item \emph{Competitor Similarity}: If competitor $j$ is more similar to film $i$, it should draw away more demand.
 For example, if $i$ is \emph{Star Wars}, it is much more enthusiastic to release against \emph{The Notebook} than against \emph{Star Trek}.
@@
 
-Let $d_{ij}$ be a measure of similarity between movies $i$ and $j$.
+Let $d_{ij}$ be a measure of similarity between films $i$ and $j$.
 In this exercise I use the cosine distance of the embedded text descriptions; however, this could be any measure of similarity.
@@
 Moreover, the $f$ function interacts multiplicatively with the $\delta_{jt}$ terms, making matrix inversion ineffective.
-With some clever rearranging, however, this model becomes more tractable:\footnote{Gory algebraic detail of this rearrangement is available in appendix \ref{app:rearrangement}.}
+With some rearranging, however, this model becomes more tractable:\footnote{Gory algebraic detail of this rearrangement is available in appendix \ref{app:rearrangement}.}
 \begin{align*}
@@
 
-Thus, conditional on the values of $f(d_{ij})$, each movie's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend plus mean-zero noise.
+Thus, conditional on the values of $f(d_{ij})$, each film's log quantity is a linear function of it and its competitors' fixed effects and ages plus a fixed effect for the relevant weekend plus mean-zero noise.
 Thus, the $\lambda$ and $\alpha$ coefficients are estimable using OLS.
@@
 As an example, consider what the data matrix will look like in the following market.
-Movie 1 released this week.
-Movie 2 released this week as well, while movie 3 released last week.
+Film 1 released this week.
+Film 2 released this week as well, while film 3 released last week.
 I omit time fixed effects in this example since we focus on a single period; similarly, I omit time subscripts on the quantities.
-In this case, we will have the following data matrix where values in ``movie'' columns are used to estimate $\alpha$ values while ``age'' columns are used to estimate $\lambda$ values:
+In this case, we will have the following data matrix where values in ``film'' columns are used to estimate $\alpha$ values while ``age'' columns are used to estimate $\lambda$ values:
 
@@
 \begin{tabular}{cccccc}
-    ln(Quantity) & Movie 1 & Movie 2 & Movie 3 & Age 0 & Age 1 \\
+    ln(Quantity) & Film 1 & Film 2 & Film 3 & Age 0 & Age 1 \\
     $\ln(q_{1})$ & 1 & $f(d_{12})$ & $f(d_{13})$ & $1 + f(d_{12})$ & $f(d_{13})$ \\
@@
 
-Note the above is all taking $f$ as given. Following \textcite{magnolfi2022}, let $f$ be a cubic polynomial in distance; that is:
+Note all the above is taking $f$ as given. Following \textcite{magnolfi2022}, let $f$ be a cubic polynomial in distance; that is:
 
@@
 To speed estimation, I first optimize $\gamma$ using data from 2000 alone (the first ``quality'' year of my data).
-Conditional on this optimized $\gamma$, I then estimate $\alpha$ and $\lambda$ on all releases from 2000-2019 (omitting 2020 onward due to the COVID-19 pandemic).
+Conditional on this optimized $\gamma$, I then estimate $\alpha$ and $\lambda$ on all releases from 2000-2019 (omitting 2020 onward due to the negative impact of the COVID-19 pandemic).
 This allows me to estimate the full model in a reasonable amount of time.
@@
 
-In some ways this model is a strange fit for ``characteristic-space'' competition, in that it includes movie fixed effects $\alpha_i$.
+In some ways this model is a strange fit for ``characteristic-space'' competition, in that it includes film fixed effects $\alpha_i$.
 One of the common virtues of using a characteristic-space model is that it allows us to estimate demand for goods which are not yet in the market by arguing based on similar characteristics.
-However, the econometrician cannot estimate $\alpha_i$ for a movie which has not yet released.
+However, the econometrician cannot estimate $\alpha_i$ for a film which has not yet released.
 
@@
 
-Moreover, the \emph{additive separability} of the $\alpha$ and $\lambda$ terms goes grants certain counterfactuals extra relevance.
+Moreover, the \emph{additive separability} of the $\alpha$ and $\lambda$ terms grants certain counterfactuals extra relevance.
 Note in the model above $\alpha$ and $\lambda$ never interact multiplicatively.
 Thus, given competitor similarity $d_{ij}$ and ages $\lambda_{t - r(k)}$, it is possible to perform comparative statics agnostic to the $\alpha$ values.
-For example, I can ask how demand for movie $i$ will compare if it releases in week 1 rather than week 2 conditional on its competitors.
+For example, I can ask how demand for film $i$ will compare if it releases in week 1 rather than week 2 conditional on its competitors.
 Since the terms containing $\alpha_i$ will be identical in these two scenarios, they drop out and leave a concrete estimate of the change in quantity which is not dependent on the $\alpha_i$ values.
@@
 Pandya is a film consultant specializing in releasing Indian films for the North American market.
-For each weekend, the website lists the top 10-20 movies by North American box office revenue.
+For each weekend, the website lists the top 10-20 films by North American box office revenue.
 
-While BoxOfficeGuru is getting its footing in the late 1990s, the data is reported less consistently; thus, I begin my sample in 2000.
-To avoid the impact of the COVID-19 pandemic, I end my sample in 2019.
+While BoxOfficeGuru was getting its footing in the late 1990s, their reports were less consistent; thus, I begin my sample in 2000.
+To avoid the negative shock of the COVID-19 pandemic, I end my sample in 2019.
 When Pandya is on vacation the website is not updated, so I have a few missing weekends in my sample interspersed thoughout my period of interest.
@@
 
-I retrieve movie characteristics from the API of TheMovieDB, a user-generated database of movies.
-Specifically, I clean movie names and years retrieved from BoxOfficeGuru and query the search API for the top result released in the relevant year.
-For each movie, I can thus retrieve the description, genres, and original release language.
-Example movie descriptions are visible in table \ref{tab:tmdb_desc}.
+I retrieve film characteristics from the API of TheMovieDB, a user-generated database of films.
+Specifically, I clean film names and years retrieved from BoxOfficeGuru and query the search API for the top result released in the relevant year.
+For each film, I can thus retrieve the description, genres, and original release language.
+Example film descriptions are visible in table \ref{tab:tmdb_desc}.
 
-TMDB also lists the primary language of each movie.
-I limit only to movies originally released in English to avoid both non-English movie descriptions and differential sales trends among non-English speaking consumers.
-In total,my sample contains 2,970 movies over the 20 year span matched between BoxOfficeGuru and TheMovieDB.
+TMDB also lists the primary language of each film.
+I limit only to films originally released in English to avoid both non-English film descriptions and differential sales trends among non-English speaking consumers.
+In total, my sample contains 2,970 films over the 20 year span matched between BoxOfficeGuru and TheMovieDB.
 
@@
         \toprule 
-        \textbf{Movie} & \textbf{Description} \\
+        \textbf{Film} & \textbf{Description} \\
         \midrule
@@
     \end{tabular} 
-    \caption{Example Movie Descriptions from TheMovieDB}
+    \caption{Example Film Descriptions from TheMovieDB}
     \label{tab:tmdb_desc}
@@
 
-I use OpenAI's embedding model \texttt{text-embedding-3-small} API to embed movie descriptions.
-This produces a 1,536-dimensional vector for each movie description.
+I use OpenAI's embedding model \texttt{text-embedding-3-small} API to embed film descriptions.
+This produces a 1,536-dimensional vector for each film description.
 The OpenAI embedding procedure trains a transformer to predict adjacent texts based on a large corpus, such that vectors with similar cosine similarity are likely to have similar next words \parencite{neelakantan2022ao, kusupati2022ao}.
@@
 I prefer the OpenAI embeddings to other text-as-data methods for a few reasons.
-First, movie descriptions are quite terse; most are around a paragraph, with a few as short as one or two sentences.
-Thus, traditional bag-of-words methods are likely to suffer, as word or bigram counts would be remarkably sparse.
-Because the OpenAI embedding model is trained on an enormous corpus of text, it can infer when descriptions are similar even when they share \emph{no} words--for example, a movie about ``wizardry'' would be similar to one about ``sorcery,'' which would not be true under a bag-of-words model \parencite{brown2020a}.
+First, film descriptions are quite terse; most are around a paragraph, with a few as short as one or two sentences.
+Thus, traditional bag of words methods are likely to suffer, as word or bigram counts would be remarkably sparse.
+Because the OpenAI embedding model is trained on an enormous corpus of text, it can infer when descriptions are similar even when they share \emph{no} words--for example, a film about ``wizardry'' would be similar to one about ``sorcery,'' which would not be true under a bag of words model \parencite{brown2020a}.
 
 Secondly, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level), while the OpenAI embeddings are generally more context-aware.
-Given the movie descriptions tend to be longer than a single sentence, I believe the OpenAI embeddings are likely to be more able to capture the full context of the movie description.
+Given the film descriptions tend to be longer than a single sentence, I believe the OpenAI embeddings are likely to be more able to capture the full context of the film description.
 
@@
 As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and emphasize clustering.
-For each movie, I identify the first genre listed on the TMDB page.
+For each film, I identify the first genre listed on the TMDB page.
 When running t-SNE I limit to only the four most popular genres (action, drama, comedy, and horror) to keep the figure legible.
@@
         \toprule
-        \textbf{Baseline Movie}    & \textbf{Nearest Neighbor \#1}   & \textbf{Nearest Neighbor \#2}       \\ \midrule
+        \textbf{Baseline Film}    & \textbf{Nearest Neighbor \#1}   & \textbf{Nearest Neighbor \#2}       \\ \midrule
         \textit{Frozen}            & \textit{Frozen II}              & \textit{Mirror Mirror}              \\
@@
         \end{tabular}
-        \caption{Nearest Neighbors to Baseline Movies by Cosine Similarity}
+        \caption{Nearest Neighbors to Baseline Films by Cosine Similarity}
         \label{tab:cosine}
@@
 
-Nearest neighbors to some popular movies are visible in table \ref{tab:cosine}.
-Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest movies to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest movies to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
+Nearest neighbors to some popular films are visible in table \ref{tab:cosine}.
+Just by inspection, it is clear this similarity score is capturing \emph{something}: the closest films to \emph{The Notebook} are other Nicholas Sparks book adaptations, and the closest films to \emph{A Bug's Life} are \emph{The Ant Bully} and \emph{Antz}.
 Moreover, the closest neighbors to \emph{Frozen} and \emph{Avengers: Endgame} are in the same franchise: \emph{Frozen II} and \emph{Avengers: Infinity War}, respectively.
-Thus, I feel comfortable using these embeddings to capture similarity between movies.
+Thus, I feel comfortable using these embeddings to capture similarity between films.
 
@@
 First, let us check the age coefficients $\lambda$.
-The movie in my sample with the longest run is \emph{My Big Fat Greek Wedding,} which ran for 43 weeks.
-Estimating a weekly fixed effect for all 43 possible ages would absorb much of my power and create colinearity issues, as few movies reach this age.
+The film in my sample with the longest run is \emph{My Big Fat Greek Wedding,} which ran for 43 weeks.
+Estimating a weekly fixed effect for all 43 possible ages would absorb much of my power and create colinearity issues, as few films reach this age.
 Instead, I topcode age.
@@
 The estimated $\lambda$ coefficients are visible in figure \ref{fig:age_coefs}.
-As hypothesized, the age fixed effects decline with time such that new movies are more popular than old ones.
+As hypothesized, the age fixed effects decline with time such that new films are more popular than old ones.
 Recall that the dependent variable is \emph{log} quantity; thus, the linear decline we observe in the $\lambda$ coefficients implies an exponential decline in sales, as is expected in the film industry.
@@
 \begin{itemize}
-    \item It is negative for similar movies, suggesting that additional competition reduces a movie's revenue
-    \item It is (roughly) monotonic and attenuating with distance, suggesting that as movies become more distant their competitive impact declines
-    \item It approaches zero as distance becomes large, suggesting that movies which are very dissimilar have no impact on each other
-    \item It is concave, suggesting that competition is most important when movies are similar but less so when they are dissimilar
+    \item It is negative for similar films, suggesting that additional competition reduces a film's revenue.
+    \item It is (roughly) monotonic and attenuating with distance, suggesting that as films become more distant their competitive impact declines.
+    \item It approaches zero as distance becomes large, suggesting that films which are very dissimilar have no impact on each other.
+    \item It is concave, suggesting that competition is most important when films are similar but less so when they are dissimilar.
 \end{itemize}
@@
 \begin{itemize}
-    \item If two movies are dissimilar enough, they may be complements rather than substutites.
+    \item If two films are dissimilar enough, they may be complements rather than substitutes.
 For example, if a teenage child wants to see \emph{Friday the 13th}, parents may drop off the child and take the younger sibling to see \emph{Frozen}, when they would not have seen \emph{Frozen} otherwise.
-    \item Theater crowding may push people to see movies they would not have otherwise seen.
+    \item Theater crowding may push people to see films they would not have otherwise seen.
 If a theater is showing \emph{Friday the 13th,} this may crowd out an extra showing of \emph{The Lego Movie}.
@@
 
-We know the average movie's $\alpha_i$ is roughly 0.5.
-Similarly, the $\lambda_{r(i) - t}$ in the week of a movie's premier is roughly 2.5.
-Thus, the average movie's appeal $\delta_{it}$ in its first week is 3.
+We know the average film's $\alpha_i$ is roughly 0.5.
+Similarly, the $\lambda_{r(i) - t}$ in the week of a film's premier is roughly 2.5.
+Thus, the average film's appeal $\delta_{it}$ in its first week is 3.
 
-Suppose movie $i$ is attempting to estimate the impact of competitor $j$'s premier, anticipating $j$ is an average movie.
+Suppose film $i$ is attempting to estimate the impact of competitor $j$'s premier, anticipating $j$ is an average film.
 If $j$ is in the 90th percentile of similarity to $i$, then $i$'s log quantity would be reduced by $f(d_{ij}) \cdot \delta_{jt} = -0.01 \cdot (0.5 + 2.5) = -0.03$, or roughly 3\%.
@@
 
-To put this magnitude in context, \textcite{follows2016SF,follows2016SFa} estimates that box office receipt makes up 42\% of a movie's revenue.
+To put this magnitude in context, \textcite{follows2016SF,follows2016SFa} estimates that box office receipt makes up 42\% of a film's revenue.
 Further (acknowledging the difficulties of managing so-called ``Hollywood accounting''), he argues that the average film earns a profit of roughly 3.7\%.
 If a film experiences a 3.3\% decline in 42\% of its revenue, its total revenue falls by 1.3\%.
-The author documents that the vast majority of a movie's cost is fixed rather than variable; thus, we can assume cost does not change given the presence of competitors.
+The author documents that the vast majority of a film's cost is fixed rather than variable; thus, we can assume cost does not change given the presence of competitors.
 These figures together imply the profit margin in the presence of the closer competitor is only 1.95\%, a 47.3\% reduction in profitability.\footnote{$\frac{R - C}{C} = 0.033 \implies R = 1.033C \implies 0.987 R = 1.0195 C \implies \frac{0.987R - C}{C} = 0.0195 \implies \frac{0.037 - 0.0195}{0.037} = 47.3\%$}
 
-To be even more concrete, let us consider specific movies.
-Ignoring the $\alpha$ coefficients for each of the following movies, suppose Disney were releasing \emph{Frozen} opposite gory historical action flick \emph{Medieval}: a movie in the bottom decile of cosine similarity to \emph{Frozen}.
-If Disney were instead facing children's fantasy romp \emph{Inkheart}--a movie in the top decile of similarity to \emph{Frozen}--they would expect a reduction in revenue similar to the one computed above as \emph{Inkheart} pulls away customers \emph{Medieval} would not.
+To be even more concrete, let us consider specific films.
+Ignoring the $\alpha$ coefficients for each of the following films, suppose Disney were releasing \emph{Frozen} opposite gory historical action flick \emph{Medieval}: a film in the bottom decile of cosine similarity to \emph{Frozen}.
+If Disney were instead facing children's fantasy romp \emph{Inkheart}--a film in the top decile of similarity to \emph{Frozen}--they would expect a reduction in revenue similar to the one computed above as \emph{Inkheart} pulls away customers \emph{Medieval} would not.
 
@@
 
-A skeptic might argue the patterns I see are merely coincidental: why would we believe movies with similar descriptions necessarily appeal to the same audience?
-In this section, I use real movie reviews to validate my text-based model, showing that movies with similar descriptions indeed appeal to similar consumers.
+A skeptic might argue the patterns I see are merely coincidental: why would we believe films with similar descriptions necessarily appeal to the same audience?
+In this section, I use real film reviews to validate my text-based model, showing that films with similar descriptions indeed appeal to similar consumers.
 
 ``Recommendation engines'' are a common tool in the tech world used to predict what products a user will like.
-We can think of this project as harnessing a new type of recommendation engine, identifying movies which are likely to be substitutable based on their descriptions.
+We can think of this project as harnessing a new type of recommendation engine, identifying films which are likely to be substitutable based on their descriptions.
 Algorithms in this space are broadly divided into two categories: collaborative filtering and content-based filtering.
-The text embeddings used above are a form of content-based filtering: using the ``content'' of the movie (its text description) to predict its appeal.
+The text embeddings used above are a form of content-based filtering: using the ``content'' of the film (its text description) to predict its appeal.
 However, one way to benchmark the quality of this approach is to compare against a \emph{collaborative} filtering algorithm.
-Unlike content-based filtering, collaborative filtering doesn't include any information about the content of a movie; instead, it uses the preferences of other users to predict what a user will like.
+Unlike content-based filtering, collaborative filtering doesn't include any information about the content of a film; instead, it uses the preferences of other users to predict what a user will like.
 For example, if I like \emph{The Dark Knight}, and most people who like \emph{The Dark Knight} also like \emph{Inception}, a collaborative filtering algorithm would predict I would like \emph{Inception}.
 
-\textcite{harper2016ATIIS} provide a public-use dataset of real movie reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of movies by users.
-Moreover, movies are tagged with their TMDB ID; that is, the same ID I use to collect movie descriptions.
-Thus, for any given pair of movies, I can compute similarity under both the text-based and collaborative filtering methods.
+\textcite{harper2016ATIIS} provide a public-use dataset of real film reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of films by users.
+Moreover, films are tagged with their TMDB ID; that is, the same ID I use to collect film descriptions.
+Thus, for any given pair of films, I can compute similarity under both the text-based and collaborative filtering methods.
 If the two measures are correlated, I can be more confident that the patterns I see in the text-based model are not coincidental.
@@
 I implement UV-decomposition for matrix completion as described in \textcite{leskovec2020MMD}.\footnote{The algorithm described in the textbook optimizes element-wise, which is quite slow.
-Based on my own vector calculus visible in appendix \ref{app:uv}, I've implemented \emph{matrix-wide} optimization, which seems to work.
-However, I have not found a reference for this approach anywhere; I emailed Professor Leskovec to confirm the result is the same, but have not heard back.} Suppose I have an $n \times m$ matrix of movie reviews, with $n$ users and $m$ reviews.
-However, the matrix is generally sparse; most users have not reviewed most movies.
+Based on my own vector calculus visible in appendix \ref{app:uv}, I've implemented \emph{matrix-wide} optimization, which seems to work as well, but I have not seen a reference to this method elsewhere.} Suppose I have an $n \times m$ matrix of film reviews, with $n$ users and $m$ reviews.
+However, the matrix is generally sparse; most users have not reviewed most films.
 UV-decomposition decomposes this matrix into two thin matrices $U$ and $V$ such that $UV'$ approximates the original matrix.
-This is a cousin of singular value decomposition, however relaxes the orthogonality constraint common in those approaches.
+This is a cousin of singular value decomposition, however it relaxes the orthogonality constraint common in those approaches.
 By minimizing the sum of squared errors among projections in the \emph{observed} entries of the matrix, we hope to project the \emph{unobserved} entries of the matrix as well.
@@
 
-With this decomposition in hand, it is straightforward to assess whether two movies are ``similar:'' we simply compute the cosine similarity of the projected reviews.
-If two movies have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
-The virtue of UV-decomposition is it works \emph{even when the movies have no reviewers in common}; that is, it can infer similarity between movies which have never been reviewed by the same person.
+With this decomposition in hand, it is straightforward to assess whether two films are ``similar'': we simply compute the cosine similarity of the projected reviews.
+If two films have similar projected reviews, we can infer they appeal to similar audiences, and so are likely substitutable.
+The virtue of UV-decomposition is it works \emph{even when the films have no reviewers in common}; that is, it can infer similarity between films which have never been reviewed by the same person.
 
@@
 
-As we might expect, the UV decomposition is better at predicting reviews for movies with more reviews.
+As we might expect, the UV decomposition is better at predicting reviews for films with more reviews.
 To assess the size of this effect, I regress the true review on the predicted review for various thresholds of review count.
-If the threshold is low, we're including movies with fewer reviews in our prediction set; if the threshold is high, we're looking only at hits like \emph{Star Wars.} The results of these regressions are visible in figure \ref{fig:uv_threshold}.
+If the threshold is low, we're including films with fewer reviews in our prediction set; if the threshold is high, we're looking only at hits like \emph{Star Wars.} The results of these regressions are visible in figure \ref{fig:uv_threshold}.
 
 For any value of the review count threshold, the coefficient on predicted reviews is positive and significant.
-This suggests that the UV decomposition is quite predictive of a movie's appeal.
-Moreover, the coefficient is increasing in the review count threshold; that is, the UV decomposition is better at predicting reviews for movies with more reviews.
-This is as expected; the more data we have on a movie, the better we can predict its appeal.
+This suggests that the UV decomposition is quite predictive of a film's appeal.
+Moreover, the coefficient is increasing in the review count threshold; that is, the UV decomposition is better at predicting reviews for films with more reviews.
+This is as expected; the more data we have on a film, the better we can predict its appeal.
 
-These results affirm for me that the collaborative filtering results are likely to be quite predictive of a movie's appeal.
-Thus, if my text-based model identifies the same pairs of movies as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
+These results affirm for me that the collaborative filtering results are likely to be quite predictive of a film's appeal.
+Thus, if my text-based model identifies the same pairs of films as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
 
@@
 
-For movies with relatively few reviews, the coefficient on text similarity is insignificant; this is to be expected, as for these movies UV decomposition struggles to capture anything of meaning.
+For films with relatively few reviews, the coefficient on text similarity is insignificant; this is to be expected, as for these films UV decomposition struggles to capture anything of meaning.
 However, as the review count threshold increases, the coefficient on text similarity becomes increasingly positive and significant.
@@
 This is a strong validation of my text-based model; the patterns I see in the text embeddings are not coincidental, but reflect consumers' true preferences.
-The quality of the match begins to decline when the review threshold becomes too high; however, this is not a fault with my method but with the decomposition, as at these high standards the number of movies with sufficient views falls to less than 10, producing fewer than 100 pairs of movies.
+The quality of the match begins to decline when the review threshold becomes too high; however, this is not a fault with my method but with the decomposition, as at these high standards the number of films with sufficient views falls to less than 10, producing fewer than 100 pairs of films.
 
@@
 
-Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of movies, as implemented by \textcite{sajani2023}.
+Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of films, as implemented by \textcite{sajani2023}.
 PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages.
@@
 
-For my application, I implement TSPR as follows. To find movies similar to specific movie $x$,:
+For my application, I implement TSPR as follows. To find films similar to specific film $x$:
 \begin{itemize}
-    \item The algorithm randomly walks over the bipartite network of users and movies. 
-    \item At each step, the algorithm has a 15\% chance of jumping back to $x$
-    \item If the algorithm does not jump to $x$, it takes a random step to a neighbor of its current node
-    \item Random draws are weighted by the number of stars in the review; e.g., a ``5-star'' review is five times as likely to be drawn as a ``1-star'' review
+    \item The algorithm randomly walks over the bipartite network of users and films. 
+    \item At each step, the algorithm has a 15\% chance of jumping back to $x$.
+    \item If the algorithm does not jump to $x$, it takes a random step to a neighbor of its current node.
+    \item Random draws are weighted by the number of stars in the review; e.g., a ``5-star'' review is five times as likely to be drawn as a ``1-star'' review.
 \end{itemize}
 
-Running this algorithm until convergence provides a measure of how central any given movie $y$ is relative to movie $x$.
-However, we are still not quite done, as some movies are just \emph{inherently} central; if everyone likes \emph{Forrest Gump,} then it's not particularly informative that fans of movie $x$ \emph{also} like \emph{Forrest Gump}.
+Running this algorithm until convergence provides a measure of how central any given film $y$ is relative to film $x$.
+However, we are still not quite done, as some films are just \emph{inherently} central; if everyone likes \emph{Forrest Gump,} then it's not particularly informative that fans of film $x$ \emph{also} like \emph{Forrest Gump}.
 Thus, I divide the TSPR score of $y$ by $y$'s \emph{non}-topic specific PageRank score.
-This gives a measure of how much more central $y$ is to $x$ than it is to the average movie.
+This gives a measure of how much more central $y$ is to $x$ than it is to the average film.
 Given this is a ratio, it has propensity to explode if $y$'s centrality is very small; thus, I prefer to take the log of this ratio for my analysis.
 
-Once I have each movie's TSPR score for each other movie, I can compare these scores to the cosine similarity of the text embeddings.
-Note that the TSPR scores are not symmetric; that is, the TSPR score of movie $x$ to movie $y$ is not necessarily the same as the TSPR score of movie $y$ to movie $x$.
-Thus, for each pair of movies, I actually observe \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
+Once I have each film's TSPR score for each other film, I can compare these scores to the cosine similarity of the text embeddings.
+Note that the TSPR scores are not symmetric; that is, the TSPR score of film $x$ to film $y$ is not necessarily the same as the TSPR score of film $y$ to film $x$.
+Thus, for each pair of films, I actually observe \emph{two} measures of similarity; one from $x$ to $y$ and another from $y$ to $x$; these pairs have the same text-based cosine similarity, but different TSPR values.
 
 A common weakness of collaborative filtering algorithms like TSPR is that they require significant data before they become effective.
-Thus, I compute TSPR values only for movies which receive at least 10,000 reviews in the MovieLens data.
-This leaves me with 258 movies, or 66,564 pairs of movies.
+Thus, I compute TSPR values only for films which receive at least 10,000 reviews in the MovieLens data.
+This leaves me with 258 films, or 66,564 pairs of films.
 To assess the importance of this margin, I then plot the coefficient of regressing TSPR on text similarity for thresholds 10,000 to 50,000.
-I also plot the number of movie pairs included in the sample for each threshold value, to assess the trade-off of ``more ratings per movie'' vs.
-``more movies.'' The results of this plot are visible in figure \ref{fig:tspr_threshold}.
+I also plot the number of film pairs included in the sample for each threshold value, to assess the trade-off of ``more ratings per film'' vs.
+``more films.'' The results of this plot are visible in figure \ref{fig:tspr_threshold}.
 
@@
 This plot shows that the coefficient on text similarity for predicting TSPR is consistently and significantly positive for nearly all thresholds.
-Thus, the content of the text embeddings \emph{does} seem to have predictive power for which movies are similar based on revealed preference; that is, the patterns I see in the text-based model are not coincidental and reflect consumers' true preferences.
+Thus, the content of the text embeddings \emph{does} seem to have predictive power for which films are similar based on revealed preference; that is, the patterns I see in the text-based model are not coincidental and reflect consumers' true preferences.
 
@@
 If we think TSPR increases in quality with data, this suggests that as TSPR becomes higher quality, it looks increasingly like text-based cosine similarity.
-This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per movie to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.
+This exercise highlights an additional strength of my cosine similarity measure; while collaborative filtering needs tens of thousands of reviews per film to generate useful predictions, the embeddings-based method can generate useful predictions with only a few sentences of text.
 
@@
 In summary, I have developed a new method for estimating competition in characteristic space.
-By developing a new model of competition in the film industry, I can harness boilerplate OpenAI embeddings to estimate impact of competition on a movie's box office revenue.
+By developing a new model of competition in the film industry, I can harness boilerplate OpenAI embeddings to estimate impact of competition on a film's box office revenue.
 
@@
 
-Movies with similar descriptions are indeed substitutes; moreover, substitutability falls as distance in characteristic space increases.
-My magnitudes are large but reasonable; changing one competitor movie from the bottom decile of similarity to the top decile can reduce revenue by 3.3\%, decreasing total profitability by 47\%.
+Films with similar descriptions are indeed substitutes; moreover, substitutability falls as distance in characteristic space increases.
+My magnitudes are large but reasonable; changing one competitor film from the bottom decile of similarity to the top decile can reduce revenue by 3.3\%, decreasing total profitability by 47\%.
 
@@
     &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \sum_{r(k)} \underbrace{\left(\lambda_{t - r(k)} \sum_{j \neq i}^{r(j) = r(k)} f(d_{ij}) \right)}_{\text{$\lambda_{t - r(k)}$ constant within group}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
-    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \underbrace{\lambda_{t - r(i)} \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij}) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(k)} \sum_j^{r(j) = r(k)} f(d_{ij}) \right)}_{\text{Break the term for $r(i)$, movies of age equal to $i$'s, out of sum}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
+    &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \lambda_{t - r(i)} + \underbrace{\lambda_{t - r(i)} \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij}) + \sum_{r(k) \neq r(i)} \left(\lambda_{t - r(k)} \sum_j^{r(j) = r(k)} f(d_{ij}) \right)}_{\text{Break the term for $r(i)$, films of age equal to $i$'s, out of sum}} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t \\
     &= \alpha_i + \sum_{j \neq i} f(d_{ij})\alpha_j + \underbrace{\lambda_{t - r(i)} \left(1 + \sum_{j \neq i}^{r(j) = r(i)} f(d_{ij})\right)}_{\text{Group matching $\lambda_{t - r(i)}$}} + \sum_{r(k) \neq r(i)}\left(\lambda_{t - r(k)} \sum_j^{r(j) = r(k)} f(d_{ij}) \right) + \xi_{it} + \sum_{j \neq i} f(d_{ij}) \xi_{jt} + \alpha_t
@@
 \begin{itemize}
-    \item First, we break up the sum and rearrange the equation so the $\alpha$ terms are together, the $\lambda$ terms are together, and the $\xi$ terms are together
-    \item Next we break down the sum over competitor $\lambda$ terms into two sums; the outer sum is over the age of the competitor, while the inner sum is over the competitors of that age
-    \item From there we pull out the element of the outer sum corresponding movies of the same age as movie $i$
-    \item Finally we can pull $\lambda_{t - r(i)}$ out of the combined sum of movie $i$'s age coefficient and its competitors' weighted age coefficients
+    \item First, we break up the sum and rearrange the equation so the $\alpha$ terms are together, the $\lambda$ terms are together, and the $\xi$ terms are together.
+    \item Next, we break down the sum over competitor $\lambda$ terms into two sums; the outer sum is over the age of the competitor, while the inner sum is over the competitors of that age.
+    \item From there we pull out the element of the outer sum corresponding films of the same age as film $i$.
+    \item Finally we can pull $\lambda_{t - r(i)}$ out of the combined sum of film $i$'s age coefficient and its competitors' weighted age coefficients.
 \end{itemize}
@@
 These matrix forms are particularly appealing since they highlight the procedure's implicit relation to linear regression.
-Philosophically, we are searching for $d$ latent characteristics of each movie (corresponding to rows of $V$) and the analogous coefficients on these characteristics which define each user's taste (corresponding to rows of $U$) such that the coefficient-characteristic pair justify the existing reviews.
-From there, we can extrapolate to missing movies by simply multiplying the latent characteristics by the coefficients for each missing user.
-Note since the problem is symmetric we could just as easily say the latent values for users are the ``characteristics'' and the values for movies are ``coefficients,'' but I prefer to think of movies as more easily defined by a finite set of characteristics than users.
+Philosophically, we are searching for $d$ latent characteristics of each film (corresponding to rows of $V$) and the analogous coefficients on these characteristics which define each user's taste (corresponding to rows of $U$) such that the coefficient-characteristic pair justify the existing reviews.
+From there, we can extrapolate to missing films by simply multiplying the latent characteristics by the coefficients for each missing user.
+Note since the problem is symmetric we could just as easily say the latent values for users are the ``characteristics'' and the values for films are ``coefficients,'' but I prefer to think of films as more easily defined by a finite set of characteristics than users.
 

### 2024-09-23 draft/draft.tex
Final tweaks before 2YP submission

@@
 \usepackage{appendix}
+\usepackage{hyperref}
 
@@
 
-Demand estimation methods are central pillars of the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
-However, these approaches are often limited to markets in which goods have quantifiable characteristics such as storage in the market for hard drives or wattage in the market for lightbulbs.
+Demand estimation methods are central to the economist toolkit, informing antitrust regulation, optimal taxation, and trade policy.
+However, these approaches are often confined to markets in which goods have quantifiable characteristics, such as storage in the market for hard drives or wattage in the market for lightbulbs.
 Markets with difficult-to-quantify characteristics should not receive less attention simply because they are difficult-to-quantify; though it is hard to put a number on how funny a film is, we understand that a studio monopolizing ``funny'' films is likely to be bad for consumers.
@@
 
-Product-level demand systems such as the Almost Ideal Demand System (AIDS) can estimate rich substitution patterns without any characteristic data by looking only at substitution patterns under variation; that is, the substitution patterns of good $i$ are uninformative about the patterns of good $j$, even if $i$ and $j$ are actually very similar \parencite{deaton1980AER}.
-These methods often require onerous quantities of data on consumer behavior.
-For example, estimation of AIDS requires sufficient variation to estimate parameters quadratic in the number of goods.
-A corollary of this is that these methods require goods to have \emph{already entered the market}.
-If a firm wishes to estimate substitution for a new product, these methods will have little to offer them.
+Product-level demand systems such as the Almost Ideal Demand System (AIDS) can estimate rich substitution patterns without any characteristic data by looking directly at substitution between each pair of goods $i$ and $j$ \parencite{deaton1980AER}.
+However, these methods often require onerous quantities of data on consumer behavior; for example, estimation of AIDS requires sufficient variation to estimate parameters quadratic in the number of goods.
+Moreover, even if goods $i$ and $j$ are very similar, these methods infer nothing about good $j$ from the substitution patterns of good $i$.
+A corollary of these limitations is that these methods require goods to have \emph{already entered the market}, since without sales data on good $i$ it is impossible to say anything about its substitution patterns.
+If a firm wishes to estimate substitution for a new product, these methods will have little to offer.
 
@@
 
-Using data from BoxOfficeGuru and TheMovieDB, I estimate the impact of competitor characteristics on film revenue for each weekend from 2000-2019.
-I find substitution patterns which are plausible in both magnitude and direction, and which are robust to the choice of distance metric.
+Using data from BoxOfficeGuru and TheMovieDB, I estimate the impact of competitor similarity on film revenue in North America for each weekend from 2000-2019.
+I identify substitution patterns which are plausible in both magnitude and direction.
 As a film becomes more similar to its competitor, it becomes more and more detrimental to the competitor's revenue.
@@
 First, there are certain prime release dates which are likely to be more profitable than others; for example, releasing a family film in November is much better than doing so in January, as you can capture the holiday season.
-Thus, if a film targeted a holiday release, moving its release date to avoid competition would potentially be quite costly since there no longer exists consumer demand.
+Thus, if a film targeted a holiday release, moving its release date to avoid competition would potentially be quite costly since other weekends face lower demand.
 Second, the continued existence of twin movies further suggests this is not a margin firms seem to adjust significantly on: if studios could easily move release dates to avoid competition, we would expect to see fewer twin movies than we do.
@@
 \subsection{Construction}
+
+\label{sec:model_construction}
 
@@
 
-Regressing quantity on the given Xs identifies the coefficients for $\alpha_1$, $\alpha_2$, $\alpha_3$, $\lambda_0$, and $\lambda_1$ via the first through fifth columns, respectively.
+Regressing quantity on the given Xs identifies the coefficients for $\alpha_1$, $\alpha_2$, $\alpha_3$, $\lambda_0$, and $\lambda_1$ via the second through sixth columns, respectively.
 These results make some amount of intuitive sense; if $\ln(q_1)$ performs below expectations, one explanation is that $\alpha_2$ is large--represented by a large coefficient on the (negative) $f(d_{12})$.
 
-Note all the above is taking $f$ as given. Following \textcite{magnolfi2022}, let $f$ be a cubic polynomial in distance; that is:
+Note all the above is taking $f$ as given. Following \textcite{mmagnolfi2024AM}, let $f$ be a cubic polynomial in distance; that is:
 
@@
 
-This functional form makes few assumptions about the relationship between distance and substitutability, allowing for flexible sign, monotonicity, and concavity.
-Thus, if the final result \emph{does} exhibit those properties, we know it is coming from the data rather than baked into the model.
+This functional form makes few assumptions about the relationship between distance and substitutability, allowing for flexible sign, monotonicity, intercept, level, and concavity.
+Thus, if the final result \emph{does} exhibit those properties, we know they are coming from the data rather than baked into the model.
 
@@
 We know the resulting model minimizes mean squared error \emph{in the set of models using this $\gamma$}.
-By adjusting the $\gamma$ vector to minimize mean squared error of these optimized models, I search across the lower envelope and know my final model minimizes mean squared error over the whole $\gamma, \alpha, \lambda$ space.
+By adjusting the $\gamma$ vector to minimize mean squared error of these optimized models, I search across the lower envelope and know my final model minimizes mean squared error over the whole $(\gamma, \alpha, \lambda)$ space.
 
@@
 
-TMDB also lists the primary language of each film.
+TheMovieDB also lists the primary language of each film.
 I limit only to films originally released in English to avoid both non-English film descriptions and differential sales trends among non-English speaking consumers.
@@
 The OpenAI embedding procedure trains a transformer to predict adjacent texts based on a large corpus, such that vectors with similar cosine similarity are likely to have similar next words \parencite{neelakantan2022ao, kusupati2022ao}.
+These embeddings are quite cheap; the bill for this project was less than \$1.
+Thus, this method should be accessible to researchers with any level of resources even for projects of significant scale.
 
@@
 
-Secondly, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level), while the OpenAI embeddings are generally more context-aware.
+Second, BERT embeddings tend to function at the sentence level (and word2vec embeddings function only at the word level), while the OpenAI embeddings are generally more context-aware.
 Given the film descriptions tend to be longer than a single sentence, I believe the OpenAI embeddings are likely to be more able to capture the full context of the film description.
@@
 As a check for the quality of the embeddings, I perform t-Stochastic Neighbor Embedding (t-SNE) on the embeddings to reduce them to two dimensions and emphasize clustering.
-For each film, I identify the first genre listed on the TMDB page.
+For each film, I identify the first genre listed on the TheMovieDB page.
 When running t-SNE I limit to only the four most popular genres (action, drama, comedy, and horror) to keep the figure legible.
@@
 
-First, let us check the age coefficients $\lambda$.
+First, I examine the age coefficients $\lambda$.
 The film in my sample with the longest run is \emph{My Big Fat Greek Wedding,} which ran for 43 weeks.
@@
 Recall that the dependent variable is \emph{log} quantity; thus, the linear decline we observe in the $\lambda$ coefficients implies an exponential decline in sales, as is expected in the film industry.
-Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are monotonically decreasing suggests my model is well specified.
+Again, the model does not impose either the sign or the shape of the $\lambda$ coefficients; the fact that they are monotonically decreasing at a linear rate suggests my model is well specified.
 
@@
 
-Now we can check the $\alpha$ coefficients.
-The distribution of $\alpha$ coefficients is visible in figure \ref{fig:alpha_dist}.
+Now I check the distribution of $\alpha$ coefficients, visible in figure \ref{fig:alpha_dist}.
 The distribution is roughly symmetrical with a mean of 0.5.
@@
 \textcite{harper2016ATIIS} provide a public-use dataset of real film reviews (``MovieLens''); it contains more than 25 million 1-to-5 star reviews of films by users.
-Moreover, films are tagged with their TMDB ID; that is, the same ID I use to collect film descriptions.
+Moreover, films are tagged with their TheMovieDB ID; that is, the same ID I use to collect film descriptions.
 Thus, for any given pair of films, I can compute similarity under both the text-based and collaborative filtering methods.
@@
 I validate my text-based model by comparing it against two frontier collaborative filtering algorithms: UV decomposition and topic-specific PageRank.
-
-\pagebreak
 
@@
 To assess the size of this effect, I regress the true review on the predicted review for various thresholds of review count.
-If the threshold is low, we're including films with fewer reviews in our prediction set; if the threshold is high, we're looking only at hits like \emph{Star Wars.} The results of these regressions are visible in figure \ref{fig:uv_threshold}.
+If the threshold is low, we're including films with fewer reviews in our prediction set; if the threshold is high, we're looking only at hits like \emph{Star Wars.} The results of these regressions are visible in figure \ref{fig:actual_vs_predicted}.
 
@@
 
-These results affirm for me that the collaborative filtering results are likely to be quite predictive of a film's appeal.
+These results affirm that the collaborative filtering estimates are likely to be quite predictive of a film's appeal.
 Thus, if my text-based model identifies the same pairs of films as substitutable, I can be more confident that the patterns I see reflect consumers' true preferences.
@@
 
-Harnessing the bipartite network nature of the reviews data, I use a topic-specific PageRank (TSPR) algorithm to compute similarity between each pair of films, as implemented by \textcite{sajani2023}.
+As a second validation, I harness the bipartite network nature of the reviews data.
+Using topic-specific PageRank (TSPR) as implemented by \textcite{sajani2023}, I compute a second review-based similarity between each pair of films.
 PageRank is an algorithm for detecting the most important nodes in a network; it was originally developed by Google to rank webpages.
@@
 
-% \section*{Acknowledgments}
-% Thank you to Liran Einav and Shoshana Vasserman for their support both in and after IO III.
-Thanks also to Arun Chandrasekhar, Nick Bloom, Janet Stefanov, Nick Scott-Hearn, and Lauren Harris for their feedback on this project, as well as the rest of the 2024 IO III class.
+\section*{Acknowledgments}
+Thank you to Liran Einav and Shoshana Vasserman for their support both in and after Industrial Organization III.
+Thanks also to Janet Stefanov, Nick Scott-Hearn, and Lauren Harris for their feedback on this project, as well as the rest of the 2024 Industrial Organization III class.
 
@@
 
-Let us start with the expression of the model as given in the model construction section: $$\ln(q_{it}) = \alpha_i + \lambda_{t - r(i)} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) (\alpha_j + \lambda_{t - r(j)} + \xi_{jt}) + \alpha_{t}$$
+Let us start with the expression of the model as given in section \ref{sec:model_construction} (``Construction''): $$\ln(q_{it}) = \alpha_i + \lambda_{t - r(i)} + \xi_{it} + \sum_{j \neq i} f(d_{ij}) (\alpha_j + \lambda_{t - r(j)} + \xi_{jt}) + \alpha_{t}$$
 
@@
 Philosophically, we are searching for $d$ latent characteristics of each film (corresponding to rows of $V$) and the analogous coefficients on these characteristics which define each user's taste (corresponding to rows of $U$) such that the coefficient-characteristic pair justify the existing reviews.
-From there, we can extrapolate to missing films by simply multiplying the latent characteristics by the coefficients for each missing user.
+From there, we can interpolate missing films by simply multiplying the latent characteristics by the coefficients for each missing user.
 Note since the problem is symmetric we could just as easily say the latent values for users are the ``characteristics'' and the values for films are ``coefficients,'' but I prefer to think of films as more easily defined by a finite set of characteristics than users.
